### 2020-10-05 msccls/MScTemplateCS.tex
Initial Overleaf Import

\documentclass[nofilelist]{cslthse-msc}
% to show a list of used packages at the end of the document, delete the nofilelist option
%\documentclass{cslthse-msc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
%\usepackage{amsfonts}
%%\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[titletoc, header, page]{appendix}
\usepackage{transparent}
\usepackage{natbib}

% used to display the used files at the end. Select nofilelist as a package option to disable this
\listfiles % initialize

%\geometry{showframe}
%better like this?
%\student{Flavius Gruian}{Flavius.Gruian@cs.lth.se}
\students{Nick Persson}{dat14npe@student.lu.se}{Filip Karabeleski}{dat14fka.student.lu.se}

\thesisnumber{LU-CS-EX: 2020-XX} % Birger Swahn will provide this number to you, once the thesis is ready for publication
% default is Master. Uncomment the following for "kandidatarbete"/Bachelor's thesis
%\thesistype{Bachelor}{Kandidatarbete}

%\title{Formatting a Master's Thesis}
\title{Keyword Categorization and Emotion Analysis in Voice Calls}

%\onelinetitle
%\twolinestitle
\threelinestitle
%\fourlinestitle

\subtitle{Subtitle WIP}
\company{Telavox AB}
\supervisors
{Sara Lindgren, \href{mailto:Sara.Lindgren@telavox.com}{\texttt{Sara.Lindgren@telavox.com}}}
{Pierre Nugues, \href{mailto:Pierre.Nugues@cs.lth.se}{\texttt{Pierre.Nugues@cs.lth.se.com}}}
\examiner
{Flavius Gruian,\href{mailto:Flavius.Gruian@cs.lth.se}{\texttt{Flavius.Gruian@cs.lth.se}}}

\date{\today}
%\date{January 16, 2015}

\acknowledgements{
If you want to thank people, do it here, on a separate right-hand page. Both the U.S. \textit{acknowledgments} and the British \textit{acknowledgements} spellings are acceptable.

We would like to thank Lennart Andersson for his feedback on this template.

We would also like thank Camilla Lekebjer for her contribution on this template, as well as Magnus Hultin for his popular science summary class and example document.

Thanks also go to the following (former) students for helping with feedback and suggestions on this template: Mikael Persson, Christoffer Lundgren, Mahmoud Nasser.
}

\theabstract{
This document describes the Master's Thesis format for the theses carried out at 
the Department of Computer Science, Lund University.

Your abstract should capture, in English, the whole thesis with focus on the problem and solution in 150 words. It should be placed on a separate right-hand page, with an additional \textit{1cm} margin on both left and right. Avoid acronyms, footnotes, and references in the abstract if possible.

Leave a \textit{2cm} vertical space after the abstract and provide a few keywords relevant for your report. Use five to six words, of which at most two should be from the title.
}

\keywords{Machine Learning, NLP, Sentiment Analysis, Keyword Extraction, (style, structure)}

%% Only used to display font sizes
\makeatletter
\newcommand\thefontsize[1]{{#1 \f@size pt\par}}
\makeatother
%%%%%%%%%%

\begin{document}
\renewcommand{\bibname}{References}

\makefrontmatter
\chapter{Introduction}

\section{Background}
Natural language processing (NLP) is a field of study mainly in computer science that deals with analyzing human language in a way that computers can understand and process. According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notations. Since natural languages continuously develop as time goes on and its adherence to structures diminish, it is difficult to define explicit rules to follow.

The LSTM architecture \citep{ntlk2009} is something....

\begin{itemize}
    \item Förklara vad systemet är
    \item Berätta om företaget Telavox
\end{itemize}

\section{Purpose}
\begin{itemize}
    \item Varför behöver problemet lösas, nytta
\end{itemize}
\subsection{Telavox}
Telavox is a telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service. 
\section{Research questions}
The primary aim of this thesis is to investigate whether or not voice call analysis using NLP can be effective enough to be used as a replacement for manual analysis. 
\textit{What level of precision can be reached with the categorization of keywords from voice calls? }

\textit{How well will our algorithm perform compared to other algorithms carrying out the same measurements?}

\textit{What level of effectiveness can be reached compared to a human manually analyzing the data?} 
\section{Terminology}
NLP, ML, Speech-to-text och (STT), API. Kan också kallas för Glossary and Abbreviations
\section{Delimitations}
Delimitations are choices made by the researcher which should be mentioned. They describe the boundaries that you have set for the study.
\section{Contributions}
Potentialen av applications 
\section{Related work}

\chapter{Approach}

\begin{itemize}
    \item Dataset till input måste följa Googles standard som de följer vid speech to text. 
\end{itemize}

\section{Theory}
NLP technique, categorization, ML, supervised/unsupervised,  
Testing the backup by writing this, hope this works.
\section{Method} Ta eventuellt bort
Word level one hot encoding, kanske relevant, avgör beroende på hur mycket mer vi behöver skriva.
\section{Implementation} Ta eventuellt bort
Overfitting \\
Data Augmentation \\
importance of finding the correct data-sets \\
\section{Bibtesting, will be removed}
\citep{emotionlinesdataset} \\
\citep{franoischollet2017learning}\\
\citep{beaver2020towards}
\chapter{Speech-to-text}
Vi antar att det är 2 speakers i samtalet, nämn för- och nackdelar. Ta med i limitations
Vi har endast stöd för engelska pga limits i datamängd, dokumentation. 
Google diarization är i Beta.
Använda kategoriserade ord i Googles STTs speechContexts och därefter använda ML igen för att få högre precision på kategorisering. 
Sample rates 8000 vs 48000 t.ex.

Cased eller uncased? Hur fungerar google speech-to-text med cased/uncased?

\chapter{Keyword Extraction}
\chapter{Emotion Analysis}
6 eller 4 känslor
Sentiment analysis, varför valde vi inte det?

\chapter{Evaluation}

\section{Results}
\section{Discussion}

\chapter{Conclusions}

% Should use consistent formatting when it comes to Names ("FirstName LastName", or "F. LastName")
%\printbibliography
\makebibliography{MyMSc}

\begin{appendices}
\chapter{Dummy Appendix}

% display used packages information unless nofilelist is used in the cslthse-msc package option
\printfilelist

%make sure we're on even page with the pop-sci
\checkoddpage
\ifoddpage
\else
   \newpage
   \thispagestyle{empty}
   \mbox{ }
\fi
\includepdf[pages={1}]{popsci/popsci.pdf}
\end{appendices}

\end{document}

### 2020-11-13 msccls/MScTemplateCS.tex
removed unnecessary garbage
Wrote a bit on theory

@@
 \usepackage{natbib}
+\usepackage{wrapfig}
 
@@
 
+\subsection{Emotion Recognition}
+Emotion has always been a key part of human language, and it therefore comes naturally that teaching a computer to understand emotion in human language is a big part of NLP.
+Initially, the focus lied in sentiment analysis, a field of study focused on determining the binary polarization of piece of text as either positive or negative.
+Godtycklig bit om sentiment analysis.
+Emotion recognition is an extension of sentiment analysis that instead of just determining the polarization of text, attributes emotion to it.
+This process is more challenging than sentiment analysis, as the differences between emotions are subtler than that of just positive and negative.
+Emotion recognition has also been extended to work on different modalities, such as for example visual and auditory input.
+
+Emotion recognition can be applied to data in various modalities, 
+such as visual, in the form of facial expressions and body language, audible and text based in the form of dialogues. To o
+
 The LSTM architecture \citep{ntlk2009} is something....
@@
 \section{Research questions}
-The primary aim of this thesis is to investigate whether or not voice call analysis using NLP can be effective enough to be used as a replacement for manual analysis. 
-\textit{What level of precision can be reached with the categorization of keywords from voice calls? }
+The primary aim of this thesis is to investigate whether or not voice call analysis using NLP can be effective enough to be used as a replacement for manual analysis. \\
+\textit{What level of precision can be reached with the categorization of keywords from voice calls?}\\
+\textit{How well will our algorithm perform compared to other algorithms carrying out the same measurements?}\\
+\textit{What level of effectiveness can be reached compared to a human manually analyzing the data?}
 
-\textit{How well will our algorithm perform compared to other algorithms carrying out the same measurements?}
-
-\textit{What level of effectiveness can be reached compared to a human manually analyzing the data?} 
 \section{Terminology}
-NLP, ML, Speech-to-text och (STT), API. Kan också kallas för Glossary and Abbreviations
+NLP, ML, Speech-to-text och (STT), API. Modality. Kan också kallas för Glossary and Abbreviations
 \section{Delimitations}
@@
 
+\chapter{Theory}
+\section{Machine Learning}
+Machine learning is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program it.
+By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem.
+In \textbf{supervised learning}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number.
+The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs information that allows deducing the label y for this feature vector.)
+The other main type of tasks in machine learning is \textbf{unsupervised learning}.
+The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector.
+The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs either another vector or a value.
+
+(Källor Andriy Burkov, Chollet)
+
+\subsection{Neural Networks and Deep Learning}
+
+Neural networks are machine learning systems loosely inspired by biological neural networks (Chollet s.8). A neural network is a collection of connected nodes or neurons that
+
+\begin{quote}
+An ANN is based on a collection of connected units or nodes called artificial neurons, which loosely model the neurons in a biological brain.
+Each connection, like the synapses in a biological brain, can transmit a signal to other neurons.
+An artificial neuron that receives a signal then processes it and can signal neurons connected to it.
+\end{quote}
+
+Deep learning = multiple layers of neural networks
+
+\section{Word Vectors}
+Word vectors are a type of word representation where texts are turned into numbers
+\subsection{One-Hot Encoding}
+The most basic way of creating a word vector is the one-hot encoding method.
+This creates a very big and sparse $n \cdot n$ matrix where $n$ is the number of words in the embedding.
+Each word is then represented by an array filled with zeroes except for one index set to one.
+For example, the one-hot encoding of the phrase \textit{Why was six afraid of seven} would start with the assignment of values to word:
+
+\begin{center}
+    \begin{tabular}{c|c|c|c|c|c}
+
+         Why & was & six & afraid & of & seven \\
+         \hline
+         1 & 2 & 3 & 4 & 5 & 6 \\
+    \end{tabular}
+\end{center}
+
+This would then be converted into the following matrix:
+
+\begin{center}
+    \begin{tabular}{lllllll}
+                            & 1 & 2 & 3 & 4 & 5 & 6 \\ \cline{2-7} 
+\multicolumn{1}{l|}{Why}    & 1 & 0 & 0 & 0 & 0 & 0 \\
+\multicolumn{1}{l|}{was}    & 0 & 1 & 0 & 0 & 0 & 0 \\
+\multicolumn{1}{l|}{six}    & 0 & 0 & 1 & 0 & 0 & 0 \\
+\multicolumn{1}{l|}{afraid} & 0 & 0 & 0 & 1 & 0 & 0 \\
+\multicolumn{1}{l|}{of}     & 0 & 0 & 0 & 0 & 1 & 0 \\
+\multicolumn{1}{l|}{seven}  & 0 & 0 & 0 & 0 & 0 & 1
+    \end{tabular}
+\end{center}
+
+As mentioned, this method is very basic and creates an unnecessarily large matrix that is largely filled up with nothing.
+If we wanted to represent a vocabulary of 20 000 words, each word would be represented by 19 999 zeroes and a single one.
+
+The other big issue with it is that no meaning can be extracted from the matrix.
+If our vocabulary contained the words "cat", "dog" and "hamster", we could easily see that these are all domestic animals but in the space of the one hot vector, "cat" is as close to "dog" as it would be to "airplane".
+
+\subsection{Word Embedding}
+To resolve the two aforementioned problems with one-hot encodings, word embeddings were created, also known as dense word vectors. \citep{neuralnetworkmethods} 
+Here, instead of creating a sparse, hardcoded vector with equal distance between words, they are learned from data and packed into a lower-dimensional vector where each dimension contains some relevant data about the word.
+According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional or 1024-dimensional are common when dealing with very large vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use one-hot encoding.
+Additionally, the geometric relationship between words can and should reflect the semantic relationship between those same words.
+This means that synonyms and words often used together are embedded closer to each other, while words seldom used together are more distant in the embedding.
+In addition to distance, direction could also be a useful tool to extract meaning from the embedding.\newpage
+
+\begin{wrapfigure}{r}{0.25\textwidth}
+\includegraphics[width=0.9\linewidth]{msccls/embedding_direction.png} 
+\caption{Small example of a word embedding space}
+\label{fig:direction}
+\end{wrapfigure}
+
+In figure \ref{fig:direction}, the four words \textit{cat}, \textit{dog}, \textit{wolf} and \textit{tiger} are represented on a geometric plane.
+We can clearly see some of the possible geometric transformations possible by using direction as a tool.
+For instance, the same vector that takes us from \textit{dog} to \textit{wolf} also takes us from \textit{cat} to \textit{tiger}.
+This could be interpreted this vector as "converting a domesticated animal to its wildlife equivalent".
+Another vector will take us both from \textit{cat} to \textit{dog} and from \textit{tiger} to \textit{wolf} which could be interpreted as "converting a feline to a canine".
+Two meaningful transformation vectors that are common in the real world are "gender" and "plural" vectors.
+By applying a "female" vector to \textit{king}, \textit{queen} is obtained and the same vector returns \textit{actress} if instead used on \textit{actor}.
+"Plural", as should seem obvious by now, if used on \textit{actor} will return \textit{actors}.
+Typically there are thousands of these types of interpretable direction vectors in a large word embedding.
+
+This bit is after the picture, that should be to the right of the last paragraph
+
+\subsection{RoBERTa}
+
+\section{Evaluation}
+\subsection{F1-score, precision, accuracy}
 \chapter{Approach}
 
+\section{Baseline - COSMIC}
 \begin{itemize}
@@
 NLP technique, categorization, ML, supervised/unsupervised,  
-Testing the backup by writing this, hope this works.
-\section{Method} Ta eventuellt bort
+Mechanical turks to 
+Finns både text och audio på MELD, \citep{zhang2019modeling} visar att text för sig är bättre än ljud men att text+ljud > text
+train, dev(validation), test, man testar på dev under development, först på test efter man är klar
+\section{Datasets}
+\subsection{Approach} 
+\section{Method}
+Google Colab
 Word level one hot encoding, kanske relevant, avgör beroende på hur mycket mer vi behöver skriva.
-\section{Implementation} Ta eventuellt bort
+\section{Implementation}
 Overfitting \\

### 2020-11-13 msccls/popsci/popsci.tex
Wrote a bit on theory

% USE PDFLatex!
% to correctly render Swedish characters

\documentclass{popsci}

\usepackage[utf8]{inputenc}
\usepackage[swedish, english]{babel}

\usepackage{fancyhdr}
\usepackage{titling}
\usepackage{color}
\usepackage{colortbl}
\usepackage{graphicx}
\usepackage{flushend}
\usepackage{lmodern}

% Please specify the presentation date
\presentationsdag{2016-04-28}

% use either of these commands to specify the title of your thesis
% \examensarbete{}
% To create a title in two rows, leave examensarbete blank and fill in examensarbeteTwoRows.
\examensarbeteTwoRows{Application Specific Instruction-set Processor Using a Parametrizable multi-SIMD}{Synthesizeable Model Supporting Design Space Exploration}
\student{Nick Persson, Filip Karabeleski}
%\students{Magnus Hultin}{Mr X}
\supervisor{Pierre Nugues (LTH)}
\examiner{Flavius Gruian (LTH)}

% Your pop-sci title should be different (more catchy) than your thesis title
\title{Parametrisk processor modell för design utforskning}

\begin{document}

% not more than 4 rows!
\theabstract{Applikations-specifika processorer är allt mer vanligt för få ut rätt prestanda med så lite resurser som möjligt. Detta arbete har en parametrisk modell för att kunna testa hur mycket resurser som behövs för en specifik applikation.}

{\noindent För att öka prestandan i dagens processorer finns det vektorenheter och flera kärnor i processorer. Vektorenheten finns till för att kunna utföra en operation på en mängd data samtidigt och flera kärnor gör att man kan utföra fler instruktioner samtidigt. Ofta är processorerna designade för att kunna stödja en mängd olika datorprogram. Detta resulterar i att det blir kompromisser som kan påverka prestandan för vissa program och vara överflödigt för andra. I t.ex. videokameror, mobiltelefoner, medicinsk utrustning, digital kameror och annan inbyggd elektronik, kan man istället använda en processor som saknar vissa funktioner men som istället är mer energieffektiv. Man kan jämföra det med att frakta ett paket med en stor lastbil istället för att använda en mindre bil där samma paketet också skulle få plats.

I mitt examensarbete har jag skrivit en modell som kan användas för att snabbt designa en processor enligt vissa parametrar. Dessa parametrar väljs utifrån vilket eller vilka program man tänkta köra på den. Vissa program kan t.ex. lättare använda flera kärnor och vissa program kan använda korta eller längre vektorenheter för dess data.

%För att kunna välja vilken typ av processor som är rätt för den specifika applikationen krävs det ofta att man snabbt kan testa olika prototyper. Att implementera dessa till hårdvara kan ofta vara tidskrävande och ifall det visar sig att implementationen inte klarar dem kraven man ställt för prestanda och energieffektivitet, måste man designa för nya parametrar och mer tid har blivit slösat. %Om den här processen istället kan göras automatiskt utifrån dessa design-parametrar kan man teoretiskt spara en massa tid.
Modellen testades med olika multimedia program. Den mest beräkningsintensiva och mest upprepande delen av programmen användes. Dessa kallas för kärnor av programmen. Kärnorna som användes var ifrån MPEG och JPEG, som används för bildkomprimering och videokomprimering.

\begin{figure}[!bth] % Use pictures in your pop-vet!
\includegraphics[width=\columnwidth]{samplePic.pdf} 
%\caption{En fin bild}
\end{figure}
Resultatet visar att det finns en prestanda vinst jämfört med generella processorer men att detta också ökar resurserna som behövs. Detta trots att den generella processorn har nästan dubbelt så hög klockfrekvens än dem applikations-specifika processorerna. Resultatet visar också att schemaläggning av instruktionerna i programmen spelar en stor roll för att kunna utnyttja resurserna som finns tillgängliga och därmed öka prestandan. Med den schemaläggningen som utnyttjade resurserna bäst var prestandan minst 79\% bättre än den generella processorn.
}

\end{document}

### 2020-12-03 msccls/MScTemplateCS.tex
Updates from Overleaf

@@
 \usepackage{wrapfig}
+\usepackage{multirow}
+\usepackage{amsmath}
 
@@
 \section{Machine Learning}
-Machine learning is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program it.
+Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program it.
 By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem.
-In \textbf{supervised learning}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number.
+In \textbf{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number.
 The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs information that allows deducing the label y for this feature vector.)
@@
 
-(Källor Andriy Burkov, Chollet)
-
 \subsection{Neural Networks and Deep Learning}
 
-Neural networks are machine learning systems loosely inspired by biological neural networks (Chollet s.8). A neural network is a collection of connected nodes or neurons that
+Neural networks are machine learning systems loosely inspired by biological neural networks.
+A neural network is a collection of connected nodes or neurons that was first mentioned in \citet{mcculloch1943}.
+A neuron \citep{dawson1998ann} is an information-processing unit that receives inputs and produces a single output that can be sent to multiple other neurons.
+To find the output of a neuron the weighted sum of all inputs is weighted by the weights all connecting to the neuron.
+This weighted sum is then passed through an activation function to provide the output.
 
 \begin{quote}
-An ANN is based on a collection of connected units or nodes called artificial neurons, which loosely model the neurons in a biological brain.
-Each connection, like the synapses in a biological brain, can transmit a signal to other neurons.
-An artificial neuron that receives a signal then processes it and can signal neurons connected to it.
+RNN gate    
 \end{quote}
 
-Deep learning = multiple layers of neural networks
+\begin{quote}
+    The learning of the error is the cost function of the network and feeding back is called backpropagation.
+
+\end{quote}
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=12cm]{msccls/explanatory_images/single_neuron.png}
+    \caption{An example of a single neuron}
+    \label{fig:neuron}
+\end{figure}
+
+\begin{equation}
+    g(x_1, x_2, x_3,...,x_n) = g(x) = \sum_{i=1}^n  xi
+\end{equation}
+
+Here, $g(x)$ takes the inputs $x_1, x_2, x_3,...,x_n$, while the activation function $f(x)$ defines the output.
+An example of a well known activation function is ReLU, Rectified linear unit, which has the following function.
+
+\begin{equation}
+y = f(g(x)) =
+\begin{cases}
+  0, & \text{if}\ x \leq 0 \\
+  x, & \text{if}\ x > 0
+\end{cases}
+\end{equation}
+
+\begin{itemize}
+    \item Neurons
+    \item Weights and connections 
+    \item Backpropagation function
+    \item gradient
+    \item Deep learning = multiple layers of neural networks
+\end{itemize}
+
+\subsection{Gated Recurrent Unit - GRU}
+A gated recurrent unit (GRU) is neural network architecture introduced in 2014 by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural network.
+Although fairly new it is already a well-established variant of recurrent neural networks.
+A GRU is quite useful for solving the vanishing gradient problem \citep{hochreiter1998} that might occur in recurrent neural networks.
+The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.
+It has also been shown that GRUs have had better performance on smaller and less frequent datasets \citep{Gruber2020AreGC}.
+GRUs is primarily constructed of an update gate, a reset gate and a current memory gate.
+The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to being disregarded henceforth.
+
+\begin{quote}
+Snacka mer om current memory gate eller ta bort helt
+\end{quote}
+
+\subsection{Long Short-Term Memory - LSTM}
+
+\begin{quote}
+    Ta bara med ifall vi bestämmer oss att ändra från GRU till LSTM i modellen
+    In comparison to GRUs, LSTMs are
+
+\end{quote}
 
@@
 This means that synonyms and words often used together are embedded closer to each other, while words seldom used together are more distant in the embedding.
-In addition to distance, direction could also be a useful tool to extract meaning from the embedding.\newpage
+In addition to distance, direction could also be a useful tool to extract meaning from the embedding.\newpage %Temp for formatting, don't forget it's here
 
 \begin{wrapfigure}{r}{0.25\textwidth}
-\includegraphics[width=0.9\linewidth]{msccls/embedding_direction.png} 
+\includegraphics[width=0.9\linewidth]{msccls/explanatory_images/embedding_direction.png} 
 \caption{Small example of a word embedding space}
@@
 \subsection{F1-score, precision, accuracy}
-\chapter{Approach}
 
@@
 train, dev(validation), test, man testar på dev under development, först på test efter man är klar
+
+\chapter{Implementation}
+
 \section{Datasets}
+
 \subsection{Approach} 
@@
 \section{Results}
+
+These are our results
+
+% Please add the following required packages to your document preamble:
+% \usepackage{multirow}
+\begin{center}
+\begin{tabular}{|c|cc|cc|cc|cc|}
+\hline
+\multirow{2}{*}{Methods} & \multicolumn{2}{c|}{MELD 34}            & \multicolumn{2}{c|}{DailyDialog 45}     & \multicolumn{2}{c|}{IEMOCAP 34}         & \multicolumn{2}{c|}{EmoryNLP 23}        \\ \cline{2-9}
+                         & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss \\ \hline
+Cosmic                   & 64.14                         & 2.68 & 60.01                         & 2.94 & 58.77                         & 0.96 & 42.55                         & 2.48 \\ \hline
+Vår modell               &                               &      &                               &      &                               &      &                               &      \\ \hline
+\end{tabular}
+\end{center}
+
 \section{Discussion}
+This is the discussion about our results
 

### 2020-12-23 msccls/MScTemplateCS.tex
Updates from Overleaf

@@
 It has also been shown that GRUs have had better performance on smaller and less frequent datasets \citep{Gruber2020AreGC}.
-GRUs is primarily constructed of an update gate, a reset gate and a current memory gate.
-The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to being disregarded henceforth.
+Basic GRUs are primarily constructed of an update gate and a reset gate.
+The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to be disregarded henceforth.
 
@@
 \subsection{Long Short-Term Memory - LSTM}
+A basic LSTM-unit has input gate, output gate and forget gate. 
+1997 Hochreiter
 

### 2020-12-29 msccls/MScTemplateCS.tex
Updates from Overleaf

@@
 \section{Terminology}
-NLP, ML, Speech-to-text och (STT), API. Modality. Kan också kallas för Glossary and Abbreviations
+NLP, ML, Speech-to-text och (STT), API. Modality. RNN, GRU, LSTM, BPTT(?)
+
+Kan också kallas för Glossary and Abbreviations
 \section{Delimitations}
@@
 \chapter{Theory}
+
+blablabla information om de olika fields of study
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[scale=0.4]{msccls/explanatory_images/map_of_techniques.png}
+    \caption{Relations between the related fields of study }
+    \label{fig:map}
+\end{figure}
+
 \section{Machine Learning}
@@
 \begin{itemize}
-    \item Neurons
-    \item Weights and connections 
-    \item Backpropagation function
-    \item gradient
+    \item A central algorithm in deep learning is Back-propagation algorithm/function.
+The fundamental trick in deep learning is to use this score as a feedback signal to adjust the value of the weights a little, in a direction that will lower the loss score for the current example.
+This is the job of the optimizer.
+    \item gradient (räcker med vanishing gradient problem?)
     \item Deep learning = multiple layers of neural networks
@@
 \subsection{Gated Recurrent Unit - GRU}
-A gated recurrent unit (GRU) is neural network architecture introduced in 2014 by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural network.
+A gated recurrent unit (GRU) is neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks.
+Basic GRU nodes are primarily constructed of an \textit{update gate} and a \textit{reset gate}.
+The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to be disregarded henceforth.
 Although fairly new it is already a well-established variant of recurrent neural networks.
-A GRU is quite useful for solving the vanishing gradient problem \citep{hochreiter1998} that might occur in recurrent neural networks.
-The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.
-It has also been shown that GRUs have had better performance on smaller and less frequent datasets \citep{Gruber2020AreGC}.
-Basic GRUs are primarily constructed of an update gate and a reset gate.
-The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to be disregarded henceforth.
+In comparison to LSTM layers, GRU layers are cheaper to run since they use less parameters, but have less representational power.
+It has been shown that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers \citep{Gruber2020AreGC}.
 
-\begin{quote}
-Snacka mer om current memory gate eller ta bort helt
-\end{quote}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[scale=0.4]{msccls/explanatory_images/gru.png}
+    \caption{A GRU node}
+    \label{fig:gru_node}
+\end{figure}
 
 \subsection{Long Short-Term Memory - LSTM}
-A basic LSTM-unit has input gate, output gate and forget gate. 
-1997 Hochreiter
+Another widely used neural network architecture is the Long Short-Term Memory (LSTM).
+It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem \citep{hochreiter1998}, that might occur in recurrent neural networks, and has barely been changed since.
+The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.
 
-\begin{quote}
-    Ta bara med ifall vi bestämmer oss att ändra från GRU till LSTM i modellen
-    In comparison to GRUs, LSTMs are
+As illustrated in \ref{fig:lstm_node} a LSTM node consists of an \textit{input gate}, an \textit{output gate}, a \textit{forget gate} and a \textit{cell state}. 
+\begin{itemize}
+    \item The input state controls the extent of data to write onto the internal cell-state.
+    \item The output state controls the extent of data to pass as output. 
+    \item The forget gate, similarly to the GRUs reset gate, decides how much of the previous data is irrelevant and will be ignored. 
+    \item The cell state stores the previous state of a node and passes it on to its following node. 
+\end{itemize}
 
-\end{quote}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[scale=0.4]{msccls/explanatory_images/lstm.png}
+    \caption{A LSTM node}
+    \label{fig:lstm_node}
+\end{figure}
 
@@
 This means that synonyms and words often used together are embedded closer to each other, while words seldom used together are more distant in the embedding.
-In addition to distance, direction could also be a useful tool to extract meaning from the embedding.\newpage %Temp for formatting, don't forget it's here
+In addition to distance, direction could also be a useful tool to extract meaning from the embedding.
+%Temp for formatting, don't forget it's here
 
@@
 \section{Evaluation}
-\subsection{F1-score, precision, accuracy}
+There are many different metrics that can be used to evaluate a machine learning model, based on what kind of model and what its purpose is.
+For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}.
+In a confusion matrix, one of the axis represent the actual class of the instance is trying to predict and the other represents what instance the model predicted.
+The most basic and common confusion matrix is a $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}.
+In this, we have four different squares:
+\begin{itemize}
+    \item TP means that the model was correct in predicting true
+    \item TN means that the model was correct in predicting false
+    \item FP means that the model was incorrect in predicting true
+    \item FN means that the model was incorrect in predicting false 
+\end{itemize}
+
+For cases where classification happens in a higher dimensionality than binary, the same logic is used, just on a bigger scale.
+
+%kan snacka om Type I och Type II errors om mer text behövs.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}
+    \caption{Binary Confusion Matrix}
+    \label{fig:confusion}
+\end{figure}
+
+\subsection{Accuracy}
+One of the simplest metrics used, accuracy quantifies how many percent of predictions that the model got right and is calculated by taking all instances where our model predicted true and dividing it with all instances.
+
+$$\frac{TP + TN}{TP + TN + FP + FN}$$
+
+If we have a balanced data set with a similar amount of true and false instances, this is very valuable information.
+If, however, our set is inbalanced and contains e.g.
+90 percent true instances and only 10 percent false, the model can achieve a 90 percent accuracy by simply predicting true on every instance so the accuracy suddenly becomes useless.
+
+\subsection{Precision}
+Precision is a metric that quantifies
+
+\subsection{Recall}
+
+\subsection{F1-score}
+
+\subsection{Loss}
+
+\section{Overfitting}
 
@@
 \section{Datasets}
-
+Filip
 \subsection{Approach} 

### 2020-12-29 msccls/MScTemplateCS.tex
2500 ord

@@
 blablabla information om de olika fields of study
+Deep learning is mostly successive neural networks but there are other types as well. Deep belief networks.
 
@@
 
-\begin{quote}
-RNN gate    
-\end{quote}
-
-\begin{quote}
-    The learning of the error is the cost function of the network and feeding back is called backpropagation.
-
-\end{quote}
-
 \begin{figure}[htp]
@@
 Here, $g(x)$ takes the inputs $x_1, x_2, x_3,...,x_n$, while the activation function $f(x)$ defines the output.
-An example of a well known activation function is ReLU, Rectified linear unit, which has the following function.
+An example of a well known activation function is \textit{ReLU}, Rectified linear unit, which has the following function.
 
@@
 
-\begin{itemize}
-    \item A central algorithm in deep learning is Back-propagation algorithm/function.
-The fundamental trick in deep learning is to use this score as a feedback signal to adjust the value of the weights a little, in a direction that will lower the loss score for the current example.
-This is the job of the optimizer.
-    \item gradient (räcker med vanishing gradient problem?)
-    \item Deep learning = multiple layers of neural networks
-\end{itemize}
+Neural networks can in turn be structured together and form multiple layers that progressively learn.
+These networks fall under the deep learning umbrella.
+Deep learning has several closely relevant definitions, where \citet{deng2014deep} mentions the following:
+
+\begin{quote}{A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification}
+\end{quote}
+
+Each of the layers in a deep learning network manipulate their input data in some way, and the result is stored in the layer's weights.
+The results are monitored with the help of the \textit{loss function}.
+The loss function returns a distance score that gives an indication on how far off the networks previous prediction was from the true values \citep{franoischollet2017learning}.
+The phenomena of reducing the distance score that the loss function produces is called training.
+Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.
 
@@
 \begin{itemize}
-    \item The input state controls the extent of data to write onto the internal cell-state.
-    \item The output state controls the extent of data to pass as output. 
+    \item The input gate controls the extent of data to write onto the internal cell-state.
+    \item The output gate controls the extent of data to pass as output. 
     \item The forget gate, similarly to the GRUs reset gate, decides how much of the previous data is irrelevant and will be ignored. 
@@
 \end{figure}
+
+\section{Natural Language Processing - NLP}
+Känns vettigt att ha som section
 
@@
 For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}.
-In a confusion matrix, one of the axis represent the actual class of the instance is trying to predict and the other represents what instance the model predicted.
-The most basic and common confusion matrix is a $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}.
-In this, we have four different squares:
-\begin{itemize}
-    \item TP means that the model was correct in predicting true
-    \item TN means that the model was correct in predicting false
-    \item FP means that the model was incorrect in predicting true
-    \item FN means that the model was incorrect in predicting false 
-\end{itemize}
-
-For cases where classification happens in a higher dimensionality than binary, the same logic is used, just on a bigger scale.
-
-%kan snacka om Type I och Type II errors om mer text behövs.
+In a confusion matrix, one of the axis represent the actual class of the instance the model is trying to predict and the other represents what instance the model predicted.
+The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}.
+In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}
 
@@
 
+\begin{itemize}
+    \item TP means that the model was correct in predicting positive
+    \item TN means that the model was correct in predicting negative
+    \item FP means that the model was incorrect in predicting positive
+    \item FN means that the model was incorrect in predicting negative
+\end{itemize}
+
+Whenever our model makes a prediction on an instance during training, we increment the corresponding cell in the matrix, letting us see and draw conclusions based on how our model performed. 
+For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative.
+The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class.
+For instance, if determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{n}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.
+
+%kan snacka om Type I och Type II errors om mer text behövs.
+
 \subsection{Accuracy}
-One of the simplest metrics used, accuracy quantifies how many percent of predictions that the model got right and is calculated by taking all instances where our model predicted true and dividing it with all instances.
+One of the simplest metrics used, accuracy quantifies how many percent of predictions that the model got right and is calculated by taking all instances where our model predicted positive and dividing it with all instances.
 
@@
 
-If we have a balanced data set with a similar amount of true and false instances, this is very valuable information.
-If, however, our set is inbalanced and contains e.g.
-90 percent true instances and only 10 percent false, the model can achieve a 90 percent accuracy by simply predicting true on every instance so the accuracy suddenly becomes useless.
+If we have a balanced data set with a similar amount of positive and negative instances, this can give valuable information on the performance of the model since getting predictions right is the goal of any model.
+If, however, our set is imbalanced and contains e.g.
+90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless.
 
 \subsection{Precision}
-Precision is a metric that quantifies
+Precision is a metric that quantifies how many of the instances predicted positive that actually were true.
+It's calculated by taking all of the instances where the model predicted positive and was correct, and dividing it by all the instances where it predicted positive.
+
+$$\frac{TP}{TP + FP}$$
+
+Precision gives us information on how accurate the model is when predicting an instance to be true.
+This is useful when dealing with instances where it's important to not give a false positive.
+If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for a innocent person to be determined a criminal.
+Precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, it's likely that said person is guilty.
+
+When looking at multiple classes, the formula remains the same but with TP and FP representing the sum of all TPs and FPs
 
@@
 
+How many did we get right, of all that were actually right
+
+False negative, such as illness
 \subsection{F1-score}

### 2021-01-14 msccls/MScTemplateCS.tex
Shit ton of changes

@@
 \usepackage{wrapfig}
+\usepackage{amsmath}
 \usepackage{multirow}
-\usepackage{amsmath}
+\usepackage{tabularx}
+\usepackage{subfig}
+\usepackage{textcomp}
+\usepackage{calligra}
+
+\newcommand{\icon}[1]{\includegraphics[height=30pt]{msccls/explanatory_images/CASTLE_logo.png}}
 
@@
 %\student{Flavius Gruian}{Flavius.Gruian@cs.lth.se}
-\students{Nick Persson}{dat14npe@student.lu.se}{Filip Karabeleski}{dat14fka.student.lu.se}
+\students{Nick Persson}{nick@persson.dev}{Filip Karabeleski}{fkarabeleski@gmail.com}
 
-\thesisnumber{LU-CS-EX: 2020-XX} % Birger Swahn will provide this number to you, once the thesis is ready for publication
+\thesisnumber{LU-CS-EX: 2021-XX} % Birger Swahn will provide this number to you, once the thesis is ready for publication
 % default is Master. Uncomment the following for "kandidatarbete"/Bachelor's thesis
@@
 %\title{Formatting a Master's Thesis}
-\title{Keyword Categorization and Emotion Analysis in Voice Calls}
+\title{Emotion Recognition in Written Conversations Using CASTLE}
 
@@
 
-\subtitle{Subtitle WIP}
+\subtitle{\protect\icon{CASTLE}}
 \company{Telavox AB}
@@
 \examiner
-{Flavius Gruian,\href{mailto:Flavius.Gruian@cs.lth.se}{\texttt{Flavius.Gruian@cs.lth.se}}}
+{Flavius Gruian, \href{mailto:Flavius.Gruian@cs.lth.se}{\texttt{Flavius.Gruian@cs.lth.se}}}
 
@@
 \theabstract{
-This document describes the Master's Thesis format for the theses carried out at 
-the Department of Computer Science, Lund University.
+%This document describes the Master's Thesis format for the theses carried out at 
+%the Department of Computer Science, Lund University.
 
-Your abstract should capture, in English, the whole thesis with focus on the problem and solution in 150 words.
+%Your abstract should capture, in English, the whole thesis with focus on the problem and solution in 150 words.
 It should be placed on a separate right-hand page, with an additional \textit{1cm} margin on both left and right.
@@
 
-Leave a \textit{2cm} vertical space after the abstract and provide a few keywords relevant for your report. Use five to six words, of which at most two should be from the title.
+%Leave a \textit{2cm} vertical space after the abstract and provide a few keywords relevant for your report. Use five to six words, of which at most two should be from the title.
 }
 
-\keywords{Machine Learning, NLP, Sentiment Analysis, Keyword Extraction, (style, structure)}
+\keywords{Machine Learning, NLP, Emotion Recognition}
 
@@
 
+%ML är automagiskt
+
 \section{Background}
 Natural language processing (NLP) is a field of study mainly in computer science that deals with analyzing human language in a way that computers can understand and process.
-According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notations.
-Since natural languages continuously develop as time goes on and its adherence to structures diminish, it is difficult to define explicit rules to follow.
+According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notation.
+Since natural languages naturally evolve as time goes on and their adherence to structures diminish, it is difficult to define explicit rules to follow.
 
-\subsection{Emotion Recognition}
-Emotion has always been a key part of human language, and it therefore comes naturally that teaching a computer to understand emotion in human language is a big part of NLP.
-Initially, the focus lied in sentiment analysis, a field of study focused on determining the binary polarization of piece of text as either positive or negative.
-Godtycklig bit om sentiment analysis.
-Emotion recognition is an extension of sentiment analysis that instead of just determining the polarization of text, attributes emotion to it.
-This process is more challenging than sentiment analysis, as the differences between emotions are subtler than that of just positive and negative.
-Emotion recognition has also been extended to work on different modalities, such as for example visual and auditory input.
+%prata om emotion recognition
 
-Emotion recognition can be applied to data in various modalities, 
-such as visual, in the form of facial expressions and body language, audible and text based in the form of dialogues. To o
+%The LSTM architecture \citep{ntlk2009} is something....
 
-The LSTM architecture \citep{ntlk2009} is something....
+%\begin{itemize}
+%    \item Förklara vad systemet är
+%    \item Berätta om företaget Telavox
+%\end{itemize}
+
+\section{Purpose}
+Customer service is centered around the relation between company and customer with the goal of pleasing the customer.
+This goal can be very difficult to achieve and there is no straightforward way to do so.
+By analyzing prior conversations it is possible for companies to gain valuable insights in the most effective way to please the customer.
+Customer service is however a massive industry and it would take countless hours to analyze the conversations from even a single company manually.
+The company sponsoring this project approached us with an idea to try and solve this problem using machine learning.
+A usual technique for these types of problems is \textit{sentiment analysis} where utterances are labeled either as positive or negative.
+We instead decided to take another route, namely \textit{emotion recognition}.
+Emotion recognition focuses on labelling utterances from a non-binary set of emotions instead of sentiment analysis' binary method.
+This allows an analysis with greater specificity.
+In order to further the academic relevance of this project we decided that we would attempt to contribute to the field of emotion recognition by achieving state-of-the-art results.
+
+\section{Telavox}
+Telavox AB is a Swedish telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service.
+As all of their customers deal in customer service they would all benefit from the solution previously discussed.
+Telavox therefore has a great interest in being able to offer this as a service.
+As Telavox has over 250 businesses as customers, the solution would facilitate the quality of customer service world wide.
+
+\section{Research questions}
+The primary aim of this thesis is to create a emotion recognition model that is able to produce state-of-the-art results.
+After this, a secondary aim is to investigate whether or not it is possible to create a dynamic model for instant response regarding customer service related conversations.
+The following questions are used as guidelines on how to realize these aims.
 
 \begin{itemize}
-    \item Förklara vad systemet är
-    \item Berätta om företaget Telavox
+    \item \textit{How does our model compare to other models in the same field evaluating on the same metrics?}
+    \item \textit{Is it possible for an LSTM architecture to produce better results than a GRU one for emotion recognition?}
+    \item \textit{After implementing a fully functioning model, how can we adapt it to predict utterances in real-time?}
 \end{itemize}
 
-\section{Purpose}
-\begin{itemize}
-    \item Varför behöver problemet lösas, nytta
-\end{itemize}
-\subsection{Telavox}
-Telavox is a telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service. 
-\section{Research questions}
-The primary aim of this thesis is to investigate whether or not voice call analysis using NLP can be effective enough to be used as a replacement for manual analysis. \\
-\textit{What level of precision can be reached with the categorization of keywords from voice calls?}\\
-\textit{How well will our algorithm perform compared to other algorithms carrying out the same measurements?}\\
-\textit{What level of effectiveness can be reached compared to a human manually analyzing the data?}
+%NLP, ML, Speech-to-text och (STT), API. Modality. RNN, GRU, LSTM, BPTT(?), utterance, speaker, dialogue,
 
-\section{Terminology}
-NLP, ML, Speech-to-text och (STT), API. Modality. RNN, GRU, LSTM, BPTT(?)
+%Hyperparameters en punkt i listan, items i hyperparameters en sublist
 
-Kan också kallas för Glossary and Abbreviations
-\section{Delimitations}
-Delimitations are choices made by the researcher which should be mentioned. They describe the boundaries that you have set for the study.
+%Kan också kallas för Glossary and Abbreviations
+
 \section{Contributions}
-Potentialen av applications 
+%Potentialen av applications 
 \section{Related work}
@@
 
-blablabla information om de olika fields of study
-Deep learning is mostly successive neural networks but there are other types as well. Deep belief networks.
+The celebrity of computer science at this time is undoubtedly artificial intelligence (AI).
+A much more complex subject than most people know.
+Most of the complex problems in AI are solved using technologies within its sub-fields, which are illustrated in Figure \ref{fig:map}, rather than traditional artificial intelligence such as robotics \citep{bajwa_2020}.
+In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model.
+Firstly, the model should be created with deep learning algorithms.
+This enables it to find patterns that it otherwise could not be able to.
+Secondly, to effectively process the utterances it has been tasked to analyze, a word vector is often used to store the most commonly used words for easier analysis.
 
@@
 By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem.
-In \textbf{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number.
+In \textit{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number.
 The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs information that allows deducing the label y for this feature vector.)
-The other main type of tasks in machine learning is \textbf{unsupervised learning}.
+The other main type of task in machine learning is \textit{unsupervised learning}.
 The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector.
@@
 
+%Weights control the signal (or the strength of the connection) between two neurons. In other words, a weight decides how much influence the input will have on the output.
+
 \begin{figure}[htp]
@@
     \includegraphics[width=12cm]{msccls/explanatory_images/single_neuron.png}
-    \caption{An example of a single neuron}
+    \caption{An example of a single neuron.}
     \label{fig:neuron}
@@
 
-Each of the layers in a deep learning network manipulate their input data in some way, and the result is stored in the layer's weights.
+Each of the layers in a deep learning network manipulates their input data in some way, and the result is stored in the layer's weights.
 The results are monitored with the help of the \textit{loss function}.
-The loss function returns a distance score that gives an indication on how far off the networks previous prediction was from the true values \citep{franoischollet2017learning}.
-The phenomena of reducing the distance score that the loss function produces is called training.
-Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.
+The loss function returns a distance score that indicates how far off the network's previous prediction was from the true values \citep{franoischollet2017learning}.
+The process of reducing the distance score that the loss function produces is called training.
+Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi-variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.
 
-\subsection{Gated Recurrent Unit - GRU}
-A gated recurrent unit (GRU) is neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks.
+\subsection{Recurrent Neural Network -- RNN}
+
+One of the simpler neural network classes used in deep learning is recurrent neural networks (RNN) and they process either sequential data or time-series data \citep{lipton2015critical}.
+Since the inputs are sequential, all the inputs are dependant on each other in one way or another.
+RNNs are therefore suitable for processing sequences of text, audio, and video where every computation will be performed on each element in the sequence.
+
+One of RNN's weaknesses is that they suffer the \textit{vanishing gradient problem} \citep{hochreiter1998}.
+Theoretically, RNNs can process arbitrarily long sequences, in practice, they are limited to the length of the sequence.
+If the sequence is too long, the network will struggle to try to look back in time longer than a few steps away.
+In a lot of cases, this will result in leaving out important information from the prediction.
+There are instead other alternatives that consider this weakness.
+
+%Time series.
+Meant to process sequences of data like text, audio, video.
+Neural networks that feed their output back into their inputs recursively.
+One of the more basic building blocks in machine learning/deep learning.
+A gate looks like this:
+
+%suffer from short-term memory.
+If a sequence is long enough, they’ll have a hard time carrying information from earlier time steps to later ones.
+So if you are trying to process a paragraph of text to do predictions, RNN’s may leave out important information from the beginning.
+
+ \begin{figure}[!ht]
+    \centering
+    \includegraphics[scale=0.3]{msccls/explanatory_images/rnn.png}
+    \caption{A simple RNN node}
+    \label{fig:rnn_node}
+\end{figure}
+
+\subsection{Gated Recurrent Unit -- GRU}
+%Byta alla "node" till "cell" för att lättare kunna referera till LSTM-celler i CASTLE
+A gated recurrent unit (GRU) is recurrent neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks.
 Basic GRU nodes  are primarily constructed of an \textit{update gate} and a \textit{reset gate}.
@@
 Although fairly new it is already a well-established variant of recurrent neural networks.
-In comparison to LSTM layers, GRU layers are cheaper to run since they use less parameters, but have less representational power.
+In comparison to LSTM layers, GRU layers are cheaper to run since they use fewer parameters, but have less representational power.
 It has been shown that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers \citep{Gruber2020AreGC}.
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
     \centering
-    \includegraphics[scale=0.4]{msccls/explanatory_images/gru.png}
+    \includegraphics[scale=0.5]{msccls/explanatory_images/gru.png}
     \caption{A GRU node}
@@
 
-\subsection{Long Short-Term Memory - LSTM}
-Another widely used neural network architecture is the Long Short-Term Memory (LSTM).
-It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem \citep{hochreiter1998}, that might occur in recurrent neural networks, and has barely been changed since.
+\subsection{Long Short-Term Memory -- LSTM}
+Another widely used recurrent neural network architecture is Long Short-Term Memory (LSTM).
+It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem that might occur in recurrent neural networks and has barely been changed since.
 The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.
@@
 
-\begin{figure}[h!]
+\begin{figure}[!ht]
     \centering
-    \includegraphics[scale=0.4]{msccls/explanatory_images/lstm.png}
+    \includegraphics[scale=0.45]{msccls/explanatory_images/lstm.png}
     \caption{A LSTM node}
@@
 
-\section{Natural Language Processing - NLP}
-Känns vettigt att ha som section
+\section{Training Process}
+In order to train a model there must exist some data set, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method, that feeds the model with input.
+When using a labeled data set the labels are regarded as the solution to the problem.
+When training a model the goal is to find values for its weights such that the model correctly labels the given input.
+The training of a model is done step by step in the shape of \textit{epochs}.
+Epochs represent the number of iterations that the model has passed through a data set intended for training the model.
+Following each epoch, a \textit{loss function} measures how bad the model's previous prediction was and produces a value for the \textit{loss}.
+The model will have a perfect prediction when the loss reaches zero, therefore minimizing the loss is desirable.
 
-\section{Word Vectors}
-Word vectors are a type of word representation where texts are turned into numbers
+In order to minimize the loss you need to know what the best values for the weights to do so are. To do this, while training the model iteratively calculates the derivatives of the
+
+We want to travel through the function and find the place where the derivative is zero.
+This is done for each weight in the network.
+It is accomplished by iteratively using backpropagation algorithm that finds a value closer the the derivative being zero.
+Gradient descent is then used to update the weights to match the result of the backpropagation algorithm.
+
+direction the weights are taking the model. This is done by evaluating the derivatives of the ...in the network and that is exactly what the \textit{gradient descent} does.
+
+%In gradient descent one is trying to reach the minimum of the loss function with respect to the parameters using the derivatives calculated in the back-propagation.
+
+%gradient descent iteratively tries to find a minimum. tries to regulate the weights in order to minimize loss.
+
+%Datasets ger inputs till model, första lagret ..... Modellen har svaren till hands och använder dessa för att styra skeppet i rätt riktning. 
+%There exist many components to training a model .....
+%In \ref{chapter}, backpropagation and gradient descnet among other temrs were mentioned but not extensively explored.
+The training process can be described as the road from a standalone model to having relevant results.
+
+% In this context, learning means finding a set of values for the weights of all layers in a network, such that the network will correctly map example inputs to their associated targets
+
+%Backpropagation
+%Gradient Descent
+%Loss (function)
+%Från början till slut, från en modell med lager till en evaluation, sen fortsätter vi snacka om det under evaluation
+%Overfitting
+%Epochs
+
+\section{Evaluation}
+Many different metrics can be used to evaluate a machine learning model, based on what kind of model and what its purpose is.
+For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}.
+In a confusion matrix, one of the axes represents the actual class of the instance the model is trying to predict and the other represents what instance the model predicted.
+The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}.
+In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}
+
+\begin{figure}[h!]
+    \centering
+    \hbox{\hspace{6em}\includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}}
+    \caption{Binary Confusion Matrix}
+    \label{fig:confusion}
+\end{figure}
+
+\begin{itemize}
+    \item TP means that the model was correct in predicting positive
+    \item TN means that the model was correct in predicting negative
+    \item FP means that the model was incorrect in predicting positive
+    \item FN means that the model was incorrect in predicting negative
+\end{itemize}
+
+Whenever our model predicts an instance during training, we increment the corresponding cell in the matrix, letting us see and draw conclusions based on how our model performed.
+
+For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative.
+The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class.
+For instance, when determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{neutral}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.
+
+%kan snacka om Type I och Type II errors om mer text behövs.
+
+\subsection{Accuracy}
+One of the simplest metrics used, accuracy quantifies how many percent of predictions that the model got right and is calculated by taking all instances where our model predicted positive and dividing it with all instances.
+
+$$ Accuracy = \frac{TP + TN}{TP + TN + FP + FN}$$
+
+If we have a balanced data set with a similar amount of positive and negative instances, this can give valuable information on the performance of the model since getting predictions right is the goal of any model.
+If, however, our set is imbalanced and contains e.g.
+90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless \citep{imbalancedlearning}.
+
+\subsection{Precision}
+Precision is a metric that quantifies how many of the instances predicted positive that were true.
+It's calculated by taking all of the instances where the model predicted positive and was correct and dividing it by all the instances where it predicted positive.
+
+$$ Precision = \frac{TP}{TP + FP}$$
+
+Precision gives us information on how accurate the model is when predicting an instance to be true.
+This is useful when dealing with instances where it's important to not give a false positive.
+If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for an innocent person to be determined as a criminal.
+However, precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, that said person is likely guilty.
+
+\subsection{Recall}
+Recall is a metric that quantifies how many percent of positive instances that our model managed to predict which we calculate by dividing all instances the model predicted positive and was right by all positive instances.
+
+$$ Recall =\frac{TP}{TP + FN}$$
+
+As a sort of counterpoint to precision which tells us how good the model is when guessing true, recall tells us how many positive instances the model missed.
+A high recall can be extremely important when dealing with subjects such as contagious diseases where a result of false-negative can be devastating.
+Recall doesn't take into account how many positive predictions turned out to be false so in cases where that doesn't matter and you want to predict positive at the slightest probability of it being true, optimizing for a high recall is essential.
+
+\subsection{F1-score}
+A combination of the two previous metrics, F1-score is calculated by creating the harmonic mean of \textit{precision} and \textit{recall}.
+
+$$ F1score = 2 \cdot \frac{Precision \cdot Recall}{Precision + Recall} $$
+
+In cases where there is no extra importance in finding either false negatives or false positives, you just want to minimize the risk of errors whatsoever.
+The harmonic mean is a kind of average that is appropriate when working with ratios, which is what both precision and recall are.
+This lets us improve the model concerning either of the variables and not have to think about if, for example, the reduction in Recall was worth the gain in Precision.
+In multi-class scenarios with imbalanced data sets there are multiple ways to calculate the F1-score, namely \textit{macro}, \textit{weighted} and \textit{micro}.
+Which one to use depends on your data set and what you prioritize in your model.
+
+The macro F1-score is derived by calculating the F1-score for each class separately and then just summing them up, leading to disproportionate weight on the classes with fewer instances.
+If we assume that we are using a data set where the \textit{neutral} class is represented in 10 000 instances and a less common emotion \textit{fear} in just 50.
+The model then managed to predict 9 995 of the neutral instances, but only 25 of the fear ones.
+When calculating the mean F1-score, the low fear score would muddy the entire score and make it seem bad, when in fact the score for the neutral case is excellent.
+If we want to optimize for the minority classes, however, this lets us do just that.
+
+To counteract this behavior of favoring the minority class, one could instead calculate the weighted F1-score.
+After calculating the F1-score for each class, they are weighted by the number of true instances for each class, thus making them proportionate before calculating the final score.
+This gives a score more reflecting on how the model performs, but can also favor identifying the majority class heavily, which might not be the goal of the model.
+
+When calculating micro F1-score, instead of calculating the f1-scores separately, the f1-score is calculated directly by using the total values for TP, FN, and FP in the Precision and Recall formulas which places no bias on either class.
+If we have an example data set with two minority classes \textit{anger} and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class.
+From the perspective of the joy class, however, it would instead be an FP.
+Since micro calculations take into account the total values for our identification labels and each FN correspond to the FP of another class, $FP = FN$, which in turn means that $Precision = Recall$ in these cases.
+Precision and Recall being equal means that their mean also is equal and thus \textit{F1-}$score_{micro} = Precision_{Micro} = Recall_{Micro}$ which incidentally also equals \textit{Accuracy} as this is the proportion of instances we got correct.
+This means that just like Accuracy this is a good metric on balanced data sets but can be deceiving in an imbalanced one.
+
+\section{Loss}
+%might be removed
+
+\section{Overfitting}
+Overfitting is a concept in statistical analysis in which the result corresponds too closely to a certain data set.
+Instead of the model learning how to notice and predict features, it simply remembers a massive amount of examples from the training set.
+This causes the model to perform very well on that specific data but would likely have a very high error rate on any new data.
+
+\section{Natural Language Processing -- NLP}
+Natural languages are the set of languages that humans have evolved naturally without any predetermined strategy like e.g English or Mandarin.
+Making computers understand and analyze natural languages has long been a challenge and one of the first practical attempts was Weaver and Booth's translation project in 1946 trying to crack enemy codes during World War II \citep{Liddy2001NLP}.
+Shortly thereafter \citet{Turing1950} created his famous \textit{Imitation game} or more commonly referred to as the Turing test, which has developed into a test to evaluate whether a computer's natural language response in a real-time written conversation can be indistinguishable from that of a human.
+One of the reasons that make natural language processing a challenge is the fact that natural languages are minted with unstructured types of data such as inflection or dialects, or using phrases to convey the opposite than the common meaning namely sarcasm.
+Natural languages are also oftentimes ambiguous and require a context-based interpretation that computers have a hard time understanding.
+
+%is used for classification, translation, emotion/sentiment analysis,
+
+Emotion is a key part of human language and it, therefore, comes naturally that teaching a computer to understand emotion in human language is a big part of NLP.
+Initially, the focus lied in sentiment analysis, a field of study focused on determining the binary polarization of a piece of text as either positive or negative.
+%Godtycklig bit om sentiment analysis. 
+Emotion recognition is an extension of sentiment analysis that instead of just determining the polarization of text, attributes emotion to it.
+This process is more challenging than sentiment analysis, as the differences between emotions are subtler than that of just positive and negative.
+Emotion recognition has also been extended to work on different modalities, such as visual and auditory input.
+
+%Emotion recognition can be applied to data in various modalities, 
+%such as visual, in the form of facial expressions and body language, audible and text-based in the form of dialogues.
+
+\subsection{Word Vectors}
+%Om vi behöver en tie in: Instead of analysing the words themselves, they are converted into numerical representations that are easier for the computer to handle.
+
+Word vectors are a type of word representation where texts are turned into a series of numbers, each word or sometimes phrase instead represented by a number.
+
 \subsection{One-Hot Encoding}
@@
 Each word is then represented by an array filled with zeroes except for one index set to one.
-For example, the one-hot encoding of the phrase \textit{Why was six afraid of seven} would start with the assignment of values to word:
+For example, the one-hot encoding of the phrase \textit{Why was six afraid of seven} would start with the following assignment of values to words:
 
@@
 The other big issue with it is that no meaning can be extracted from the matrix.
-If our vocabulary contained the words "cat", "dog" and "hamster", we could easily see that these are all domestic animals but in the space of the one hot vector, "cat" is as close to "dog" as it would be to "airplane".
+If our vocabulary contained the words "cat", "dog" and "hamster", we could easily see that these are all domestic animals but in the space of the one-hot vector, "cat" is as close to "dog" as it would be to "airplane".
 
@@
 Here, instead of creating a sparse, hardcoded vector with equal distance between words, they are learned from data and packed into a lower-dimensional vector where each dimension contains some relevant data about the word.
-According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional or 1024-dimensional are common when dealing with very large vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use one-hot encoding.
+According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional, or 1024-dimensional are common when dealing with very large vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use a one-hot encoding.
 Additionally, the geometric relationship between words can and should reflect the semantic relationship between those same words.
 This means that synonyms and words often used together are embedded closer to each other, while words seldom used together are more distant in the embedding.
-In addition to distance, direction could also be a useful tool to extract meaning from the embedding.
+In addition to distance, the direction could also be a useful tool to extract meaning from the embedding.
 %Temp for formatting, don't forget it's here
@@
 \includegraphics[width=0.9\linewidth]{msccls/explanatory_images/embedding_direction.png} 
-\caption{Small example of a word embedding space}
+\caption{Small example of a word embedding space. From \citet{franoischollet2017learning}.}
 \label{fig:direction}
@@
 
-\section{Evaluation}
-There are many different metrics that can be used to evaluate a machine learning model, based on what kind of model and what its purpose is.
-For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}.
-In a confusion matrix, one of the axis represent the actual class of the instance the model is trying to predict and the other represents what instance the model predicted.
-The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}.
-In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}
+%Pytorch
+%GRU
+%Context-based 
+%Commonsense Features
+%Förklara bilden
 
-\begin{figure}[h!]
+%\section{Theory}
+%NLP technique, categorization, ML, supervised/unsupervised,  
+%Mechanical turks to 
+%Finns både text och audio på MELD, \citep{zhang2019modeling} visar att text för sig är bättre än ljud men att text+ljud > text
+%train, dev(validation), test, man testar på dev under development, först på test efter man är klar
+
+\chapter{Implementation}
+%Summary of what the chapter is about
+The following chapter is mainly focused on giving a clear description of what we have done for the duration of the project.
+Firstly we will describe our model \textit{CASTLE}, how it has been implemented and what changes were made from the baseline \textit{COSMIC}.
+Our initial goal was to have our model produce state-of-the-art results in emotion recognition which later on changed to a more primary focus on optimizing the run time instead.
+We based our research strategy on the work by \citet{bouthillier:hal-02447823} that summarizes and reviews the experimental procedures that are most commonly used in the field of machine learning.
+
+\section{Baseline -- COSMIC}
+When trying to improve existing results in many machine learning fields, instead of starting from scratch, a baseline is used.
+A baseline is an already functioning model that provides basic results to the problem at hand.
+This provides the opportunity to allocate resources on improving the model instead of building it from the ground up and allows the machine learning community to expand on and draw benefits from each other's work.
+
+This project has been developed using the COSMIC model \citep{ghosal2020cosmic} as a baseline.
+COSMIC is a GRU-based model implemented with PyTorch that has achieved state-of-the-art results across four data sets commonly used in emotion recognition.
+The idea behind the model is when classifying an utterance, the emotions of the utterances both preceding and succeeding it are also taken into account.
+This way the context of the conversation can be used to more accurately predict the emotions of individual utterances.
+
+%COMSICs arkitektur bygger på att de skapar egna GRU-celler och GRU-lager som bygger på PyTorchs respektive. 
+%Filip
+
+Figure \ref{fig:COSMIC_arch} provides a simplified description of how COSMIC operates in a conversation between person A and person B.
+For each time step $t$ there is an utterance said by either A or B with the other person as a listener.
+At the next time step $t+1$ the roles are alternated.
+For each utterance, there exist multiple GRU-cells that predict both the intent of the speaker and the response of the listener at a certain time step.
+These cells are used not only to predict the emotion of the current utterance but also used for the prediction of emotions in other utterances related to said utterance.
+
+%Softmax, relu mm
+
+\begin{figure}[!ht]
     \centering
-    \includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}
-    \caption{Binary Confusion Matrix}
-    \label{fig:confusion}
+    \includegraphics[scale=0.3]{msccls/explanatory_images/COSMIC_arch.png}
+    \caption{COSMIC architecture. From \citet{ghosal2020cosmic}.}
+    \label{fig:COSMIC_arch}
 \end{figure}
 
+\section{CASTLE}
+
+COSMIC-based AnalySer Taking advantage of LSTM for Emotion recognition, CASTLE, was developed as the solution to our research questions.
+We have made several improvements on the COSMIC baseline, one of the fundamental being replacing the internal GRU architecture with an LSTM-based architecture as the name of the model indicates.
+At the start of the project we read some recently published research in the field of deep learning, mainly the article by \citet{edseee.922172720200601} that states that GRU is not as suited for small pieces of text and large data sets as LSTM.
+This is where we got the inspiration to test whether or not the architectural change to LSTM could improve the results.
+
+COSMIC takes advantage of the \texttt{cudnn} framework in PyTorch to assist in multiple ways such as the \texttt{backends.cudnn.benchmark} which attempts to automatically optimize the algorithm during runtime.
+As we aimed to obtain as much control as possible over how the model behaved, we excluded this and other third party features that restricted our domain of control.
+COSMIC is also implemented with a \textit{residual} connection \citep{he2015deep} which allows the model to redefine the layers as learning residual functions referencing the layer inputs.
+This is supposed to facilitate the training of the model.
+While we originally wanted to implement this feature in CASTLE, there was simply not space to do it within the scope of the project.
+
+%Didnt implement residual for CASTLE tidsbrist
+
+%Removed CUDA Backend since we wanted as little noise as possible interfering with our model We didnt want the reason for the results to be out of our reach. Third party.
+
+%COSMIC-based AnalySer Taking advantage of LSTM for Emotion recognition
+
+\section{Experimental Setup}
+
+%We chose COSMIC as our baseline because SOTA, more experienced in PyTorch
+
+When deciding what configuration to use for our experiments, we based our decisions on the empirical studies made by  \citet{bouthillier:hal-02447823}.
+In their report they surveyed authors whose papers were presented at the peer-reviewed conferences NeurIPS 2019 and ICLR 2020.
+As the work is based on answers from relevant authors in the machine learning community, we found it appropriate to use the answers as guidelines for our setup.
+Instead of relying on our own intuition, basing it on the broad consensus grants the experiment a certain sense of validity.
+In Figure \ref{fig:HAL} we have presented the questions and responses we took into consideration.
+
+CASTLE is fine-tuned on three hyperparameters where a combination of \textit{grid search} and \textit{random search} was used.
+On average we performed 20 sets of different values for the three hyperparameters across all the data sets.
+CASTLE is was intended to be evaluated on four different corpora, or datasets, but ultimately only provided results for three of them.
+Our results are based upon 10 consecutive runs with the same setup.
+
+Comparing this to the responses observed in \ref{fig:HAL}, we are in line with the majority in all but one instance, (f), where the majority base their results on 5 or less samples and we instead chose to base it on 10 samples as a higher sample size tends to become a more reliable result.
+
+%How many sets of hyperparameters we evaluated? Vi hade i snitt 20:ish per dataset
+%4 dataset for the experiment
+%Reported 10 samples as one averge.
+
+%Stycke där vi jömför med HAL
+
+\begin{figure}[!ht]
+
+\centering
+\subfloat[Did you optimize your hyperparameters?] {\includegraphics[width=\linewidth/2]{msccls/Questionnaire/Q2.png}}\hfil
+\subfloat[If yes to (A), how did you tune them?] {\includegraphics[width=\linewidth/2]{msccls/Questionnaire/Q3.png}}
+
+\subfloat[How many hyperparameters did you optimize?] {\includegraphics[width=\linewidth/2]{msccls/Questionnaire/Q4.png}}\hfil  
+\subfloat[How many trials/experiments in total during the optimization?] {\includegraphics[width=\linewidth/2]{msccls/Questionnaire/Q5.png}}
+
+\subfloat[How many datasets or tasks did you compare on?] {\includegraphics[width=\linewidth/2]{msccls/Questionnaire/Q8.png}}\hfil
+\subfloat[How many results did you report for each model?] {\includegraphics[width=\linewidth/2]{msccls/Questionnaire/Q9.png}}
+\caption{Collection of questionnaire answers regarding machine learning experiments. From \citet{bouthillier:hal-02447823}.}
+\label{fig:HAL}
+\end{figure}
+
+%Hur vi "räknade ut" våra resultat, 10 avg på best valid ->( test)
+%Epochs
+%Google Colab
+%Ray Tune för att studera hyperparametrar och automatisera
+%HAL - diskussion
+
+%Lack of resources
+\section{Optimizing Hyperparameters}
+As previously mentioned, CASTLE is optimized on three different hyperparameters.
+The ones we chose are \textit{dropout}\citep{hyperparameters}, \textit{batch size}, and \textit{learning rate}.
+Dropout is a value between 0 and 1 that regulates what percentage of learned data that the model randomly chooses to forget.
+This is used in order to combat overfitting and make sure the model is actually predicting instead of just learning the structure of the data set.
+Batch size governs how many training examples will be utilized in each iteration.
+A larger value increases the accuracy of the gradient in trade off for a higher memory cost.
+Learning rate defines the speed of which the model updates its parameters.
+A high learning rate speeds up the learning at the risk of not converging while a low value will more likely converge, sacrificing speed in the process.
+
+To optimize our hyperparameters, we implemented the \texttt{Tune} framework offered by \texttt{Ray}.
+This framework allowed us to efficiently experiment with many different configurations and provided a simple way to run trials concurrently.
+Most of our optimization were done with a combination of grid search and random search.
+Random search is a function with two parameters \textit{start} and \textit{end}, and for each trial it randomly selects a float from in between the two values to use for the given hyperparameter.
+Grid search instead takes in a list of floats as a parameter and runs a trial for each value in the list.
+Multiple grid searches can very easily take a large amount of time to run as all possible combination of values across all grid searches will be tested in separate trials.
+
+Our approach was to start out with random search on parameters to get a sense on what range of values produced optimal results.
+Once we had established the magnitude of which we wanted our parameters to be, we tested out various combinations with grid search, keeping the ones that produced promising results and replacing the others with new trials.
+
+%Hyperparametrization with Tune
+%Learning rate
+%Dropout
+%Batch_size [RAM limitations]
+%Definitioner finns redan i terminologi
+\section{Corpora}
+
+To evaluate our model we used four popular corpora that are recurring in emotion recognition research work.
+We chose these data sets as we wanted our results to be comparable with the current state-of-the-art, namely COSMIC as well as the common practice observed in Figure \ref{fig:HAL}(e).
+As COSMIC compared its results to several other models using these data sets, the results of CASTLE will be directly comparable with COSMIC and the other models.
+
+All the following data sets include a \textit{neutral} emotion that is not very elucidatory.
+An emotion is classified as neutral when it can not be classified as any of the other emotions within the corpus using each data sets corresponding methods for classifying the emotions.
+%mention mechanical turks as example? [MELD?]
+
+%Ranges in size
+%These datasets are recurring in many of the papers on this subject. -> our results comparable
+%All these data sets include a "neutral" emotion.
+This is when an emotion couldn't fit into any of the other categories within the data set using their corresponding methods for classifying the emotions.
+%Fyra populära dataset inom NLP
+
+%train-valid-test i vårt fall
+
 \begin{itemize}
-    \item TP means that the model was correct in predicting positive
-    \item TN means that the model was correct in predicting negative
-    \item FP means that the model was incorrect in predicting positive
-    \item FN means that the model was incorrect in predicting negative
+    \item \textit{MELD} \citep{poria2019meld} is based on about 13000 utterances and 1400 dialogues from the TV series \textit{Friends} and is an extension to the EmotionLines dataset \citep{chen2018emotionlines}.
+MELD is multimodal and  frequently has more than two speakers per dialogue.
+The utterances are labeled with the following emotions: \textit{anger}, \textit{disgust}, \textit{sadness}, \textit{joy}, \textit{surprise}, \textit{fear} and \textit{neutral}.
+
+    \item \textit{IEMOCAP} \citep{busso2008iemocap} is a multimodal, multispeaker corpus with utterances from acting and improvisation scenarios emphasizing on emotional expressions.
+The utterances are labeled with the following emotions: \textit{anger}, \textit{happiness}, \textit{excitement}, \textit{sadness}, \textit{frustration}, \textit{fear}, \textit{surprise}, \textit{other} and \textit{neutral state}.
+
+    \item \textit{EmoryNLP} \citep{zahiri2017emoryNLP} is also a corpus with transcripts from the TV series \textit{Friends}.
+In comparison to MELD, EmoryNLP has over 60.000 utterances.
+The utterances are labeled with the following emotions: neutral, joyful, peaceful, powerful, scared, mad and sad.
+
+    \item \textit{DailyDialog} \citep{lietal2017dailydialog} encompasses numerous daily life topics with human-written dialogues to reflect usual day-to-day conversations.
+The corpus has an average of circa 8 speakers per dialogue and has a total of over 13.000 dialogues with their emotions manually labeled.
+The utterances are labeled with the following emotions: anger, disgust, fear, joy, neutral, sadness, and surprise.
 \end{itemize}
 
-Whenever our model makes a prediction on an instance during training, we increment the corresponding cell in the matrix, letting us see and draw conclusions based on how our model performed. 
-For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative.
-The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class.
-For instance, if determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{n}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.
+Following are sample dialogues from two of the corpora to showcase how the speakers, dialogues and their respective emotions are structured.
+Additionally, we want to demonstrate the difference in topics of the dialogues that might be important to notice when selecting a corpus for a problem.
 
-%kan snacka om Type I och Type II errors om mer text behövs.
+\subsection{Corpus Structure}
+Each of our corpora are structured similarly.
+They contain a train set, a test set, and development set, usually shortened to dev or val(for validation).
+These sets are all structured identically and containing utterances randomly selected from the full corpus, with the bulk of the utterances in train, and test and val of approximately the same size.
+As the name implies, the train set contains the data that the models sets out to train on.
+The other two are used together to produce our results.
+While the test set is where we get our results from, we do not want to directly access our results from it as there is a risk that we accidentally manually overfit our model when tuning it based on our intermediate results.
+Instead, the result that gets reported is the F1-score in the test set for the epoch where the dev set produced the best F1-score.
 
-\subsection{Accuracy}
-One of the simplest metrics used, accuracy quantifies how many percent of predictions that the model got right and is calculated by taking all instances where our model predicted positive and dividing it with all instances.
+\subsection{MELD}
 
-$$\frac{TP + TN}{TP + TN + FP + FN}$$
+\newenvironment{pag}{\fontfamily{pag}\selectfont}{\par}
 
-If we have a balanced data set with a similar amount of positive and negative instances, this can give valuable information on the performance of the model since getting predictions right is the goal of any model.
-If, however, our set is imbalanced and contains e.g.
-90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless.
+\begin{table}[!ht]
+    \begin{pag}
 
-\subsection{Precision}
-Precision is a metric that quantifies how many of the instances predicted positive that actually were true.
-It's calculated by taking all of the instances where the model predicted positive and was correct, and dividing it by all the instances where it predicted positive.
+    \centering
+    \begin{tabular}{lll}
+Rachel: & Thank you. Oh Joey and look at this crib! It's so cute! & \textit{Joy} \\
+Joey: & I know! I found it on the street. & \textit{Joy} \\
+Rachel: & Are you serious, Really?! It's in such good condition. & \textit{Surprise} \\
+Joey: & Yeah. & \textit{Neutral} \\
+Rachel: & Wow! Whoa-whoa what's under the covers? & \textit{Surprise} \\
+Joey: & I don't know. & \textit{Neutral} \\
+Rachel: & It's moving. & \textit{Fear} \\
+Joey: & Ew. & \textit{Disgust} \\
+Rachel: & It's still. It's got a tail! Get it out of here! Get it out of here!! & \textit{Fear} \\
+Joey: & Ooh! Ah! Okay! & \textit{Fear} \\
+    \end{tabular}
+    \label{tab:dialogue_meld}
+\end{pag}
+\end{table}
 
-$$\frac{TP}{TP + FP}$$
+\subsection{IEMOCAP}
+IEMOCAP determines the emotion of an utterance through the response of four independent observers.
+If the observers can reach a clear consensus, that consensus is chosen as the motion, otherwise, it is put as undecided.
 
-Precision gives us information on how accurate the model is when predicting an instance to be true.
-This is useful when dealing with instances where it's important to not give a false positive.
-If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for a innocent person to be determined a criminal.
-Precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, it's likely that said person is guilty.
+\begin{table}[!ht]
+    \begin{pag}
 
-When looking at multiple classes, the formula remains the same but with TP and FP representing the sum of all TPs and FPs
+    \centering
+    \begin{tabular}{lll}
+Female: & Uh-, So I got some good news. & \textit{Happiness} \\
+Male: & You're not pregnant or anything. & \textit{Neutral} \\
+Female: & Pregnant? What are you talking about? No. & \textit{Surprise} \\
+Male: & I know you've been trying to have a baby, right? & \textit{Neutral} \\
+Female: & I have some better news than that. & \textit{Happiness} \\
+Male: & Oh, really? & \textit{Undecided} \\
+    \end{tabular}
+    \label{tab:dialogue_iemocap}
+\end{pag}
+\end{table}
 
-\subsection{Recall}
+%female 2ggr för olika känslor mitt i dialog
 
-How many did we get right, of all that were actually right
+\subsection{DailyDialog}
 
-False negative, such as illness
-\subsection{F1-score}
-
-\subsection{Loss}
-
-\section{Overfitting}
-
-\section{Baseline - COSMIC}
-\begin{itemize}
-    \item Dataset till input måste följa Googles standard som de följer vid speech to text. 
-\end{itemize}
-
-\section{Theory}
-NLP technique, categorization, ML, supervised/unsupervised,  
-Mechanical turks to 
-Finns både text och audio på MELD, \citep{zhang2019modeling} visar att text för sig är bättre än ljud men att text+ljud > text
-train, dev(validation), test, man testar på dev under development, först på test efter man är klar
-
-\chapter{Implementation}
-
-\section{Datasets}
+\ifx false
+Subject to removal
 Filip
@@
 \citep{beaver2020towards}
-\chapter{Speech-to-text}
-Vi antar att det är 2 speakers i samtalet, nämn för- och nackdelar. Ta med i limitations
-Vi har endast stöd för engelska pga limits i datamängd, dokumentation. 
-Google diarization är i Beta.
-Använda kategoriserade ord i Googles STTs speechContexts och därefter använda ML igen för att få högre precision på kategorisering. 
-Sample rates 8000 vs 48000 t.ex.
 
-Cased eller uncased? Hur fungerar google speech-to-text med cased/uncased?
+\fi
 
-\chapter{Keyword Extraction}
-\chapter{Emotion Analysis}
-6 eller 4 känslor
-Sentiment analysis, varför valde vi inte det?
-
-\chapter{Evaluation}
-
-\section{Results}
-
-These are our results
+\chapter{Results}
 
@@
 % \usepackage{multirow}
+\begin{table}[h!]
 \begin{center}
-\begin{tabular}{|c|cc|cc|cc|cc|}
+\begin{tabular}{lcccc}
 \hline
-\multirow{2}{*}{Methods} & \multicolumn{2}{c|}{MELD 34}            & \multicolumn{2}{c|}{DailyDialog 45}     & \multicolumn{2}{c|}{IEMOCAP 34}         & \multicolumn{2}{c|}{EmoryNLP 23}        \\ \cline{2-9}
-                         & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss \\ \hline
-Cosmic                   & 64.14                         & 2.68 & 60.01                         & 2.94 & 58.77                         & 0.96 & 42.55                         & 2.48 \\ \hline
-Vår modell               &                               &      &                               &      &                               &      &                               &      \\ \hline
+\textbf{W-avg F1 score} & \textbf{MELD}           & \textbf{IEMOCAP}        & \textbf{EmoryNLP}  \\ \hline
+KET                     & 53.37           & 59.56          & 34.39 \\
+CNN                     & 55.02           & 52.04          & 32.59 \\
+DialougeRNN             & 57.03           & 62.57          & 31.70 \\
+DialogXL                & 62.41           & 65.94          & 34.73 \\
+COSMIC                  & 65.21           & 65.28          & \textbf{38.11} \\ \hline
+CASTLE                  & \textbf{65.27} & \textbf{66.13} & 37.86\\ \hline
 \end{tabular}
 \end{center}
+\caption{Comparison of the weighted average F1-score between CASTLE and other models taken from \citet{ghosal2020cosmic} and \citet{Shen2020DialogXLAX}.}
+\end{table}
 
-\section{Discussion}
-This is the discussion about our results
+%Ha samma storlek på tabellerna
+
+\ifx false
+\begin{table}[h!]
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+Time per epoch & MELD             & IEMOCAP          & EmoryNLP          \\ \hline
+COSMIC         & XXX s            & XXX s          & XXX s                  \\ \hline
+CASTLE         & \textbf{46.16 s} & \textbf{27.10 s} & \textbf{43.02 s}       \\ \hline
+\end{tabular}
+\end{table}
+\fi
+
+%As we can see CASTLE obtains sota
+%Prata om att dessa resultaten är state-of-the-art och vad det innebär
+
+%These are our results
+
+\chapter{Discussion}
+%Changed architecture for increased performance, then focused primarily on reducing runtime
+
+%Vårt main mål var att förbättra performance med LSTM, snabbare tid blev en biprodukt.
+LSTM gjorde EmoryNLP snabbare ty default parametrar.
+LSTM+hyperparametrar gjorde MELD och IEMOCAP snabbare.
+Diskutera vidare.
+
+%We compare our results to the other baselines results that COMSIC presents in their report.
+%When we started working with the project our main goal was in creating that produced state-of-the-art results
+
+%\section{Discoveries}
+
+\section{Architectural Differences}
+%LSTM är enligt teorin vi har läst bättre men tar längre tid
+%LSTM är en mer avancerad design. Leder ofta till bättre men tar längre tid. Tar upp mer resurser
+
+\section{Performance}
+%Kan inte skriva förrän state-of-the-art på Emory
+%Our results are the average of  10 samples since we wanted them to be comparable to COSMICs.
+
+\section{Execution Time}
+%LSTM tar längre tid (Tror vi)
+
+\section{Limitations}
+
+When this project started, the scope was a lot larger than the one explored during this thesis.
+Originally the model was intended to also perform keyword extraction in addition to emotion recognition.
+This was based on discussions with the company sponsoring this project, where the idea was a model that could identify the topic of the conversation and the emotions of the parties involved.
+There, unfortunately, does not exist a corpus fitting for this task
+
+%Tid på Colab max 24h
+%Tid på exjobb max 20 veckor
+%Tid för hyperparametrization
+%Reducerat scope.
+Ju mer överdrivet desto bättre.
+Vi tänkte från början ha recognition OCH keyword extraction.
+Speech-to-text.
+Vårt originalscope baserades på att vi hade ett privat dataset att jobba på t.ex.
+telekom.
+%Kan inte köra ML på våra datorer pga ingen GPU -> kör på cloud
+\subsection*{DailyDialog}
+%No resources available to run this with LSTM 
+Originally we intended to have CASTLE run on the same corpora that COSMIC produced results for.
+However, due to the previously discussed hardware limitations, we did not have enough memory to run CASTLE on DailyDialog.
+Despite continuous attempts at optimization, we were unable to complete a finished run with the corpus and instead repeatedly ran out of memory.
+The reason for this is twofold.
+Firstly, DailyDialog is close to a magnitude larger than the other corpora used in this experiment.
+Secondly, LSTM is a complex architecture with more gates and calculations which leads to higher demand in hardware.
+As Google Colab has a limited amount of memory for us to utilize, we did not have to option to run the corpus on the service.
+On the other hand, with access to more memory there is a great possibility that we would produce close to, if not, state-of-the-art results for DailyDialog as well.
+
+\section{Future Work}
+
+Even though CASTLE achieves state-of-the-art across all corpora, there are still opportunities for further enhancements. 
+%For a future usage of CASTLE it would be beneficial to have access to machines with 
+%Provided with sufficient hardware resources CASTLE would be able to use all of its potential.
+
+A natural continuation for the project is to expand the scope to what it initially was defined as.
+The first proposed goal of the project was to create a model for analyzing written customer service conversations and using keyword extraction as a complement to said analysis.
+CASTLE is undoubtedly able to analyze written conversations with great success.
+However, the corpora used for training CASTLE are not completely suited for customer service specific dialogues.
+When searching for appropriate corpora there was no customer-service themed data set annotated with emotions available.
+One possibility is to create such a corpus either by using an already existing customer service related corpus and annotating it or by creating a corpus from scratch using recorded customer service conversations as a basis.
+
+%Lack of resources, vad man kan göra med mer resources. Man är inte restricted till services som t.ex. Colab som hindrar en att hogga resurser längre än 24h. 
+%Utöka scope till vad det var från början vilket var, emotion recognition i samtal. Både i realtid och transcribed calls. Keyword extraction kan användas som ett supplement. 
+%Skapa ett dataset specifikt för customer service
+%Ytterligare ändringar som kan ge state-of-the-art
+%Mer utförliga hyperparameter optimization.
 
 \chapter{Conclusions}
+%Ett tydligt exempelt på när teori inte överensstämmer med verkligheten. 
+%Able to produce 
+%Ta med lägga till som residual i future work.
 

### 2021-01-16 msccls/MScTemplateCS.tex
pierre comments taken care of

@@
 \usepackage{calligra}
+\usepackage{nameref}
 
@@
 \acknowledgements{
-If you want to thank people, do it here, on a separate right-hand page. Both the U.S. \textit{acknowledgments} and the British \textit{acknowledgements} spellings are acceptable.
 
-We would like to thank Lennart Andersson for his feedback on this template.
+We would like express our gratitude, first and foremost, to our supervisor Pierre. There is no way that this project would have been completed without your guidance, at least not in time.
 
-We would also like thank Camilla Lekebjer for her contribution on this template, as well as Magnus Hultin for his popular science summary class and example document.
+We would also like to thank \textit{Telavox} and our company supervisor Sara for our weekly meetings, where much of our planning and reflection were made.
 
-Thanks also go to the following (former) students for helping with feedback and suggestions on this template: Mikael Persson, Christoffer Lundgren, Mahmoud Nasser.
+As the COVID-19 epidemic created some unusual circumstances, we would like to give a special thanks to \textit{Åke P Fågel \& Vilt} for providing us with a stable office to work out of. 
 }
@@
 
-%ML är automagiskt
+%ML är automagiskt - PopSci
 
@@
 According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notation.
-Since natural languages naturally evolve as time goes on and their adherence to structures diminish, it is difficult to define explicit rules to follow.
+Since natural languages naturally evolve as time goes on and their adherence to structure diminish, it is difficult to define explicit rules to follow.
+For this reason, there is a need for machine learning models to aid the computer in identifying these diffuse rules.
 
-%prata om emotion recognition
+In order to feed the model with effective information, the input is structured and inserted into a data set.
+This way, the computer has an easier way of processing and analyzing the data.
+When handling data sets constructed using exclusively textual input, the data sets are instead referred to as \textit{corpora}.
 
-%The LSTM architecture \citep{ntlk2009} is something....
-
-%\begin{itemize}
-%    \item Förklara vad systemet är
-%    \item Berätta om företaget Telavox
-%\end{itemize}
+%?? and hence the need of machine learning to replace rules?? ?? You need also to introduce more specifically your task ??
 
@@
 This goal can be very difficult to achieve and there is no straightforward way to do so.
-By analyzing prior conversations it is possible for companies to gain valuable insights in the most effective way to please the customer.
+By analyzing prior conversations, it is possible for companies to gain valuable insights in the most effective way to please the customer.
 Customer service is however a massive industry and it would take countless hours to analyze the conversations from even a single company manually.
+
 The company sponsoring this project approached us with an idea to try and solve this problem using machine learning.
-A usual technique for these types of problems is \textit{sentiment analysis} where utterances are labeled either as positive or negative.
+A usual technique for these types of problems is \textit{sentiment analysis}, where utterances are labeled either as positive or negative.
 We instead decided to take another route, namely \textit{emotion recognition}.
@@
 This allows an analysis with greater specificity.
+
 In order to further the academic relevance of this project we decided that we would attempt to contribute to the field of emotion recognition by achieving state-of-the-art results.
 
+\section{State-of-the-art}
+In this paper, the term state-of-the-art will be recurringly mentioned as it is a fundamental part of our aim and result.
+The term refers to the highest level of achievement accomplished.
+A common goal in machine learning projects is to attempt to raise the bar for the current state-of-the-art results across various metrics for the task that the project is trying to solve.
+Presenting state-of-the-art results is an excellent approach in proving that an implementation or algorithm is valid and valuable.
+
+%Description of what it is and why it plays a big part in advancing algorithms and models. Deepminds Alphafold has led to groundbreaking results in a competitive environment.
+
 \section{Telavox}
-Telavox AB is a Swedish telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service.
-As all of their customers deal in customer service they would all benefit from the solution previously discussed.
+We carried out this project at Telavox AB.
+Telavox is a Swedish telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service.
+As all of their customers deal in customer service, they would all benefit from the solution previously discussed.
 Telavox therefore has a great interest in being able to offer this as a service.
@@
 \section{Research questions}
+\label{research_questions}
 The primary aim of this thesis is to create a emotion recognition model that is able to produce state-of-the-art results.
@@
 \begin{itemize}
+    \item \textit{Is it possible to create a model that fulfills our aims?}
+%?? THe first question is can we create such a model??
+    \item \textit{Could a different architecture than the current state-of-the-art produce better results?}
+%?? here your statement is too specific as you did not introduce the terms. Use vague words such as What are the different option to design an architecture and train a model?? 
+%?? I have swapped the items ??
     \item \textit{How does our model compare to other models in the same field evaluating on the same metrics?}
-    \item \textit{Is it possible for an LSTM architecture to produce better results than a GRU one for emotion recognition?}
     \item \textit{After implementing a fully functioning model, how can we adapt it to predict utterances in real-time?}
+    %Snacka om detta i limitations
 \end{itemize}
@@
 
-\section{Contributions}
-%Potentialen av applications 
-\section{Related work}
-
 \chapter{Theory}
 
-The celebrity of computer science at this time is undoubtedly artificial intelligence (AI).
+The celebrity of computer science at this time is undoubtedly \emph{artificial intelligence} (AI).
 A much more complex subject than most people know.
 Most of the complex problems in AI are solved using technologies within its sub-fields, which are illustrated in Figure \ref{fig:map}, rather than traditional artificial intelligence such as robotics \citep{bajwa_2020}.
-In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model.
+
+%In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model.
 Firstly, the model should be created with deep learning algorithms.
@@
 Secondly, to effectively process the utterances it has been tasked to analyze, a word vector is often used to store the most commonly used words for easier analysis.
+??
+I propose this instead??
+
+In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model:
+\begin{enumerate}
+\item As shown by the recent progress in the field \citep{deep_survey}, the model should be created with deep learning algorithms.
+This should enable it to find patterns that it otherwise could not be able to.
+\item To effectively process the utterances, deep learning requires to collect an annotated corpus and design an appropriate representation for the words.
+\end{enumerate}
 
@@
 \section{Machine Learning}
-Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program it.
+Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program the system.
 By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem.
 In \textit{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number.
-The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs information that allows deducing the label y for this feature vector.)
+The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector $\mathbf{x}$ as input and outputs information that allows deducing the label $y$ for this feature vector.)
+
 The other main type of task in machine learning is \textit{unsupervised learning}.
 The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector.
-The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs either another vector or a value.
+The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector $\mathbf{x}$ as input and outputs either another vector or a value.
 
@@
 A neural network is a collection of connected nodes or neurons that was first mentioned in \citet{mcculloch1943}.
-A neuron \citep{dawson1998ann} is an information-processing unit that receives inputs and produces a single output that can be sent to multiple other neurons.
+A neuron \citep{dawson1998ann} is an information-processing unit that receives multiple inputs and produces a single output that can be sent to multiple other neurons.
 To find the output of a neuron the weighted sum of all inputs is weighted by the weights all connecting to the neuron.
@@
     \centering
-    \includegraphics[width=12cm]{msccls/explanatory_images/single_neuron.png}
-    \caption{An example of a single neuron.}
+    \includegraphics[width=12cm]{msccls/explanatory_images/neuron_new.png}
+    \caption{An example of a single neuron. From \citet{wiki:neuron_image}.}
     \label{fig:neuron}
@@
 These networks fall under the deep learning umbrella.
-Deep learning has several closely relevant definitions, where \citet{deng2014deep} mentions the following:
+Deep learning has several closely related definitions, where \citet{deng2014deep} mentions the following:
 
-\begin{quote}{A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification}
+\begin{quote}{A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification.}
 \end{quote}
 
-Each of the layers in a deep learning network manipulates their input data in some way, and the result is stored in the layer's weights.
+Each of the layers in a deep learning network manipulates its input data in some way, and the result is stored in the layer's weights.
 The results are monitored with the help of the \textit{loss function}.
 The loss function returns a distance score that indicates how far off the network's previous prediction was from the true values \citep{franoischollet2017learning}.
-The process of reducing the distance score that the loss function produces is called training.
+The concept of reducing the distance score that the loss function produces is called training.
+
 Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi-variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.
@@
 
-One of the simpler neural network classes used in deep learning is recurrent neural networks (RNN) and they process either sequential data or time-series data \citep{lipton2015critical}.
-Since the inputs are sequential, all the inputs are dependant on each other in one way or another.
+One of the simplest neural network classes used in deep learning is \emph{recurrent neural networks} (RNN).
+RNN process either sequential data or time-series data \citep{lipton2015critical}.
+Since the inputs are sequential, all the inputs are dependent on each other in one way or another.
 RNNs are therefore suitable for processing sequences of text, audio, and video where every computation will be performed on each element in the sequence.
 
-One of RNN's weaknesses is that they suffer the \textit{vanishing gradient problem} \citep{hochreiter1998}.
-Theoretically, RNNs can process arbitrarily long sequences, in practice, they are limited to the length of the sequence.
+One of RNN's weaknesses is that they suffer from the \textit{vanishing gradient problem} \citep{hochreiter1998}.
+While RNN's theoretically can process arbitrarily long sequences, in practice, they are limited to the length of the sequence.
 If the sequence is too long, the network will struggle to try to look back in time longer than a few steps away.
@@
     \centering
-    \includegraphics[scale=0.3]{msccls/explanatory_images/rnn.png}
+    \includegraphics[scale=0.5]{msccls/explanatory_images/rnn.png}
     \caption{A simple RNN node}
@@
 
+\subsection{Long Short-Term Memory -- LSTM}
+%?? I would have this section before GRU ??
+
+Another widely used recurrent neural network architecture is Long Short-Term Memory (LSTM).
+It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem that might occur in recurrent neural networks and has barely been changed since.
+The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.
+
+As illustrated in Figure \ref{fig:lstm_node}, a LSTM node consists of an \textit{input gate}, an \textit{output gate}, a \textit{forget gate} and a \textit{cell state}. 
+\begin{itemize}
+    \item The input gate controls the extent of data to write onto the internal cell-state.
+    \item The output gate controls the extent of data to pass as output. 
+    \item The forget gate that decides how much of the previous data is irrelevant and will be ignored. 
+    \item The cell state stores the previous state of a node and passes it on to its following node. 
+\end{itemize}
+
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[scale=0.5]{msccls/explanatory_images/lstm.png}
+    \caption{A LSTM node} 
+    %?? You need to mention the source of your figure if it is not yours??
+    \label{fig:lstm_node}
+\end{figure}
+
 \subsection{Gated Recurrent Unit -- GRU}
 %Byta alla "node" till "cell" för att lättare kunna referera till LSTM-celler i CASTLE
-A gated recurrent unit (GRU) is recurrent neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks.
+A \emph{gated recurrent unit} (GRU) is recurrent neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks.
+
 Basic GRU nodes  are primarily constructed of an \textit{update gate} and a \textit{reset gate}.
-The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to be disregarded henceforth.
-Although fairly new it is already a well-established variant of recurrent neural networks.
-In comparison to LSTM layers, GRU layers are cheaper to run since they use fewer parameters, but have less representational power.
-It has been shown that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers \citep{Gruber2020AreGC}.
+The update gate regulates the amount of information that is passed along to the next state, while the reset gate, much like the forget gate of LSTM, decides the amount of past information that is going to be disregarded henceforth.
+
+Although fairly new, it is already a well-established variant of recurrent neural networks.
+In comparison to \textit{LSTM} layers, GRU layers are cheaper to run since they use fewer parameters, but have less representational power.
+It has been shown by \citet{Gruber2020AreGC} that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers.
+
+%?? You cannot cite LSTM without describing them first. In addition GRU are newer than LSTMs??
 
@@
 
-\subsection{Long Short-Term Memory -- LSTM}
-Another widely used recurrent neural network architecture is Long Short-Term Memory (LSTM).
-It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem that might occur in recurrent neural networks and has barely been changed since.
-The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.
+\section{Training Process}
+In order to train a model, there must exist some data set, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method, that feeds the model with input.
+When using a labeled data set, the labels are regarded as the solution to the problem.
+When training a model, the goal is to find values for its weights such that the model correctly labels the given input.
 
-As illustrated in \ref{fig:lstm_node} a LSTM node consists of an \textit{input gate}, an \textit{output gate}, a \textit{forget gate} and a \textit{cell state}. 
-\begin{itemize}
-    \item The input gate controls the extent of data to write onto the internal cell-state.
-    \item The output gate controls the extent of data to pass as output. 
-    \item The forget gate, similarly to the GRUs reset gate, decides how much of the previous data is irrelevant and will be ignored. 
-    \item The cell state stores the previous state of a node and passes it on to its following node. 
-\end{itemize}
-
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[scale=0.45]{msccls/explanatory_images/lstm.png}
-    \caption{A LSTM node}
-    \label{fig:lstm_node}
-\end{figure}
-
-\section{Training Process}
-In order to train a model there must exist some data set, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method, that feeds the model with input.
-When using a labeled data set the labels are regarded as the solution to the problem.
-When training a model the goal is to find values for its weights such that the model correctly labels the given input.
-The training of a model is done step by step in the shape of \textit{epochs}.
-Epochs represent the number of iterations that the model has passed through a data set intended for training the model.
-Following each epoch, a \textit{loss function} measures how bad the model's previous prediction was and produces a value for the \textit{loss}.
+The training of a model is done step by step in the form of \textit{epochs}.
+Epochs represent the number of times that the model iterates through a data set that is intended to train the model.
+Following each epoch, a \textit{loss function} measures how incorrect the model's previous prediction was and produces a value for the \textit{loss}.
 The model will have a perfect prediction when the loss reaches zero, therefore minimizing the loss is desirable.
 
-In order to minimize the loss you need to know what the best values for the weights to do so are. To do this, while training the model iteratively calculates the derivatives of the
+The \textit{backpropagation} algorithm is a fundamental tool for minimizing the loss.
+The algorithm starts with the loss and works its way \textit{backwards} through the layers of the network and applies the chain rule to determine how much impact each of the weights has had on the loss.
+This impact is then used for the \textit{gradient descent} algorithm to alter the weights in the direction of flattening out the derivative.
+While travelling through the function, the ultimate goal is to find where the derivative is zero and this is done for each weight in the network.
+Gradient descent updates the weights to match the results from the backpropagation algorithm.
 
-We want to travel through the function and find the place where the derivative is zero.
+%bild på training process
+
+%Man får inte minimera lossen för aggressivt, det finns en balans, annars sker overfitting vilket är......
+
+%Overfitting is a concept in statistical analysis in which the result corresponds too closely to a certain data set.
+??
+This is not the definition of overfitting.
+Overfitting is when the model fits well the training but fails outside it??
+Instead of the model learning how to notice and predict features, it simply remembers a massive amount of examples from the training set.
+This causes the model to perform very well on that specific data but would likely have a very high error rate on any new data.
+
+%bild på overfitting
+
+%We want to travel through the function and find the place where the derivative is zero.
 This is done for each weight in the network.
@@
 
-direction the weights are taking the model. This is done by evaluating the derivatives of the ...in the network and that is exactly what the \textit{gradient descent} does.
+%In order to minimize the loss you need to know what the best values for the weights to do so are. To do this, while training the model iteratively calculates the derivatives of the
+
+%We want to travel through the function and find the place where the derivative is zero.
+This is done for each weight in the network.
+It is accomplished by iteratively using backpropagation algorithm that finds a value closer the the derivative being zero.
+Gradient descent is then used to update the weights to match the result of the backpropagation algorithm.
+
+%Backpropagation, givet en loss och weights, så tar reda på hur mcyekt varje weight har påverkat lossen. Gradient decent använder denna infon för att uppdatera sina vikter och gör om
+
+%direction the weights are taking the model. This is done by evaluating the derivatives of the ...in the network and that is exactly what the \textit{gradient descent} does.
+
+%so all you have to do is find all the points where the derivative goes to 0 and check for which of these points the function has the lowest value.
+The engine of neural networks: gradient-based optimization.
+Applied to a neural network, that means finding analytically the combination of weight values that yields the smallest possible loss function.
 
 %In gradient descent one is trying to reach the minimum of the loss function with respect to the parameters using the derivatives calculated in the back-propagation.
+
+%Gradient Descent is an optimization technique that is used to find the parameters that will give the least loss.
 
@@
 For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}.
+
 In a confusion matrix, one of the axes represents the actual class of the instance the model is trying to predict and the other represents what instance the model predicted.
@@
 For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative.
+
 The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class.
@@
 \subsection{Accuracy}
-One of the simplest metrics used, accuracy quantifies how many percent of predictions that the model got right and is calculated by taking all instances where our model predicted positive and dividing it with all instances.
+Accuracy quantifies how many percent of predictions the model got right and is calculated by taking all instances, where our model predicted positive and dividing it with all instances:
 
@@
 
-If we have a balanced data set with a similar amount of positive and negative instances, this can give valuable information on the performance of the model since getting predictions right is the goal of any model.
+If we have a balanced dataset with a similar amount of positive and negative instances, then this can give valuable information on the performance of the model since getting predictions right is the goal of any model.
 If, however, our set is imbalanced and contains e.g.
@@
 
-$$ Precision = \frac{TP}{TP + FP}$$
+$$ Precision = \frac{TP}{TP + FP}.$$
 
-Precision gives us information on how accurate the model is when predicting an instance to be true.
-This is useful when dealing with instances where it's important to not give a false positive.
+Precision gives us information on how correct the model is when predicting an instance to be true.
+This is useful when dealing with instances, where it's important to not give a false positive.
 If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for an innocent person to be determined as a criminal.
+
 However, precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, that said person is likely guilty.
@@
 
-$$ Recall =\frac{TP}{TP + FN}$$
+$$ Recall =\frac{TP}{TP + FN}.$$
 
@@
 A high recall can be extremely important when dealing with subjects such as contagious diseases where a result of false-negative can be devastating.
-Recall doesn't take into account how many positive predictions turned out to be false so in cases where that doesn't matter and you want to predict positive at the slightest probability of it being true, optimizing for a high recall is essential.
 
-\subsection{F1-score}
-A combination of the two previous metrics, F1-score is calculated by creating the harmonic mean of \textit{precision} and \textit{recall}.
+Recall does not take into account how many positive predictions turned out to be false so in cases, where that does not matter and you want to predict positive at the slightest probability of it being true, optimizing for a high recall is essential.
 
-$$ F1score = 2 \cdot \frac{Precision \cdot Recall}{Precision + Recall} $$
+\subsection{F1 score}
+A combination of the two previous metrics, the F1 score, is calculated by creating the harmonic mean of precision and recall:
 
-In cases where there is no extra importance in finding either false negatives or false positives, you just want to minimize the risk of errors whatsoever.
-The harmonic mean is a kind of average that is appropriate when working with ratios, which is what both precision and recall are.
+$$ F1score = 2 \cdot \frac{Precision \cdot Recall}{Precision + Recall}.$$
+
+In cases where there is no extra cost associated with either false negatives or false positives, the goal is usually to minimize the total amount of errors.
+The harmonic mean is a type of average that is appropriate when working with ratios, which is what both precision and recall are.
+
 This lets us improve the model concerning either of the variables and not have to think about if, for example, the reduction in Recall was worth the gain in Precision.
@@
 
-The macro F1-score is derived by calculating the F1-score for each class separately and then just summing them up, leading to disproportionate weight on the classes with fewer instances.
+The macro F1-score is derived by calculating the F1-score for each class separately and then just summing them up, leading to an excessive focus on the classes with fewer instances.
 If we assume that we are using a data set where the \textit{neutral} class is represented in 10 000 instances and a less common emotion \textit{fear} in just 50.
@@
 
-When calculating micro F1-score, instead of calculating the f1-scores separately, the f1-score is calculated directly by using the total values for TP, FN, and FP in the Precision and Recall formulas which places no bias on either class.
+When calculating micro F1 score, instead of calculating the F1 scores separately, the F1 score is calculated directly by using the total values for TP, FN, and FP in the Precision and Recall formulas which places no bias on either class.
 If we have an example data set with two minority classes \textit{anger} and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class.
@@
 Since micro calculations take into account the total values for our identification labels and each FN correspond to the FP of another class, $FP = FN$, which in turn means that $Precision = Recall$ in these cases.
+
 Precision and Recall being equal means that their mean also is equal and thus \textit{F1-}$score_{micro} = Precision_{Micro} = Recall_{Micro}$ which incidentally also equals \textit{Accuracy} as this is the proportion of instances we got correct.
-This means that just like Accuracy this is a good metric on balanced data sets but can be deceiving in an imbalanced one.
-
-\section{Loss}
-%might be removed
-
-\section{Overfitting}
-Overfitting is a concept in statistical analysis in which the result corresponds too closely to a certain data set.
-Instead of the model learning how to notice and predict features, it simply remembers a massive amount of examples from the training set.
-This causes the model to perform very well on that specific data but would likely have a very high error rate on any new data.
+This means that just like Accuracy, this is a good metric on balanced data sets but can be deceiving in an imbalanced one.
 
 \section{Natural Language Processing -- NLP}
-Natural languages are the set of languages that humans have evolved naturally without any predetermined strategy like e.g English or Mandarin.
+Natural languages are the set of languages that humans have evolved naturally without any predetermined strategy such as English or Mandarin.
 Making computers understand and analyze natural languages has long been a challenge and one of the first practical attempts was Weaver and Booth's translation project in 1946 trying to crack enemy codes during World War II \citep{Liddy2001NLP}.
-Shortly thereafter \citet{Turing1950} created his famous \textit{Imitation game} or more commonly referred to as the Turing test, which has developed into a test to evaluate whether a computer's natural language response in a real-time written conversation can be indistinguishable from that of a human.
+Shortly thereafter \citet{Turing1950} created his famous \textit{Imitation game}, or more commonly referred to as the Turing test, which has developed into a test to evaluate whether a computer's natural language response in a real-time written conversation can be indistinguishable from that of a human.
+
 One of the reasons that make natural language processing a challenge is the fact that natural languages are minted with unstructured types of data such as inflection or dialects, or using phrases to convey the opposite than the common meaning namely sarcasm.
@@
 
+Instead of working directly with raw input, it first need to be converted into a numerical representation in order for the model to process it \citep{franoischollet2017learning}.
+In the case of textual input this is done by transforming it into a \textit{word vector}.
+
 %Emotion recognition can be applied to data in various modalities, 
 %such as visual, in the form of facial expressions and body language, audible and text-based in the form of dialogues.
-
-\subsection{Word Vectors}
-%Om vi behöver en tie in: Instead of analysing the words themselves, they are converted into numerical representations that are easier for the computer to handle.
-
-Word vectors are a type of word representation where texts are turned into a series of numbers, each word or sometimes phrase instead represented by a number.
 
@@
 The other big issue with it is that no meaning can be extracted from the matrix.
-If our vocabulary contained the words "cat", "dog" and "hamster", we could easily see that these are all domestic animals but in the space of the one-hot vector, "cat" is as close to "dog" as it would be to "airplane".
+If our vocabulary contained the words ``cat'', ``dog'' and ``hamster'', we could easily see that these are all domestic animals but in the space of the one-hot vector, ``cat'' could be as close to ``dog'' as it could be to ``airplane''.
 
 \subsection{Word Embedding}
-To resolve the two aforementioned problems with one-hot encodings, word embeddings were created, also known as dense word vectors. \citep{neuralnetworkmethods} 
+To resolve the two aforementioned problems with one-hot encodings, word embeddings were created, also known as dense word vectors \citep{neuralnetworkmethods}. 
 Here, instead of creating a sparse, hardcoded vector with equal distance between words, they are learned from data and packed into a lower-dimensional vector where each dimension contains some relevant data about the word.
 According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional, or 1024-dimensional are common when dealing with very large vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use a one-hot encoding.
+
 Additionally, the geometric relationship between words can and should reflect the semantic relationship between those same words.
@@
 
-\begin{wrapfigure}{r}{0.25\textwidth}
-\includegraphics[width=0.9\linewidth]{msccls/explanatory_images/embedding_direction.png} 
+\begin{figure}[!ht]
+\centering
+\includegraphics[scale=0.8]{msccls/explanatory_images/embedding_direction.png} 
 \caption{Small example of a word embedding space. From \citet{franoischollet2017learning}.}
 \label{fig:direction}
-\end{wrapfigure}
+\end{figure}
 
-In figure \ref{fig:direction}, the four words \textit{cat}, \textit{dog}, \textit{wolf} and \textit{tiger} are represented on a geometric plane.
+In Figure \ref{fig:direction}, the four words \textit{cat}, \textit{dog}, \textit{wolf} and \textit{tiger} are represented on a geometric plane.
 We can clearly see some of the possible geometric transformations possible by using direction as a tool.
 For instance, the same vector that takes us from \textit{dog} to \textit{wolf} also takes us from \textit{cat} to \textit{tiger}.
-This could be interpreted this vector as "converting a domesticated animal to its wildlife equivalent".
-Another vector will take us both from \textit{cat} to \textit{dog} and from \textit{tiger} to \textit{wolf} which could be interpreted as "converting a feline to a canine".
-Two meaningful transformation vectors that are common in the real world are "gender" and "plural" vectors.
-By applying a "female" vector to \textit{king}, \textit{queen} is obtained and the same vector returns \textit{actress} if instead used on \textit{actor}.
-"Plural", as should seem obvious by now, if used on \textit{actor} will return \textit{actors}.
+This could be interpreted this vector as ``converting a domesticated animal to its wildlife equivalent''.
+
+Another vector will take us both from \textit{cat} to \textit{dog} and from \textit{tiger} to \textit{wolf} which could be interpreted as ``converting a feline to a canine''.
+Two meaningful transformation vectors that are common in the real world are ``gender'' and ``plural'' vectors.
+By applying a ``female'' vector to \textit{king}, \textit{queen} is obtained and the same vector returns \textit{actress} if instead used on \textit{actor}.
+``Plural'', as should seem obvious by now, if used on \textit{actor} will return \textit{actors}.
+
 Typically there are thousands of these types of interpretable direction vectors in a large word embedding.
-
-This bit is after the picture, that should be to the right of the last paragraph
 
@@
 %Summary of what the chapter is about
-The following chapter is mainly focused on giving a clear description of what we have done for the duration of the project.
-Firstly we will describe our model \textit{CASTLE}, how it has been implemented and what changes were made from the baseline \textit{COSMIC}.
-Our initial goal was to have our model produce state-of-the-art results in emotion recognition which later on changed to a more primary focus on optimizing the run time instead.
+This chapter is mainly focused on giving a clear description of what we have done for the duration of the project.
+Firstly we will describe the model that we used as a baseline.
+Following, we will cover our model \textit{CASTLE}, how it has been implemented and what changes were made from the baseline \textit{COSMIC}.
+Lastly, we will describe our setup and what corpora we used.
+%??In your text, you present COSMIC first. Change your intro so that it corresponds to your text??.
+
 We based our research strategy on the work by \citet{bouthillier:hal-02447823} that summarizes and reviews the experimental procedures that are most commonly used in the field of machine learning.
@@
 This project has been developed using the COSMIC model \citep{ghosal2020cosmic} as a baseline.
-COSMIC is a GRU-based model implemented with PyTorch that has achieved state-of-the-art results across four data sets commonly used in emotion recognition.
+COSMIC is a GRU-based model implemented with PyTorch that has achieved state-of-the-art results across four corpora commonly used in emotion recognition.
 The idea behind the model is when classifying an utterance, the emotions of the utterances both preceding and succeeding it are also taken into account.
@@
 For each utterance, there exist multiple GRU-cells that predict both the intent of the speaker and the response of the listener at a certain time step.
-These cells are used not only to predict the emotion of the current utterance but also used for the prediction of emotions in other utterances related to said utterance.
+These cells are used not only to predict the emotion of the current utterance, but also used for the prediction of emotions in other utterances related to said utterance.
+
+COSMIC makes use of five different states to determine what emotion to predict. 
+The four following states interact heavily with each other in a codependent manner.
+
+\begin{itemize}
+    \item \textit{Context state} is a state shared between speaker and listener. It manages the current overall emotional state of the conversation. Affects both the Internal and External state.
+    \item \textit{Internal state} tracks for each participant what emotions they are feeling but not necessarily showing. Affects the Intent state and the Context state.
+    \item \textit{External state} portrays for each participant what emotions they are expressing to the other participants through both utterances and reactions. Affects the Context state.
+    \item \textit{Intent state} is a state only present for the current speaker.
+As the name suggests, it represents the speaker's objective is for the current utterance.
+For example, if paying the listener a compliment, the intent could be to make the listener happier.
+This does not affect any other of the three states.
+\end{itemize}
+Internal state and External state are especially codependent as they combine into the sum of a participant's emotional state.
+
+The final state, Emotion state, takes into account the current utterance, the summed emotional state through the Internal and External state, as well as the Intent state to ultimately make a prediction for the current utterance.
+This process is repeated for every utterance in every conversation in the corpus.
 
@@
 
+%?? You could write a bit more on COSMIC ??
+
 \section{CASTLE}
+Building on COSMIC, we designed CASTLE, COSMIC-based AnalySer Taking advantage of LSTM for Emotion recognition, to answer our research questions.
+We have made several improvements on the COSMIC baseline, one of the fundamental being replacing the internal GRU architecture with an LSTM-based architecture as the name of the model indicates.
 
-COSMIC-based AnalySer Taking advantage of LSTM for Emotion recognition, CASTLE, was developed as the solution to our research questions.
-We have made several improvements on the COSMIC baseline, one of the fundamental being replacing the internal GRU architecture with an LSTM-based architecture as the name of the model indicates.
-At the start of the project we read some recently published research in the field of deep learning, mainly the article by \citet{edseee.922172720200601} that states that GRU is not as suited for small pieces of text and large data sets as LSTM.
+At the start of the project, we read some recently published research in the field of deep learning, mainly the article by \citet{edseee.922172720200601} that states that GRU is not as suited for small pieces of text and large corpora as LSTM.
 This is where we got the inspiration to test whether or not the architectural change to LSTM could improve the results.
@@
 As we aimed to obtain as much control as possible over how the model behaved, we excluded this and other third party features that restricted our domain of control.
+
 COSMIC is also implemented with a \textit{residual} connection \citep{he2015deep} which allows the model to redefine the layers as learning residual functions referencing the layer inputs.
@@
 %We chose COSMIC as our baseline because SOTA, more experienced in PyTorch
+\subsection{Hyperparameters}
+CASTLE's performance depends on three different hyperparameters. The ones we chose are \textit{dropout}, \textit{batch size}, and \textit{learning rate}. \citep{hyperparameters}:
+\begin{enumerate}
+    \item Dropout is a value between 0 and 1 that regulates what percentage of learned data that the model randomly chooses to forget.
+This is used in order to combat overfitting and make sure the model is actually predicting instead of just learning the structure of the corpus.
+    \item  Batch size governs how many training examples will be utilized in each iteration. A larger value increases the accuracy of the gradient in trade off for a higher memory cost.
+    \item  Learning rate defines the speed of which the model updates its parameters.
+A high learning rate speeds up the learning at the risk of not converging while a low value will more likely converge, sacrificing speed in the process.
+\end{enumerate}
 
+To optimize our hyperparameters, we implemented the \texttt{Tune} framework offered by \texttt{Ray}.
+This framework allowed us to efficiently experiment with many different configurations and provided a simple way to run trials concurrently.
+Most of our optimization were done with a combination of grid search and random search.
+Random search is a function with two parameters \textit{start} and \textit{end}, and for each trial it randomly selects a float from in between the two values to use for the given hyperparameter.
+Grid search instead takes in a list of floats as a parameter and runs a trial for each value in the list.
+Multiple grid searches can very easily take a large amount of time to run as all possible combination of values across all grid searches will be tested in separate trials.
+
+Our approach was to start out with random search on parameters to get a sense on what range of values produced optimal results.
+Once we had established the magnitude of which we wanted our parameters to be, we tested out various combinations with grid search, keeping the ones that produced promising results and replacing the others with new trials.
+
+\subsection{Comparison with Current Practices}
 When deciding what configuration to use for our experiments, we based our decisions on the empirical studies made by  \citet{bouthillier:hal-02447823}.
-In their report they surveyed authors whose papers were presented at the peer-reviewed conferences NeurIPS 2019 and ICLR 2020.
+In their report, they surveyed authors whose papers were presented at the peer-reviewed conferences NeurIPS 2019 and ICLR 2020.
 As the work is based on answers from relevant authors in the machine learning community, we found it appropriate to use the answers as guidelines for our setup.
 Instead of relying on our own intuition, basing it on the broad consensus grants the experiment a certain sense of validity.
-In Figure \ref{fig:HAL} we have presented the questions and responses we took into consideration.
+In Figure \ref{fig:HAL}, we have presented the questions and responses we took into consideration.
 
 CASTLE is fine-tuned on three hyperparameters where a combination of \textit{grid search} and \textit{random search} was used.
-On average we performed 20 sets of different values for the three hyperparameters across all the data sets.
-CASTLE is was intended to be evaluated on four different corpora, or datasets, but ultimately only provided results for three of them.
+On average, we performed 20 sets of different values for the three hyperparameters across all corpora.
+CASTLE was intended to be evaluated on four different corpora but ultimately only provided results for three of them.
 Our results are based upon 10 consecutive runs with the same setup.
@@
 Comparing this to the responses observed in \ref{fig:HAL}, we are in line with the majority in all but one instance, (f), where the majority base their results on 5 or less samples and we instead chose to base it on 10 samples as a higher sample size tends to become a more reliable result.
+
+\subsection{Equipment}
+As for our hardware setup, we were not able to run and train the model on our laptops as they lack a GPU.
+To execute our code, we instead utilized Google Colab, a cloud based service where users can remotely run their code on Google servers that are built with more powerful hardware and can handle high demand programs such as machine learning model.
+Using a remote service proved to be perfectly fine for three of the four corpora we aimed to run our model on.
+The fourth one, \textit{DailyDialog}, unfortunately required more memory than we were allocated and could not complete run.
+
+%Hur vi "räknade ut" våra resultat, 10 avg på best valid ->( test)
+%Epochs
+%Google Colab
+%Ray Tune för att studera hyperparametrar och automatisera
+%HAL - diskussion
 
@@
 
-%Hur vi "räknade ut" våra resultat, 10 avg på best valid ->( test)
-%Epochs
-%Google Colab
-%Ray Tune för att studera hyperparametrar och automatisera
-%HAL - diskussion
-
 %Lack of resources
-\section{Optimizing Hyperparameters}
-As previously mentioned, CASTLE is optimized on three different hyperparameters.
-The ones we chose are \textit{dropout}\citep{hyperparameters}, \textit{batch size}, and \textit{learning rate}.
-Dropout is a value between 0 and 1 that regulates what percentage of learned data that the model randomly chooses to forget.
-This is used in order to combat overfitting and make sure the model is actually predicting instead of just learning the structure of the data set.
-Batch size governs how many training examples will be utilized in each iteration.
-A larger value increases the accuracy of the gradient in trade off for a higher memory cost.
-Learning rate defines the speed of which the model updates its parameters.
-A high learning rate speeds up the learning at the risk of not converging while a low value will more likely converge, sacrificing speed in the process.
-
-To optimize our hyperparameters, we implemented the \texttt{Tune} framework offered by \texttt{Ray}.
-This framework allowed us to efficiently experiment with many different configurations and provided a simple way to run trials concurrently.
-Most of our optimization were done with a combination of grid search and random search.
-Random search is a function with two parameters \textit{start} and \textit{end}, and for each trial it randomly selects a float from in between the two values to use for the given hyperparameter.
-Grid search instead takes in a list of floats as a parameter and runs a trial for each value in the list.
-Multiple grid searches can very easily take a large amount of time to run as all possible combination of values across all grid searches will be tested in separate trials.
-
-Our approach was to start out with random search on parameters to get a sense on what range of values produced optimal results.
-Once we had established the magnitude of which we wanted our parameters to be, we tested out various combinations with grid search, keeping the ones that produced promising results and replacing the others with new trials.
 
@@
 
-To evaluate our model we used four popular corpora that are recurring in emotion recognition research work.
-We chose these data sets as we wanted our results to be comparable with the current state-of-the-art, namely COSMIC as well as the common practice observed in Figure \ref{fig:HAL}(e).
-As COSMIC compared its results to several other models using these data sets, the results of CASTLE will be directly comparable with COSMIC and the other models.
+To evaluate our model, we used four popular corpora that are frequently used in emotion recognition research work.
+We chose these corpora as we wanted our results to be comparable with the current state-of-the-art, namely COSMIC as well as the common practice observed in Figure \ref{fig:HAL}(e).
+As COSMIC compared its results to several other models using these corpora, the results of CASTLE will be directly comparable with COSMIC and the other models.
 
-All the following data sets include a \textit{neutral} emotion that is not very elucidatory.
-An emotion is classified as neutral when it can not be classified as any of the other emotions within the corpus using each data sets corresponding methods for classifying the emotions.
+All the following corpora include a \textit{neutral} emotion that is not very descriptive.
+An emotion is classified as neutral when it can not be classified as any of the other emotions.
+Each corpus uses its own set of emotions and method of classification to label said emotions.
 %mention mechanical turks as example? [MELD?]
@@
 
-\begin{itemize}
-    \item \textit{MELD} \citep{poria2019meld} is based on about 13000 utterances and 1400 dialogues from the TV series \textit{Friends} and is an extension to the EmotionLines dataset \citep{chen2018emotionlines}.
+\begin{description}
+    \item[MELD] \citep{poria2019meld} is based on about 13,000 utterances and 1,400 dialogues from the TV series \textit{Friends} and is an extension to the EmotionLines dataset \citep{chen2018emotionlines}.
 MELD is multimodal and  frequently has more than two speakers per dialogue.
-The utterances are labeled with the following emotions: \textit{anger}, \textit{disgust}, \textit{sadness}, \textit{joy}, \textit{surprise}, \textit{fear} and \textit{neutral}.
+The utterances are labeled with the following emotions: \textit{anger}, \textit{disgust}, \textit{sadness}, \textit{joy}, \textit{surprise}, \textit{fear}, and \textit{neutral}.
 
-    \item \textit{IEMOCAP} \citep{busso2008iemocap} is a multimodal, multispeaker corpus with utterances from acting and improvisation scenarios emphasizing on emotional expressions.
-The utterances are labeled with the following emotions: \textit{anger}, \textit{happiness}, \textit{excitement}, \textit{sadness}, \textit{frustration}, \textit{fear}, \textit{surprise}, \textit{other} and \textit{neutral state}.
+    \item[IEMOCAP] \citep{busso2008iemocap} is a multimodal, multispeaker corpus with utterances from acting and improvisation scenarios emphasizing on emotional expressions.
+The utterances are labeled with the following emotions: \textit{anger}, \textit{happiness}, \textit{excitement}, \textit{sadness}, \textit{frustration}, \textit{fear}, \textit{surprise}, \textit{other}, and \textit{neutral state}.
 
-    \item \textit{EmoryNLP} \citep{zahiri2017emoryNLP} is also a corpus with transcripts from the TV series \textit{Friends}.
-In comparison to MELD, EmoryNLP has over 60.000 utterances.
-The utterances are labeled with the following emotions: neutral, joyful, peaceful, powerful, scared, mad and sad.
+    \item[EmoryNLP] \citep{zahiri2017emoryNLP} is also a corpus with transcripts from the TV series \textit{Friends}.
+In comparison to MELD, EmoryNLP has over 60,000 utterances.
+The utterances are labeled with the following emotions: \textit{neutral}, \textit{joyful}, \textit{peaceful}, \textit{powerful}, \textit{scared}, \textit{mad} and \textit{sad}.
 
-    \item \textit{DailyDialog} \citep{lietal2017dailydialog} encompasses numerous daily life topics with human-written dialogues to reflect usual day-to-day conversations.
+    \item[DailyDialog] \citep{lietal2017dailydialog} encompasses numerous daily life topics with human-written dialogues to reflect usual day-to-day conversations.
 The corpus has an average of circa 8 speakers per dialogue and has a total of over 13.000 dialogues with their emotions manually labeled.
 The utterances are labeled with the following emotions: anger, disgust, fear, joy, neutral, sadness, and surprise.
-\end{itemize}
+\end{description}
 
@@
 Each of our corpora are structured similarly.
-They contain a train set, a test set, and development set, usually shortened to dev or val(for validation).
+They contain a training set, a test set, and development set, usually shortened to dev or val (for validation).
 These sets are all structured identically and containing utterances randomly selected from the full corpus, with the bulk of the utterances in train, and test and val of approximately the same size.
-As the name implies, the train set contains the data that the models sets out to train on.
+As the name implies, the training set contains the data that the models sets out to train on.
 The other two are used together to produce our results.
 While the test set is where we get our results from, we do not want to directly access our results from it as there is a risk that we accidentally manually overfit our model when tuning it based on our intermediate results.
-Instead, the result that gets reported is the F1-score in the test set for the epoch where the dev set produced the best F1-score.
+Instead, the result that gets reported is the F1 score in the test set for the epoch where the dev set produced the best F1 score.
 
-\subsection{MELD}
+\subsection{Corpus Excerpts}
 
+\paragraph{MELD.} The annotated emotions in MELD are chosen by the majority verdict of five independent \textit{Amazon mechanical turks}.
+Each one of these are tasked to manually examine utterances one by one and then select what emotion out of the options is most suited for that utterance.
+%?? Here have afew intro words as for IEMOCAP ??
 \newenvironment{pag}{\fontfamily{pag}\selectfont}{\par}
@@
 
-\subsection{IEMOCAP}
+\paragraph{IEMOCAP.}
 IEMOCAP determines the emotion of an utterance through the response of four independent observers.
@@
 
-\subsection{DailyDialog}
-
 \ifx false
@@
 \chapter{Results}
+\label{result_chapter}
 
@@
 % \usepackage{multirow}
-\begin{table}[h!]
+\begin{table}[!ht]
 \begin{center}
@@
 \end{center}
-\caption{Comparison of the weighted average F1-score between CASTLE and other models taken from \citet{ghosal2020cosmic} and \citet{Shen2020DialogXLAX}.}
+\caption{Comparison of the weighted average F1-score between CASTLE and other models taken from \citet{ghosal2020cosmic} and \citet{Shen2020DialogXLAX}. State-of-the-art results in bold. }
+\label{results_table}
 \end{table}
+
+As Table \ref{results_table} shows, CASTLE achieves state-of-the-art results for both the MELD and IEMOCAP corpora.
+While not producing state-of-the-art results for EmoryNLP, CASTLE manages to produce a score within 99$\%$ of the current state-of-the-art.
 
@@
 \chapter{Discussion}
-%Changed architecture for increased performance, then focused primarily on reducing runtime
-
-%Vårt main mål var att förbättra performance med LSTM, snabbare tid blev en biprodukt.
-LSTM gjorde EmoryNLP snabbare ty default parametrar.
-LSTM+hyperparametrar gjorde MELD och IEMOCAP snabbare.
-Diskutera vidare.
+In this chapter, we discuss and analyze the results of CASTLE as well as the project itself.
 
@@
 
-\section{Architectural Differences}
+\section{Performance}
+
+As stated in the \nameref{result_chapter} Chapter, CASTLE accomplishes state-of-the-art results in multiple corpora annotated for emotion recognition.
+This makes the case for CASTLE being an improvement over current state-of-the-art models, namely COSMIC for MELD and DialogXL for IEMOCAP.
+EmoryNLP was the only corpus that we managed to run CASTLE on without achieving state-of-the-art results.
+
+There might be several reasons for this but we believe that one of them is more likely than the others.
+As EmoryNLP was the last corpus that we optimized CASTLE for, there was not enough time to thoroughly tune the hyperparameters.
+The method we used for optimizing hyperparameters was to initially use random search to find relevant sets of parameters, where the leading values were then iterated through with the help of grid search.
+
+Working in this way is somewhat of a scattershot approach where we depend unpleasantly much on the results of random searches.
+While the random searches clearly produced competitive results, there is nothing to indicate that whether or not an optimal set of values was overlooked.
+Since CASTLE produces results within 99$\%$ of current state-of-the-art, it is our belief that with additional time and resources CASTLE could produce state-of-the-art results for EmoryNLP as well.
+
+The evidence demonstrates that CASTLE outperforms its competitors on the MELD and IEMOCAP corpora while remaining competitive on EmoryNLP and with inconclusive results for DailyDialog.
+This fulfills the primary aim of our thesis that we defined in section \ref{research_questions} and makes CASTLE one of, if not, the most viable systems to adapt for practical emotion recognition applications.
+
+The primary improvement we implemented was replacing an existing GRU layer with a more complex LSTM one.
+As mentioned earlier, the theory states that LSTM layers is likely to outperform GRU layers, when there is enough data, at the expense of computation power.
+Since LSTM is a more complex structure with an additional gate as well as having the concept of utilizing memory in a cell state, it is no surprise that it has the potential to outperform GRU layers.
+When debating if to use CASTLE or some other model, it would boil down to whether or not increase in certainty is worth the cost in computational power.
+
+%adding feautrés costs resources and time
+
+%Vårt main mål var att förbättra performance med LSTM.
+%Kan inte skriva förrän state-of-the-art på Emory
+%Our results are the average of  10 samples since we wanted them to be comparable to COSMICs. 
+%Avvägning mellan performance och tid
+
+%Changed architecture for increased performance, then focused primarily on reducing runtime
 %LSTM är enligt teorin vi har läst bättre men tar längre tid
@@
 
-\section{Performance}
-%Kan inte skriva förrän state-of-the-art på Emory
-%Our results are the average of  10 samples since we wanted them to be comparable to COSMICs.
-
-\section{Execution Time}
-%LSTM tar längre tid (Tror vi)
-
 \section{Limitations}
 
-When this project started, the scope was a lot larger than the one explored during this thesis.
+At the start of this project, the scope was a lot larger than the one explored during this thesis.
 Originally the model was intended to also perform keyword extraction in addition to emotion recognition.
-This was based on discussions with the company sponsoring this project, where the idea was a model that could identify the topic of the conversation and the emotions of the parties involved.
-There, unfortunately, does not exist a corpus fitting for this task
+This was based on discussions with Telavox, where the idea was a model that could identify the topic of the conversation and the emotions of the parties involved.
+Unfortunately, a corpus fitting for this task does not exist at this time.
+
+%crossroads i att skapa ett corpus som passar. gick inte för tidsbrist
 
@@
 %Kan inte köra ML på våra datorer pga ingen GPU -> kör på cloud
+%real time translation
 \subsection*{DailyDialog}
@@
 Despite continuous attempts at optimization, we were unable to complete a finished run with the corpus and instead repeatedly ran out of memory.
+
 The reason for this is twofold.
@@
 As Google Colab has a limited amount of memory for us to utilize, we did not have to option to run the corpus on the service.
-On the other hand, with access to more memory there is a great possibility that we would produce close to, if not, state-of-the-art results for DailyDialog as well.
+On the other hand, with access to more memory, there is a great possibility that we would produce close to, if not, state-of-the-art results for DailyDialog as well.
 
@@
 
-Even though CASTLE achieves state-of-the-art across all corpora, there are still opportunities for further enhancements. 
+Considering the fact that CASTLE did not produce state-of-the-art results for the EmoryNLP corpus, the most natural continuation for the project is to have CASTLE achieve this goal.
+
 %For a future usage of CASTLE it would be beneficial to have access to machines with 
@@
 
-A natural continuation for the project is to expand the scope to what it initially was defined as.
+Another natural continuation for the project is to expand the scope to what it initially was defined as.
 The first proposed goal of the project was to create a model for analyzing written customer service conversations and using keyword extraction as a complement to said analysis.
+
 CASTLE is undoubtedly able to analyze written conversations with great success.
 However, the corpora used for training CASTLE are not completely suited for customer service specific dialogues.
-When searching for appropriate corpora there was no customer-service themed data set annotated with emotions available.
+When searching for appropriate corpora, there was no customer-service themed corpus annotated with emotions available.
 One possibility is to create such a corpus either by using an already existing customer service related corpus and annotating it or by creating a corpus from scratch using recorded customer service conversations as a basis.
+
+%som vi nämner i limitations
 
@@
 \chapter{Conclusions}
+When we started this project, our goal was to find answers to the research questions mentioned in section \ref{research_questions}.
+
 %Ett tydligt exempelt på när teori inte överensstämmer med verkligheten. 

### 2021-01-18 msccls/MScTemplateCS.tex
klara med exjobbet, nästan

@@
 \usepackage{nameref}
+\usepackage{float}
 
@@
 
- \begin{figure}[!ht]
+ \begin{figure}[H]
     \centering
-    \includegraphics[scale=0.5]{msccls/explanatory_images/rnn.png}
+    \includegraphics[scale=0.35]{msccls/explanatory_images/rnn.png}
     \caption{A simple RNN node}
@@
 
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[scale=0.5]{msccls/explanatory_images/lstm.png}
-    \caption{A LSTM node} 
-    %?? You need to mention the source of your figure if it is not yours??
-    \label{fig:lstm_node}
-\end{figure}
-
 \subsection{Gated Recurrent Unit -- GRU}
@@
     \centering
-    \includegraphics[scale=0.5]{msccls/explanatory_images/gru.png}
+    \includegraphics[scale=0.58]{msccls/explanatory_images/lstm.png}
+    \caption{A LSTM node} 
+    %?? You need to mention the source of your figure if it is not yours??
+    \label{fig:lstm_node}
+\end{figure}
+
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[scale=0.58]{msccls/explanatory_images/gru.png}
     \caption{A GRU node}
@@
 While travelling through the function, the ultimate goal is to find where the derivative is zero and this is done for each weight in the network.
-Gradient descent updates the weights to match the results from the backpropagation algorithm.
+Gradient descent updates the weights to match the results from the backpropagation algorithm \citep{franoischollet2017learning}.
 
@@
 %Man får inte minimera lossen för aggressivt, det finns en balans, annars sker overfitting vilket är......
-
-%Overfitting is a concept in statistical analysis in which the result corresponds too closely to a certain data set.
-??
-This is not the definition of overfitting.
-Overfitting is when the model fits well the training but fails outside it??
-Instead of the model learning how to notice and predict features, it simply remembers a massive amount of examples from the training set.
-This causes the model to perform very well on that specific data but would likely have a very high error rate on any new data.
+%When training a model 
+%Overfitting is a concept in statistical analysis in which the  model fits well on the training set but fails when introducing new examples.
+Instead of the model learning how to notice and predict features, it simply remembers an unreasonable amount of examples from the training set.
+This causes the model to perform very well on that specific data but would likely have a very high error rate on any unseen data.
 
 %bild på overfitting
-
-%We want to travel through the function and find the place where the derivative is zero.
-This is done for each weight in the network.
-It is accomplished by iteratively using backpropagation algorithm that finds a value closer the the derivative being zero.
-Gradient descent is then used to update the weights to match the result of the backpropagation algorithm.
-
-%In order to minimize the loss you need to know what the best values for the weights to do so are. To do this, while training the model iteratively calculates the derivatives of the
-
-%We want to travel through the function and find the place where the derivative is zero.
-This is done for each weight in the network.
-It is accomplished by iteratively using backpropagation algorithm that finds a value closer the the derivative being zero.
-Gradient descent is then used to update the weights to match the result of the backpropagation algorithm.
-
-%Backpropagation, givet en loss och weights, så tar reda på hur mcyekt varje weight har påverkat lossen. Gradient decent använder denna infon för att uppdatera sina vikter och gör om
-
-%direction the weights are taking the model. This is done by evaluating the derivatives of the ...in the network and that is exactly what the \textit{gradient descent} does.
-
-%so all you have to do is find all the points where the derivative goes to 0 and check for which of these points the function has the lowest value.
-The engine of neural networks: gradient-based optimization.
-Applied to a neural network, that means finding analytically the combination of weight values that yields the smallest possible loss function.
-
-%In gradient descent one is trying to reach the minimum of the loss function with respect to the parameters using the derivatives calculated in the back-propagation.
-
-%Gradient Descent is an optimization technique that is used to find the parameters that will give the least loss.
-
-%gradient descent iteratively tries to find a minimum. tries to regulate the weights in order to minimize loss.
-
-%Datasets ger inputs till model, första lagret ..... Modellen har svaren till hands och använder dessa för att styra skeppet i rätt riktning. 
-%There exist many components to training a model .....
-%In \ref{chapter}, backpropagation and gradient descnet among other temrs were mentioned but not extensively explored.
-The training process can be described as the road from a standalone model to having relevant results.
-
-% In this context, learning means finding a set of values for the weights of all layers in a network, such that the network will correctly map example inputs to their associated targets
-
-%Backpropagation
-%Gradient Descent
-%Loss (function)
-%Från början till slut, från en modell med lager till en evaluation, sen fortsätter vi snacka om det under evaluation
-%Overfitting
-%Epochs
 
@@
 This chapter is mainly focused on giving a clear description of what we have done for the duration of the project.
-Firstly we will describe the model that we used as a baseline.
-Following, we will cover our model \textit{CASTLE}, how it has been implemented and what changes were made from the baseline \textit{COSMIC}.
+Firstly we will describe the model that we used as a baseline, \textit{COSMIC}.
+Following, we will cover our model \textit{CASTLE}, how it has been implemented and what changes were made from COSMIC.
 Lastly, we will describe our setup and what corpora we used.
@@
 \subsection{Corpus Excerpts}
-
+\label{corpus_excerpts}
 \paragraph{MELD.} The annotated emotions in MELD are chosen by the majority verdict of five independent \textit{Amazon mechanical turks}.
@@
 
-\begin{table}[!ht]
+\begin{table}[H]
     \begin{pag}
@@
 \chapter{Discussion}
-In this chapter, we discuss and analyze the results of CASTLE as well as the project itself.
+In this chapter we start by analyzing the results that CASTLE achieved.
+Consequently, the limitations that impeded the project will be mentioned and discussed.
+Finally we will mention in what ways the project can further be explored.
 
@@
 \section{Limitations}
+\label{limitations}
 
@@
 Unfortunately, a corpus fitting for this task does not exist at this time.
+This meant that in order to fulfill the original scope, we would have to both create and annotate such a corpus.
+Such a task could be considered a master's thesis by itself and certainly was not possible to complete in tandem with creating a model, at least not in the given time frame.
+As the project is only intended to run for 20 weeks, we had to make a decision about how to proceed and we decided that a model producing state-of-the-art results but not in customer service would be more rewarding, for both us and Telavox, than a specialized customer service model providing average results.
 
-%crossroads i att skapa ett corpus som passar. gick inte för tidsbrist
+When deciding the scope, we also underestimated just how time consuming this project was going to be.
+The initial thought of both emotion recognition and keyword extraction, even with a specialized corpus available from the start, would be a difficult challenge to complete in the time frame.
+As we feel that more time could have been spent on optimization with potential better results for EmoryNLP, adding a wholly new task in the form of keyword extraction would most likely have produced lacking results for both tasks.
+
+As the laptops we were using did not contain a GPU, running our code locally was not possible and we were instead bound by the rules of Google Colab.
+One of these rules is that no notebook can run for more than 24 hours straight.
+Hyperparameter optimization is a time consuming task that is solved by trial and error, and this restriction meant that we could not run as many trials at a time as we wanted which in turn led to less optimization as there simply was not enough time.
 
@@
 Considering the fact that CASTLE did not produce state-of-the-art results for the EmoryNLP corpus, the most natural continuation for the project is to have CASTLE achieve this goal.
+Our proposal is to simply spend more time with hyperparameter optimization and explore what results that might provide.
+Even though CASTLE managed to achieve state-of-the-results for MELD and IEMOCAP, an extensive hyperparameter optimization could possibly improve the results further.
 
-%For a future usage of CASTLE it would be beneficial to have access to machines with 
-%Provided with sufficient hardware resources CASTLE would be able to use all of its potential.
+%DailyDialog?
 
 Another natural continuation for the project is to expand the scope to what it initially was defined as.
-The first proposed goal of the project was to create a model for analyzing written customer service conversations and using keyword extraction as a complement to said analysis.
+As stated in \nameref{limitations}, the initial goal of the project was to create a model for analyzing written customer service conversations and using keyword extraction as a complement to said analysis.
+CASTLE could very well be used as a framework for completing this type of task.
 
@@
 When searching for appropriate corpora, there was no customer-service themed corpus annotated with emotions available.
-One possibility is to create such a corpus either by using an already existing customer service related corpus and annotating it or by creating a corpus from scratch using recorded customer service conversations as a basis.
+One possibility is to create such a corpus using an already existing unannotated customer service related corpus and labelling it with emotions.
+A couple of methods has previously been mentioned in section \nameref{corpus_excerpts} for how a corpus can be annotated with emotions.
+
+In the same manner that CASTLE built upon the work of COSMIC, other projects can achieve even better results by building upon CASTLE.
 
@@
 
-%Lack of resources, vad man kan göra med mer resources. Man är inte restricted till services som t.ex. Colab som hindrar en att hogga resurser längre än 24h. 
 %Utöka scope till vad det var från början vilket var, emotion recognition i samtal. Både i realtid och transcribed calls. Keyword extraction kan användas som ett supplement. 
@@
 When we started this project, our goal was to find answers to the research questions mentioned in section \ref{research_questions}.
+After completing this project, it is safe to say that creating a model capable of fulfilling our primary aim was possible, as CASTLE is just that.
+CASTLE also proves that an LSTM architecture can in fact outperform a GRU one.
+When comparing CASTLE to the previous state-of-the-art models we show, as presented earlier, that CASTLE surpasses those models.
+Our secondary aim and last research question unfortunately has to remain unanswered due to time limitations.
+All in all, we consider CASTLE a success as it produces state-of-the-art results for two corpora and has the potential of achieving it for the remaining two as well, thus fulfilling our primary aim.
 
-%Ett tydligt exempelt på när teori inte överensstämmer med verkligheten. 
 %Able to produce
-%Ta med lägga till som residual i future work.
 

### 2021-01-27 msccls/MScTemplateCS.tex
report done!

@@
 %\student{Flavius Gruian}{Flavius.Gruian@cs.lth.se}
-\students{Nick Persson}{nick@persson.dev}{Filip Karabeleski}{fkarabeleski@gmail.com}
+\students{Filip Karabeleski}{fkarabeleski@gmail.com}{Nick Persson}{nick@persson.dev}
 
@@
 
-We would like express our gratitude, first and foremost, to our supervisor Pierre. There is no way that this project would have been completed without your guidance, at least not in time.
+We would like to express our gratitude, first and foremost, to our supervisor Pierre. There is no way that this project would have been completed without your guidance, at least not in time.
 
@@
 
-As the COVID-19 epidemic created some unusual circumstances, we would like to give a special thanks to \textit{Åke P Fågel \& Vilt} for providing us with a stable office to work out of. 
+As the COVID-19 pandemic created some unusual circumstances, we would like to give a special thanks to \textit{Åke P Fågel \& Vilt} for providing us with an office to use. 
 }
@@
 %Leave a \textit{2cm} vertical space after the abstract and provide a few keywords relevant for your report. Use five to six words, of which at most two should be from the title.
-}
-
-\keywords{Machine Learning, NLP, Emotion Recognition}
+
+Conversational emotion recognition is a notoriously challenging task for computers as conversations often contain subtle nuances that are difficult to identify.
+In this work, we investigated the recognition of conversational emotions in a set of dialogue corpora.
+We started from an existing baseline and, using empirical observation; we combined it with a previously untested set of architectures.
+We created a new framework that builds upon the previous state of the art and complements it with more effective recurrent devices.
+As a result, using this framework, CASTLE, we could improve the F1-score of previously published results on two of the three corpora we evaluated.}
+% produces state-of the art results on several corpora commonly used as benchmarks for these types of tasks.
+% possibility of improving the results in the field of
+
+\keywords{NLP, emotion recognition, recurrent networks, deep learning}
 
@@
 \section{Background}
+
 Natural language processing (NLP) is a field of study mainly in computer science that deals with analyzing human language in a way that computers can understand and process.
 According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notation.
-Since natural languages naturally evolve as time goes on and their adherence to structure diminish, it is difficult to define explicit rules to follow.
+
+Concretely, NLP fields of application include machine translation, information extraction, and text categorization.
+For the last field, text categorization consists of automatically assigning a category to a text, for instance, to decide if a message is spam or not.
+
+Since natural languages naturally evolve as time goes on and their adherence to structure diminishes, it is difficult to define explicit rules to follow.
 For this reason, there is a need for machine learning models to aid the computer in identifying these diffuse rules.
 
-In order to feed the model with effective information, the input is structured and inserted into a data set.
-This way, the computer has an easier way of processing and analyzing the data.
-When handling data sets constructed using exclusively textual input, the data sets are instead referred to as \textit{corpora}.
+To build a machine-learning model, we need structured information as input in the form of a dataset.
+This way, the model will be able to learn parameters that are used to analyze and automatically categorize the data.
+Datasets using exclusively textual input are usually referred to as \textit{corpora}.
 
@@
 \section{Purpose}
-Customer service is centered around the relation between company and customer with the goal of pleasing the customer.
-This goal can be very difficult to achieve and there is no straightforward way to do so.
-By analyzing prior conversations, it is possible for companies to gain valuable insights in the most effective way to please the customer.
-Customer service is however a massive industry and it would take countless hours to analyze the conversations from even a single company manually.
+Customer service is centered around the relationship between company and customer with the goal of pleasing the customer.
+This goal can be challenging to achieve, and there is no straightforward way to do so.
+By analyzing prior conversations, companies can gain valuable insights in the most effective way to please the customer.
+However, customer service is a massive industry, and it would take countless hours to analyze the conversations from even a single company manually.
 
 The company sponsoring this project approached us with an idea to try and solve this problem using machine learning.
-A usual technique for these types of problems is \textit{sentiment analysis}, where utterances are labeled either as positive or negative.
+A usual technique for these problems is \textit{sentiment analysis}, where utterances are labeled either positive or negative.
 We instead decided to take another route, namely \textit{emotion recognition}.
-Emotion recognition focuses on labelling utterances from a non-binary set of emotions instead of sentiment analysis' binary method.
-This allows an analysis with greater specificity.
-
-In order to further the academic relevance of this project we decided that we would attempt to contribute to the field of emotion recognition by achieving state-of-the-art results.
-
-\section{State-of-the-art}
-In this paper, the term state-of-the-art will be recurringly mentioned as it is a fundamental part of our aim and result.
+Emotion recognition focuses on labeling utterances from a non-binary set of emotions instead of sentiment analysis' binary method.
+This allows analysis with greater specificity.
+
+To further the academic relevance of this project, we decided that we would attempt to contribute to the field of emotion recognition by achieving state-of-the-art results.
+
+\section{State of the art}
+%The progress of emotion recognition has been accomplished by 
+%The previous work carried out in the field of emotion recognition has
+
+In this report, the term \emph{state of the art} will be recurringly mentioned as it is a fundamental part of our aim and result.
 The term refers to the highest level of achievement accomplished.
-A common goal in machine learning projects is to attempt to raise the bar for the current state-of-the-art results across various metrics for the task that the project is trying to solve.
+A common goal in machine learning projects is to attempt to raise the bar for the current
+results across various metrics for the task that the project is trying to solve.
 Presenting state-of-the-art results is an excellent approach in proving that an implementation or algorithm is valid and valuable.
+
+%Previous work
 
@@
 Telavox is a Swedish telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service.
-As all of their customers deal in customer service, they would all benefit from the solution previously discussed.
-Telavox therefore has a great interest in being able to offer this as a service.
-As Telavox has over 250 businesses as customers, the solution would facilitate the quality of customer service world wide.
+As all of their customers deal in customer service, they would benefit from the previously discussed solution.
+Telavox, therefore, has a great interest in being able to offer this as a service.
+As Telavox has over 250 businesses as customers, the solution would facilitate the customer service quality worldwide.
 
@@
 \label{research_questions}
-The primary aim of this thesis is to create a emotion recognition model that is able to produce state-of-the-art results.
-After this, a secondary aim is to investigate whether or not it is possible to create a dynamic model for instant response regarding customer service related conversations.
+This thesis' primary aim is to create an emotion recognition model that can produce state-of-the-art results.
+After this, a secondary aim is to investigate whether it is possible to create a dynamic model to provide interactive support in customer service-related conversations.
 The following questions are used as guidelines on how to realize these aims.
@@
 %?? THe first question is can we create such a model??
-    \item \textit{Could a different architecture than the current state-of-the-art produce better results?}
+    \item \textit{Can we build an architecture that improves the state of the art?}
 %?? here your statement is too specific as you did not introduce the terms. Use vague words such as What are the different option to design an architecture and train a model?? 
 %?? I have swapped the items ??
-    \item \textit{How does our model compare to other models in the same field evaluating on the same metrics?}
-    \item \textit{After implementing a fully functioning model, how can we adapt it to predict utterances in real-time?}
+    \item \textit{If yes, how does our model compare to other models in the same field evaluating on the same metrics?}
+    \item \textit{After implementing a fully functioning model, how can we adapt it to predict utterances in real-time and in the context of an interactive service?}
     %Snacka om detta i limitations
@@
 
-In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model:
+In order to solve most emotion recognition problems, two main concepts are fundamental to build an accurate model:
 \begin{enumerate}
-\item As shown by the recent progress in the field \citep{deep_survey}, the model should be created with deep learning algorithms.
-This should enable it to find patterns that it otherwise could not be able to.
-\item To effectively process the utterances, deep learning requires to collect an annotated corpus and design an appropriate representation for the words.
+\item As shown by the recent progress in the field \citep{deep_survey}, the model should be created with deep learning algorithms. This should enable it to find patterns that it otherwise could not. 
+\item To effectively process the utterances, deep learning requires collecting an annotated corpus and design an appropriate representation for the words.
 \end{enumerate}
@@
 \section{Machine Learning}
-Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program the system.
+Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem rather than to program the system explicitly.
 By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem.
-In \textit{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number.
-The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector $\mathbf{x}$ as input and outputs information that allows deducing the label $y$ for this feature vector.)
+In \textit{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow, and $y_i$ is a label, which is either a finite set of classes or a real number.
+Supervised learning aims to use the dataset to produce a model that takes a feature vector $\mathbf{x}$ as input and outputs information that allows deducing the label $y$ for this feature vector.)
 
 The other main type of task in machine learning is \textit{unsupervised learning}.
-The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector.
+The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$, where $\mathbf{x}_i$ is a feature vector.
 The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector $\mathbf{x}$ as input and outputs either another vector or a value.
@@
 Neural networks are machine learning systems loosely inspired by biological neural networks.
-A neural network is a collection of connected nodes or neurons that was first mentioned in \citet{mcculloch1943}.
-A neuron \citep{dawson1998ann} is an information-processing unit that receives multiple inputs and produces a single output that can be sent to multiple other neurons.
-To find the output of a neuron the weighted sum of all inputs is weighted by the weights all connecting to the neuron.
+A neural network is a collection of connected nodes or neurons first mentioned in \citet{mcculloch1943}.
+A neuron \citep{dawson1998ann} is an information-processing unit that receives multiple inputs and produces a single output sent to multiple other neurons.
+To find the output of a neuron, all inputs are weighted and then summed.
 This weighted sum is then passed through an activation function to provide the output.
@@
     \includegraphics[width=12cm]{msccls/explanatory_images/neuron_new.png}
-    \caption{An example of a single neuron. From \citet{wiki:neuron_image}.}
+    \caption{Example of a single neuron. From \citet{wiki:neuron_image}.}
     \label{fig:neuron}
@@
 
-Neural networks can in turn be structured together and form multiple layers that progressively learn.
+Neural networks can, in turn, be structured together and form multiple layers that progressively learn.
 These networks fall under the deep learning umbrella.
@@
 
-Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi-variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.
+Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi-variable generalization of the derivative, which is then used to adjust the weights so that the distance score produced by the loss function is minimized.
 
@@
 Since the inputs are sequential, all the inputs are dependent on each other in one way or another.
-RNNs are therefore suitable for processing sequences of text, audio, and video where every computation will be performed on each element in the sequence.
+RNNs are suitable for processing sequences of text, audio, and video, where every computation will be performed on each element in the sequence.
 
 One of RNN's weaknesses is that they suffer from the \textit{vanishing gradient problem} \citep{hochreiter1998}.
-While RNN's theoretically can process arbitrarily long sequences, in practice, they are limited to the length of the sequence.
+While RNN's theoretically can process arbitrarily long sequences, they are, in practice, limited to the length of the sequence.
 If the sequence is too long, the network will struggle to try to look back in time longer than a few steps away.
-In a lot of cases, this will result in leaving out important information from the prediction.
+In many of cases, this will result in leaving out important information from the prediction.
 There are instead other alternatives that consider this weakness.
@@
     \includegraphics[scale=0.35]{msccls/explanatory_images/rnn.png}
-    \caption{A simple RNN node}
+    \caption{Illustration of a simple RNN cell}
     \label{fig:rnn_node}
@@
 
-As illustrated in Figure \ref{fig:lstm_node}, a LSTM node consists of an \textit{input gate}, an \textit{output gate}, a \textit{forget gate} and a \textit{cell state}. 
+As illustrated in Figure \ref{fig:lstm_node}, an LSTM node consists of an \textit{input gate}, an \textit{output gate}, a \textit{forget gate}, and a \textit{cell state}. 
 \begin{itemize}
@@
 Basic GRU nodes  are primarily constructed of an \textit{update gate} and a \textit{reset gate}.
-The update gate regulates the amount of information that is passed along to the next state, while the reset gate, much like the forget gate of LSTM, decides the amount of past information that is going to be disregarded henceforth.
+The update gate regulates the amount of information passed along to the next state, while the reset gate, much like the forget gate of LSTM, decides the amount of past information that is going to be disregarded henceforth.
 
 Although fairly new, it is already a well-established variant of recurrent neural networks.
-In comparison to \textit{LSTM} layers, GRU layers are cheaper to run since they use fewer parameters, but have less representational power.
+Compared to \textit{LSTM} layers, GRU layers are cheaper to run since they use fewer parameters but have less representational power.
 It has been shown by \citet{Gruber2020AreGC} that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers.
@@
     \includegraphics[scale=0.58]{msccls/explanatory_images/lstm.png}
-    \caption{A LSTM node} 
+    \caption{Illustration of an LSTM cell} 
     %?? You need to mention the source of your figure if it is not yours??
@@
     \includegraphics[scale=0.58]{msccls/explanatory_images/gru.png}
-    \caption{A GRU node}
+    \caption{Illustration of a GRU cell}
     \label{fig:gru_node}
@@
 \section{Training Process}
-In order to train a model, there must exist some data set, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method, that feeds the model with input.
-When using a labeled data set, the labels are regarded as the solution to the problem.
+To train a model, there must exist some dataset, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method that feeds the model with input.
+When using a labeled dataset, the labels are regarded as the solution to the problem.
 When training a model, the goal is to find values for its weights such that the model correctly labels the given input.
 
-The training of a model is done step by step in the form of \textit{epochs}.
-Epochs represent the number of times that the model iterates through a data set that is intended to train the model.
+The training of a model is done step by step in the form of \textit{epochs}. Epochs represent the number of times 
+the model iterates through a dataset intended to train the model.
 Following each epoch, a \textit{loss function} measures how incorrect the model's previous prediction was and produces a value for the \textit{loss}.
@@
 The \textit{backpropagation} algorithm is a fundamental tool for minimizing the loss.
-The algorithm starts with the loss and works its way \textit{backwards} through the layers of the network and applies the chain rule to determine how much impact each of the weights has had on the loss.
-This impact is then used for the \textit{gradient descent} algorithm to alter the weights in the direction of flattening out the derivative.
-While travelling through the function, the ultimate goal is to find where the derivative is zero and this is done for each weight in the network.
+The algorithm starts with the loss and works its way \textit{backward} through the layers of the network, and applies the chain rule to determine how much impact each of the weights has had on the loss.
+This impact is then used for the \textit{gradient descent} algorithm to alter the weights to flatten out the derivative.
+While traveling through the function, the ultimate goal is to find where the derivative is zero, and this is done for each weight in the network.
 Gradient descent updates the weights to match the results from the backpropagation algorithm \citep{franoischollet2017learning}.
@@
 
-\section{Evaluation}
-Many different metrics can be used to evaluate a machine learning model, based on what kind of model and what its purpose is.
-For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}.
-
-In a confusion matrix, one of the axes represents the actual class of the instance the model is trying to predict and the other represents what instance the model predicted.
-The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}.
-In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}
-
-\begin{figure}[h!]
-    \centering
-    \hbox{\hspace{6em}\includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}}
-    \caption{Binary Confusion Matrix}
-    \label{fig:confusion}
-\end{figure}
-
-\begin{itemize}
-    \item TP means that the model was correct in predicting positive
-    \item TN means that the model was correct in predicting negative
-    \item FP means that the model was incorrect in predicting positive
-    \item FN means that the model was incorrect in predicting negative
-\end{itemize}
-
-Whenever our model predicts an instance during training, we increment the corresponding cell in the matrix, letting us see and draw conclusions based on how our model performed.
-
-For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative.
-
-The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class.
-For instance, when determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{neutral}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.
-
-%kan snacka om Type I och Type II errors om mer text behövs.
-
-\subsection{Accuracy}
-Accuracy quantifies how many percent of predictions the model got right and is calculated by taking all instances, where our model predicted positive and dividing it with all instances:
-
-$$ Accuracy = \frac{TP + TN}{TP + TN + FP + FN}$$
-
-If we have a balanced dataset with a similar amount of positive and negative instances, then this can give valuable information on the performance of the model since getting predictions right is the goal of any model.
-If, however, our set is imbalanced and contains e.g.
-90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless \citep{imbalancedlearning}.
-
-\subsection{Precision}
-Precision is a metric that quantifies how many of the instances predicted positive that were true.
-It's calculated by taking all of the instances where the model predicted positive and was correct and dividing it by all the instances where it predicted positive.
-
-$$ Precision = \frac{TP}{TP + FP}.$$
-
-Precision gives us information on how correct the model is when predicting an instance to be true.
-This is useful when dealing with instances, where it's important to not give a false positive.
-If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for an innocent person to be determined as a criminal.
-
-However, precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, that said person is likely guilty.
-
-\subsection{Recall}
-Recall is a metric that quantifies how many percent of positive instances that our model managed to predict which we calculate by dividing all instances the model predicted positive and was right by all positive instances.
-
-$$ Recall =\frac{TP}{TP + FN}.$$
-
-As a sort of counterpoint to precision which tells us how good the model is when guessing true, recall tells us how many positive instances the model missed.
-A high recall can be extremely important when dealing with subjects such as contagious diseases where a result of false-negative can be devastating.
-
-Recall does not take into account how many positive predictions turned out to be false so in cases, where that does not matter and you want to predict positive at the slightest probability of it being true, optimizing for a high recall is essential.
-
-\subsection{F1 score}
-A combination of the two previous metrics, the F1 score, is calculated by creating the harmonic mean of precision and recall:
-
-$$ F1score = 2 \cdot \frac{Precision \cdot Recall}{Precision + Recall}.$$
-
-In cases where there is no extra cost associated with either false negatives or false positives, the goal is usually to minimize the total amount of errors.
-The harmonic mean is a type of average that is appropriate when working with ratios, which is what both precision and recall are.
-
-This lets us improve the model concerning either of the variables and not have to think about if, for example, the reduction in Recall was worth the gain in Precision.
-In multi-class scenarios with imbalanced data sets there are multiple ways to calculate the F1-score, namely \textit{macro}, \textit{weighted} and \textit{micro}.
-Which one to use depends on your data set and what you prioritize in your model.
-
-The macro F1-score is derived by calculating the F1-score for each class separately and then just summing them up, leading to an excessive focus on the classes with fewer instances.
-If we assume that we are using a data set where the \textit{neutral} class is represented in 10 000 instances and a less common emotion \textit{fear} in just 50.
-The model then managed to predict 9 995 of the neutral instances, but only 25 of the fear ones.
-When calculating the mean F1-score, the low fear score would muddy the entire score and make it seem bad, when in fact the score for the neutral case is excellent.
-If we want to optimize for the minority classes, however, this lets us do just that.
-
-To counteract this behavior of favoring the minority class, one could instead calculate the weighted F1-score.
-After calculating the F1-score for each class, they are weighted by the number of true instances for each class, thus making them proportionate before calculating the final score.
-This gives a score more reflecting on how the model performs, but can also favor identifying the majority class heavily, which might not be the goal of the model.
-
-When calculating micro F1 score, instead of calculating the F1 scores separately, the F1 score is calculated directly by using the total values for TP, FN, and FP in the Precision and Recall formulas which places no bias on either class.
-If we have an example data set with two minority classes \textit{anger} and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class.
-From the perspective of the joy class, however, it would instead be an FP.
-Since micro calculations take into account the total values for our identification labels and each FN correspond to the FP of another class, $FP = FN$, which in turn means that $Precision = Recall$ in these cases.
-
-Precision and Recall being equal means that their mean also is equal and thus \textit{F1-}$score_{micro} = Precision_{Micro} = Recall_{Micro}$ which incidentally also equals \textit{Accuracy} as this is the proportion of instances we got correct.
-This means that just like Accuracy, this is a good metric on balanced data sets but can be deceiving in an imbalanced one.
-
 \section{Natural Language Processing -- NLP}
-Natural languages are the set of languages that humans have evolved naturally without any predetermined strategy such as English or Mandarin.
-Making computers understand and analyze natural languages has long been a challenge and one of the first practical attempts was Weaver and Booth's translation project in 1946 trying to crack enemy codes during World War II \citep{Liddy2001NLP}.
-Shortly thereafter \citet{Turing1950} created his famous \textit{Imitation game}, or more commonly referred to as the Turing test, which has developed into a test to evaluate whether a computer's natural language response in a real-time written conversation can be indistinguishable from that of a human.
-
-One of the reasons that make natural language processing a challenge is the fact that natural languages are minted with unstructured types of data such as inflection or dialects, or using phrases to convey the opposite than the common meaning namely sarcasm.
+One of the reasons that make natural language processing a challenge is that natural languages are minted with unstructured types of data such as inflection, dialects, or sarcasm.
 Natural languages are also oftentimes ambiguous and require a context-based interpretation that computers have a hard time understanding.
@@
 
-Emotion is a key part of human language and it, therefore, comes naturally that teaching a computer to understand emotion in human language is a big part of NLP.
-Initially, the focus lied in sentiment analysis, a field of study focused on determining the binary polarization of a piece of text as either positive or negative.
+Emotion is a key part of human language, and therefore it comes naturally that teaching a computer to understand emotion in human language is a big part of NLP.
+Initially, the focus laid on sentiment analysis, a field of study focused on determining the binary polarization of a piece of text as either positive or negative.
 %Godtycklig bit om sentiment analysis. 
-Emotion recognition is an extension of sentiment analysis that instead of just determining the polarization of text, attributes emotion to it.
+Emotion recognition is an extension of sentiment analysis that, instead of just determining text's polarization, attributes emotion to it.
 This process is more challenging than sentiment analysis, as the differences between emotions are subtler than that of just positive and negative.
@@
 
-Instead of working directly with raw input, it first need to be converted into a numerical representation in order for the model to process it \citep{franoischollet2017learning}.
-In the case of textual input this is done by transforming it into a \textit{word vector}.
+Instead of working directly with raw input, it first needs to be converted into a numerical representation in order for the model to process it \citep{franoischollet2017learning}.
+In the case of textual input, this is done by transforming it into a \textit{word vector}.
 
@@
 The most basic way of creating a word vector is the one-hot encoding method.
-This creates a very big and sparse $n \cdot n$ matrix where $n$ is the number of words in the embedding.
+This creates a massive and sparse $n \cdot n$ matrix where $n$ is the number of words in the embedding.
 Each word is then represented by an array filled with zeroes except for one index set to one.
@@
 
-As mentioned, this method is very basic and creates an unnecessarily large matrix that is largely filled up with nothing.
+As mentioned, this method is elementary and creates an unnecessarily large matrix that is largely filled up with nothing.
 If we wanted to represent a vocabulary of 20 000 words, each word would be represented by 19 999 zeroes and a single one.
@@
 The other big issue with it is that no meaning can be extracted from the matrix.
-If our vocabulary contained the words ``cat'', ``dog'' and ``hamster'', we could easily see that these are all domestic animals but in the space of the one-hot vector, ``cat'' could be as close to ``dog'' as it could be to ``airplane''.
+If our vocabulary contained the words ``cat'', ``dog'', and ``hamster'', we could easily see that these are all domestic animals, but in the space of the one-hot vector, ``cat'' could be as close to ``dog'' as it could be to ``airplane''.
 
 \subsection{Word Embedding}
-To resolve the two aforementioned problems with one-hot encodings, word embeddings were created, also known as dense word vectors \citep{neuralnetworkmethods}. 
-Here, instead of creating a sparse, hardcoded vector with equal distance between words, they are learned from data and packed into a lower-dimensional vector where each dimension contains some relevant data about the word.
-According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional, or 1024-dimensional are common when dealing with very large vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use a one-hot encoding.
+Word embeddings were created to resolve the two aforementioned problems with one-hot encodings, also known as dense word vectors \citep{neuralnetworkmethods}. 
+Instead of creating a sparse, hardcoded vector with equal distance between words, they are learned from data and packed into a lower-dimensional vector where each dimension contains some relevant data about the word.
+According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional, or 1024-dimensional are common when dealing with extensive vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use one-hot encoding.
 
@@
 
-In Figure \ref{fig:direction}, the four words \textit{cat}, \textit{dog}, \textit{wolf} and \textit{tiger} are represented on a geometric plane.
+Figure \ref{fig:direction} shows the four words \textit{cat}, \textit{dog}, \textit{wolf}, and \textit{tiger} are represented on a geometric plane.
 We can clearly see some of the possible geometric transformations possible by using direction as a tool.
@@
 
-Another vector will take us both from \textit{cat} to \textit{dog} and from \textit{tiger} to \textit{wolf} which could be interpreted as ``converting a feline to a canine''.
-Two meaningful transformation vectors that are common in the real world are ``gender'' and ``plural'' vectors.
-By applying a ``female'' vector to \textit{king}, \textit{queen} is obtained and the same vector returns \textit{actress} if instead used on \textit{actor}.
-``Plural'', as should seem obvious by now, if used on \textit{actor} will return \textit{actors}.
+Another vector will take us both from \textit{cat} to \textit{dog} and from \textit{tiger} to \textit{wolf}, which could be interpreted as ``converting a feline to a canine''.
+Two meaningful transformation vectors common in the real world are ``gender'' and ``plural'' vectors.
+Applying a ``female'' vector to \textit{king}, returns \textit{queen}, and the same vector returns \textit{actress} if instead used on \textit{actor}.
+``Plural'', following the same logic, if used on \textit{actor} will return \textit{actors}.
 
@@
 
-\subsection{RoBERTa}
+%\subsection{RoBERTa}
 
@@
 
+\section{Evaluation}
+Many different metrics can be used to evaluate a machine learning model, depending on what kind of model and what its purpose is.
+For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}.
+
+In a confusion matrix, one of the axes represents the actual class of the instance the model is trying to predict, and the other represents what instance the model predicted.
+The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}.
+In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}
+
+\begin{figure}[h!]
+    \centering
+    \hbox{\hspace{6em}\includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}}
+    \caption{Binary Confusion Matrix}
+    \label{fig:confusion}
+\end{figure}
+
+\begin{itemize}
+    \item TP means that the model was correct in predicting positive
+    \item TN means that the model was correct in predicting negative
+    \item FP means that the model was incorrect in predicting positive
+    \item FN means that the model was incorrect in predicting negative
+\end{itemize}
+
+Whenever our model predicts an instance during training, we increment the corresponding cell in the matrix, letting us see and draw conclusions based on how our model performed.
+
+For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative.
+
+The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class.
+For instance, when determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{neutral}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.
+
+%kan snacka om Type I och Type II errors om mer text behövs.
+
+\subsection{Accuracy}
+Accuracy quantifies how many percent of predictions the model got right and is calculated by taking all instances, where our model predicted positive and dividing it with all instances:
+
+$$ Accuracy = \frac{TP + TN}{TP + TN + FP + FN}$$
+
+If we have a balanced dataset with a similar amount of positive and negative instances, this can give valuable information on the model's performance since getting predictions right is the goal of any model.
+If, however, our set is imbalanced and contains, e.g., 90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance, so the accuracy suddenly becomes useless \citep{imbalancedlearning}.
+
+\subsection{Precision}
+Precision is a metric that quantifies how many of the instances predicted positive that were true.
+It is calculated by taking all of the instances where the model predicted positive and was correct and dividing it by all the instances where it predicted positive.
+
+$$ Precision = \frac{TP}{TP + FP}$$
+
+Precision gives us information on how correct the model is when predicting an instance to be true.
+This is useful when dealing with instances where it is important not to give a false positive.
+If we have a model trying to predict whether or not a suspect is guilty of a certain crime, a high precision will make it unlikely for an innocent person to be determined as a criminal.
+
+However, precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, that said person is likely guilty.
+
+\subsection{Recall}
+Recall is a metric that quantifies how many percent of positive instances that our model managed to predict, which we calculate by dividing all instances the model predicted positive and was right by all positive instances.
+
+$$ Recall =\frac{TP}{TP + FN}$$
+
+As a counterpoint to precision, which tells us how good the model is when guessing true, recall tells us how many positive instances the model missed.
+A high recall can be critical when dealing with e.g.
+contagious diseases where a result of false-negative can be devastating.
+
+Recall does not consider how many positive predictions turned out to be false, so in cases where that does not matter, and you want to predict positive at the slightest probability of it being true, optimizing for a high recall is essential.
+
+\subsection{F1 score}
+A combination of the two previous metrics, the F1 score, is calculated by creating the harmonic mean of precision and recall:
+
+$$ F1score = 2 \cdot \frac{Precision \cdot Recall}{Precision + Recall}$$
+
+In cases where there is no extra cost associated with either false negatives or false positives, the goal is usually to minimize the total amount of errors.
+The harmonic mean is a type of average appropriate when working with ratios, which is what both precision and recall are.
+
+This lets us improve the model concerning either of the variables and not have to think about if, for example, the reduction in Recall was worth the gain in Precision.
+There are multiple ways to calculate the F1-score in multi-class scenarios with imbalanced datasets, namely \textit{macro}, \textit{weighted} and \textit{micro}.
+Which one to use depends on your dataset and what you prioritize in your model.
+
+The macro F1-score is derived by calculating the F1-score for each class separately and then summing them up, leading to an excessive focus on the classes with fewer instances.
+If we assume that we are using a dataset where the \textit{neutral} class is represented in 10 000 instances and a less common emotion \textit{fear} in just 50.
+The model then managed to predict 9 995 of the neutral instances, but only 25 of the fear ones.
+When calculating the mean F1-score, the low fear score would muddy the entire score and make it appear worse, when in fact, the score for the neutral case is excellent.
+If we want to optimize for the minority classes, however, this lets us do just that.
+
+To counteract this behavior of favoring the minority class, one could instead calculate the weighted F1-score.
+After calculating the F1-score for each class, they are weighted by the number of true instances for each class, making them proportionate before calculating the final score.
+This gives a score more reflecting on how the model performs, but can also favor identifying the majority class heavily, which might not be the model's goal.
+
+When calculating the micro F1 score, instead of calculating the F1 scores separately, the F1 score is calculated directly using the total values for TP, FN, and FP in the Precision and Recall formulas places no bias on either class.
+If we have an example dataset with two minority classes, \textit{anger}, and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class.
+From the perspective of the joy class, however, it would instead be an FP.
+Since micro calculations consider the total values for our identification labels and each FN correspond to the FP of another class, $FP = FN$, which means that $Precision = Recall$ in these cases.
+
+Precision and Recall being equal means that their mean also is equal, and thus \textit{F1-}$score_{micro} = Precision_{Micro} = Recall_{Micro}$, which incidentally also equals \textit{Accuracy} as this is the proportion of instances we got correct.
+This means that just like Accuracy, this is a good metric on balanced datasets but can be deceiving in an imbalanced one.
+
 \chapter{Implementation}
 %Summary of what the chapter is about
-This chapter is mainly focused on giving a clear description of what we have done for the duration of the project.
+This chapter is mainly focused on giving a clear description of what we have done for the project's duration.
 Firstly we will describe the model that we used as a baseline, \textit{COSMIC}.
-Following, we will cover our model \textit{CASTLE}, how it has been implemented and what changes were made from COSMIC.
+We will cover our model \textit{CASTLE}, its implementation, and what changes were made from COSMIC.
 Lastly, we will describe our setup and what corpora we used.
@@
 \section{Baseline -- COSMIC}
-When trying to improve existing results in many machine learning fields, instead of starting from scratch, a baseline is used.
+When trying to improve existing results in many machine learning fields, a baseline is used instead of starting from scratch.
 A baseline is an already functioning model that provides basic results to the problem at hand.
@@
 COSMIC is a GRU-based model implemented with PyTorch that has achieved state-of-the-art results across four corpora commonly used in emotion recognition.
-The idea behind the model is when classifying an utterance, the emotions of the utterances both preceding and succeeding it are also taken into account.
-This way the context of the conversation can be used to more accurately predict the emotions of individual utterances.
+The idea behind the model is that when classifying an utterance, the emotions of the utterances, both preceding and succeeding, are also taken into account.
+This way, the context of the conversation can be used to predict the emotions of individual utterances more accurately.
 
@@
 Figure \ref{fig:COSMIC_arch} provides a simplified description of how COSMIC operates in a conversation between person A and person B.
-For each time step $t$ there is an utterance said by either A or B with the other person as a listener.
-At the next time step $t+1$ the roles are alternated.
-For each utterance, there exist multiple GRU-cells that predict both the intent of the speaker and the response of the listener at a certain time step.
-These cells are used not only to predict the emotion of the current utterance, but also used for the prediction of emotions in other utterances related to said utterance.
+For each time step $t$, there is an utterance said by either A or B with the other person as a listener.
+At the next time-step $t+1$, the roles are alternated.
+For each utterance, multiple GRU-cells exist that predict both the speaker's intent and listener's response at a certain time-step.
+These cells are used to predict the emotion of the current utterance and the prediction of emotions in other utterances related to said utterance.
 
@@
     \item \textit{Internal state} tracks for each participant what emotions they are feeling but not necessarily showing. Affects the Intent state and the Context state.
-    \item \textit{External state} portrays for each participant what emotions they are expressing to the other participants through both utterances and reactions. Affects the Context state.
+    \item \textit{External state} portrays, for each participant, what emotions they are expressing to the other participants through both utterances and reactions. Affects the Context state.
     \item \textit{Intent state} is a state only present for the current speaker.
@@
 
-The final state, Emotion state, takes into account the current utterance, the summed emotional state through the Internal and External state, as well as the Intent state to ultimately make a prediction for the current utterance.
+\emph{Emotion state}, the final state, considers the current utterance, the summed emotional state through the Internal and External state, and the Intent state to predict the current utterance.
 This process is repeated for every utterance in every conversation in the corpus.
@@
 Building on COSMIC, we designed CASTLE, COSMIC-based AnalySer Taking advantage of LSTM for Emotion recognition, to answer our research questions.
-We have made several improvements on the COSMIC baseline, one of the fundamental being replacing the internal GRU architecture with an LSTM-based architecture as the name of the model indicates.
+We have made several improvements on the COSMIC baseline, one of the fundamental being replacing the internal GRU architecture with an LSTM-based architecture, as the name of the model indicates.
 
 At the start of the project, we read some recently published research in the field of deep learning, mainly the article by \citet{edseee.922172720200601} that states that GRU is not as suited for small pieces of text and large corpora as LSTM.
-This is where we got the inspiration to test whether or not the architectural change to LSTM could improve the results.
-
-COSMIC takes advantage of the \texttt{cudnn} framework in PyTorch to assist in multiple ways such as the \texttt{backends.cudnn.benchmark} which attempts to automatically optimize the algorithm during runtime.
-As we aimed to obtain as much control as possible over how the model behaved, we excluded this and other third party features that restricted our domain of control.
-
-COSMIC is also implemented with a \textit{residual} connection \citep{he2015deep} which allows the model to redefine the layers as learning residual functions referencing the layer inputs.
+This is where we got the inspiration to test whether the architectural change to LSTM could improve the results.
+
+COSMIC takes advantage of the \texttt{cudnn} framework in PyTorch to assist in multiple ways, such as the \texttt{backends.cudnn.benchmark}, which attempts to optimize the algorithm during runtime automatically.
+As we aimed to obtain as much control as possible over how the model behaved, we excluded this and other third-party features that restricted our domain of control.
+
+COSMIC is also implemented with a \textit{residual} connection \citep{he2015deep}, which allows the model to redefine the layers as learning residual functions referencing the layer inputs.
 This is supposed to facilitate the training of the model.
-While we originally wanted to implement this feature in CASTLE, there was simply not space to do it within the scope of the project.
+While we originally wanted to implement this feature in CASTLE, there was simply no space to do it within the project's scope.
 
@@
 \subsection{Hyperparameters}
-CASTLE's performance depends on three different hyperparameters. The ones we chose are \textit{dropout}, \textit{batch size}, and \textit{learning rate}. \citep{hyperparameters}:
+CASTLE's performance depends on three different hyperparameters. The ones we chose are \textit{dropout}, \textit{batch size}, and \textit{learning rate} \citep{hyperparameters}:
 \begin{enumerate}
-    \item Dropout is a value between 0 and 1 that regulates what percentage of learned data that the model randomly chooses to forget.
-This is used in order to combat overfitting and make sure the model is actually predicting instead of just learning the structure of the corpus.
+    \item Dropout is a value between 0 and 1 that regulates what percentage of learned data the model randomly chooses to forget.
+This is used to combat overfitting and make sure the model is actually predicting instead of just learning the corpus structure.
     \item  Batch size governs how many training examples will be utilized in each iteration. A larger value increases the accuracy of the gradient in trade off for a higher memory cost.
-    \item  Learning rate defines the speed of which the model updates its parameters.
-A high learning rate speeds up the learning at the risk of not converging while a low value will more likely converge, sacrificing speed in the process.
+    \item  Learning rate defines the speed at which the model updates its parameters.
+A high learning rate speeds up the learning at the risk of not converging, while a low value will more likely converge, sacrificing speed in the process.
 \end{enumerate}
@@
 To optimize our hyperparameters, we implemented the \texttt{Tune} framework offered by \texttt{Ray}.
-This framework allowed us to efficiently experiment with many different configurations and provided a simple way to run trials concurrently.
-Most of our optimization were done with a combination of grid search and random search.
-Random search is a function with two parameters \textit{start} and \textit{end}, and for each trial it randomly selects a float from in between the two values to use for the given hyperparameter.
+This framework allowed us to experiment efficiently with many different configurations and provided a simple way to run trials concurrently.
+Most of our optimization was done with a combination of grid search and random search.
+Random search is a function with two parameters, \textit{start} and \textit{end}, and for each trial, it randomly selects a float from in between the two values to use for the given hyperparameter.
 Grid search instead takes in a list of floats as a parameter and runs a trial for each value in the list.
-Multiple grid searches can very easily take a large amount of time to run as all possible combination of values across all grid searches will be tested in separate trials.
-
-Our approach was to start out with random search on parameters to get a sense on what range of values produced optimal results.
+Multiple grid searches can easily take a large amount of time to run as all possible combinations of values across all grid searches will be tested in separate trials.
+
+Our approach was to begin with random search on parameters to get a sense of what range of values produced optimal results.
 Once we had established the magnitude of which we wanted our parameters to be, we tested out various combinations with grid search, keeping the ones that produced promising results and replacing the others with new trials.
@@
 When deciding what configuration to use for our experiments, we based our decisions on the empirical studies made by  \citet{bouthillier:hal-02447823}.
-In their report, they surveyed authors whose papers were presented at the peer-reviewed conferences NeurIPS 2019 and ICLR 2020.
+Their report surveyed authors whose papers were presented at the peer-reviewed conferences NeurIPS 2019 and ICLR 2020.
 As the work is based on answers from relevant authors in the machine learning community, we found it appropriate to use the answers as guidelines for our setup.
@@
 On average, we performed 20 sets of different values for the three hyperparameters across all corpora.
-CASTLE was intended to be evaluated on four different corpora but ultimately only provided results for three of them.
+CASTLE was intended to be evaluated on four different corpora, but it only provided results for three of them due to hardware limitations.
 Our results are based upon 10 consecutive runs with the same setup.
 
-Comparing this to the responses observed in \ref{fig:HAL}, we are in line with the majority in all but one instance, (f), where the majority base their results on 5 or less samples and we instead chose to base it on 10 samples as a higher sample size tends to become a more reliable result.
+Comparing this to the responses observed in \ref{fig:HAL}, we are in line with the majority in all but one instance, (f), where the majority base their results on 5 or fewer samples, and we instead chose to base it on 10 samples as a higher sample size tends to become a more reliable result.
 
 \subsection{Equipment}
-As for our hardware setup, we were not able to run and train the model on our laptops as they lack a GPU.
-To execute our code, we instead utilized Google Colab, a cloud based service where users can remotely run their code on Google servers that are built with more powerful hardware and can handle high demand programs such as machine learning model.
+As for our hardware setup, we could not run and train the model on our laptops as they lack a GPU.
+To execute our code, we instead utilized Google Colab, a cloud-based service where users can remotely run their code on Google servers built with more powerful hardware and can handle high-demand programs such as machine learning models.
 Using a remote service proved to be perfectly fine for three of the four corpora we aimed to run our model on.
-The fourth one, \textit{DailyDialog}, unfortunately required more memory than we were allocated and could not complete run.
+The fourth one, \textit{DailyDialog}, unfortunately, required more memory than we were allocated and could not complete the run.
 
@@
 
-To evaluate our model, we used four popular corpora that are frequently used in emotion recognition research work.
-We chose these corpora as we wanted our results to be comparable with the current state-of-the-art, namely COSMIC as well as the common practice observed in Figure \ref{fig:HAL}(e).
+To evaluate our model, we used four popular corpora frequently used in emotion recognition research work.
+We chose these corpora as we wanted our results to be comparable with the previous state of the art, namely COSMIC and the common practice observed in Figure \ref{fig:HAL}(e).
 As COSMIC compared its results to several other models using these corpora, the results of CASTLE will be directly comparable with COSMIC and the other models.
@@
 %These datasets are recurring in many of the papers on this subject. -> our results comparable
-%All these data sets include a "neutral" emotion.
-This is when an emotion couldn't fit into any of the other categories within the data set using their corresponding methods for classifying the emotions.
+%All these datasets include a "neutral" emotion.
+This is when an emotion couldn't fit into any of the other categories within the dataset using their corresponding methods for classifying the emotions.
 %Fyra populära dataset inom NLP
@@
 
-    \item[IEMOCAP] \citep{busso2008iemocap} is a multimodal, multispeaker corpus with utterances from acting and improvisation scenarios emphasizing on emotional expressions.
+    \item[IEMOCAP] \citep{busso2008iemocap} is a multimodal, multispeaker corpus with utterances from acting and improvisation scenarios emphasizing emotional expressions.
 The utterances are labeled with the following emotions: \textit{anger}, \textit{happiness}, \textit{excitement}, \textit{sadness}, \textit{frustration}, \textit{fear}, \textit{surprise}, \textit{other}, and \textit{neutral state}.
@@
 In comparison to MELD, EmoryNLP has over 60,000 utterances.
-The utterances are labeled with the following emotions: \textit{neutral}, \textit{joyful}, \textit{peaceful}, \textit{powerful}, \textit{scared}, \textit{mad} and \textit{sad}.
+The utterances are labeled with the following emotions: \textit{neutral}, \textit{joyful}, \textit{peaceful}, \textit{powerful}, \textit{scared}, \textit{mad}, and \textit{sad}.
 
@@
 The corpus has an average of circa 8 speakers per dialogue and has a total of over 13.000 dialogues with their emotions manually labeled.
-The utterances are labeled with the following emotions: anger, disgust, fear, joy, neutral, sadness, and surprise.
+The utterances are labeled with the following emotions: \textit{anger}, \textit{disgust}, \textit{fear}, \textit{joy}, \textit{sadness}, and \textit{surprise}.
 \end{description}
 
-Following are sample dialogues from two of the corpora to showcase how the speakers, dialogues and their respective emotions are structured.
-Additionally, we want to demonstrate the difference in topics of the dialogues that might be important to notice when selecting a corpus for a problem.
+Following are sample dialogues from two corpora to showcase how the speakers, dialogues and their respective emotions are structured.
+Additionally, we want to demonstrate the difference in the dialogues' topics that might be important to notice when selecting a corpus for a problem.
 
 \subsection{Corpus Structure}
-Each of our corpora are structured similarly.
-They contain a training set, a test set, and development set, usually shortened to dev or val (for validation).
-These sets are all structured identically and containing utterances randomly selected from the full corpus, with the bulk of the utterances in train, and test and val of approximately the same size.
-As the name implies, the training set contains the data that the models sets out to train on.
+Each of our corpora is structured similarly.
+They contain a training set, a test set, and a development set, usually shortened to dev or val (for validation).
+These sets are all structured identically and containing utterances randomly selected from the full corpus, with the bulk of the utterances in train and test and val of approximately the same size.
+For our corpora, the train set contains between 75$\%$ to 85$\%$ of the total amount of utterances, while the rest is split somewhat evenly between dev and test.
+As the name implies, the training set contains the data that the models set out to train on.
 The other two are used together to produce our results.
 While the test set is where we get our results from, we do not want to directly access our results from it as there is a risk that we accidentally manually overfit our model when tuning it based on our intermediate results.
-Instead, the result that gets reported is the F1 score in the test set for the epoch where the dev set produced the best F1 score.
+Instead, the result reported is the F1 score in the test set for the epoch where the dev set produced the best F1 score.
 
@@
 \label{corpus_excerpts}
-\paragraph{MELD.} The annotated emotions in MELD are chosen by the majority verdict of five independent \textit{Amazon mechanical turks}.
-Each one of these are tasked to manually examine utterances one by one and then select what emotion out of the options is most suited for that utterance.
+
+Following this paragraph, we have selected one excerpt each from the MELD and IEMOCAP corpora. These excerpts are displayed in order to demonstrate what differences and similarities exist.
+
+\paragraph{MELD.} The annotated emotions in MELD are chosen by the majority verdict of five independent \textit{Amazon Mechanical Turks}.
+Each one of these is tasked to manually examine utterances one by one and then select what emotion out of the options is most suited for that utterance.
 %?? Here have afew intro words as for IEMOCAP ??
@@
 IEMOCAP determines the emotion of an utterance through the response of four independent observers.
-If the observers can reach a clear consensus, that consensus is chosen as the motion, otherwise, it is put as undecided.
+If the observers can reach a clear consensus, that consensus is chosen as the emotion; otherwise, it is undecided.
 
@@
 \end{table}
+
+We can notice a similarity in type between the two excerpts, both being a friendly conversation between friends.
+Observing the emotion classification, it is clear that MELD uses more emotionally loaded words for its classification, while IEMOCAP has a more neutral spectrum.
+Using Joey's first line in the MELD excerpt as an example, \textit{Joy} might be considered an exaggerated description considering the utterance.
 
@@
 \begin{center}
-\begin{tabular}{lcccc}
+\begin{tabular}{lccc}
 \hline
@@
 \end{center}
-\caption{Comparison of the weighted average F1-score between CASTLE and other models taken from \citet{ghosal2020cosmic} and \citet{Shen2020DialogXLAX}. State-of-the-art results in bold. }
+\caption{Comparison of the weighted average F1-score between CASTLE and other models. Taken from \citet{ghosal2020cosmic} and \citet{Shen2020DialogXLAX}. State-of-the-art results in bold. }
 \label{results_table}
@@
 As Table \ref{results_table} shows, CASTLE achieves state-of-the-art results for both the MELD and IEMOCAP corpora.
-While not producing state-of-the-art results for EmoryNLP, CASTLE manages to produce a score within 99$\%$ of the current state-of-the-art.
+While not producing better results for EmoryNLP, CASTLE manages to produce a score within 99$\%$ of it.
+
+\begin{table}[!ht]
+\begin{center}
+\begin{tabular}{lrrr}
+\hline
+\textbf{Hyperparameter values} & \textbf{Batch size}           & \textbf{Dropout}        & \textbf{Learning rate}  \\ \hline
+MELD                     & 128          & 0.35          & 0.0007 \\
+IEMOCAP                     & 16           & 0.25          & 0.001 \\
+EmoryNLP             & 64           & 0.3          & 0.0003 \\ \hline
+\end{tabular}
+\end{center}
+\caption{The hyperparameters used for the F1-scores produced by CASTLE in Table \ref{results_table}}
+\label{hyper_table}
+\end{table}
 
@@
 %As we can see CASTLE obtains sota
-%Prata om att dessa resultaten är state-of-the-art och vad det innebär
+%Prata om att dessa resultaten är state-of the art och vad det innebär
 
@@
 \chapter{Discussion}
-In this chapter we start by analyzing the results that CASTLE achieved.
+In this chapter, we start by analyzing the results that CASTLE achieved.
 Consequently, the limitations that impeded the project will be mentioned and discussed.
-Finally we will mention in what ways the project can further be explored.
+Finally, we will mention in what ways the project can further be explored.
 
 %We compare our results to the other baselines results that COMSIC presents in their report.
-%When we started working with the project our main goal was in creating that produced state-of-the-art results
+%When we started working with the project our main goal was in creating that produced state-of the art results
 
@@
 
-As stated in the \nameref{result_chapter} Chapter, CASTLE accomplishes state-of-the-art results in multiple corpora annotated for emotion recognition.
-This makes the case for CASTLE being an improvement over current state-of-the-art models, namely COSMIC for MELD and DialogXL for IEMOCAP.
-EmoryNLP was the only corpus that we managed to run CASTLE on without achieving state-of-the-art results.
-
-There might be several reasons for this but we believe that one of them is more likely than the others.
+As stated in the \nameref{result_chapter} chapter, CASTLE accomplishes state-of-the-art results on two of the three corpora annotated for emotion recognition.
+This makes CASTLE an improvement over current models, namely COSMIC for MELD and DialogXL for IEMOCAP.
+EmoryNLP was the only corpus that we managed to run CASTLE on without achieving better results.
+
+There might be several reasons for this, but we believe that one of them is more likely than the others.
 As EmoryNLP was the last corpus that we optimized CASTLE for, there was not enough time to thoroughly tune the hyperparameters.
@@
 
-Working in this way is somewhat of a scattershot approach where we depend unpleasantly much on the results of random searches.
-While the random searches clearly produced competitive results, there is nothing to indicate that whether or not an optimal set of values was overlooked.
-Since CASTLE produces results within 99$\%$ of current state-of-the-art, it is our belief that with additional time and resources CASTLE could produce state-of-the-art results for EmoryNLP as well.
-
-The evidence demonstrates that CASTLE outperforms its competitors on the MELD and IEMOCAP corpora while remaining competitive on EmoryNLP and with inconclusive results for DailyDialog.
-This fulfills the primary aim of our thesis that we defined in section \ref{research_questions} and makes CASTLE one of, if not, the most viable systems to adapt for practical emotion recognition applications.
+Working in this way is somewhat of a scattershot approach, where we depend unpleasantly much on the results of random searches.
+While the random searches clearly produced competitive results, there is nothing to indicate whether an optimal set of values was overlooked.
+Since CASTLE produces results within 99$\%$ of the current state of the art, we believe that CASTLE could produce state-of-the-art results for EmoryNLP with additional time and resources.
+
+The evidence demonstrates that CASTLE exceeded the best F1-scores on the MELD and IEMOCAP corpora while remaining competitive on EmoryNLP and with inconclusive results for DailyDialog.
+This fulfills our thesis' primary aim that we defined in section \ref{research_questions} and makes CASTLE one of the most viable systems to adapt for practical emotion recognition applications.
 
 The primary improvement we implemented was replacing an existing GRU layer with a more complex LSTM one.
-As mentioned earlier, the theory states that LSTM layers is likely to outperform GRU layers, when there is enough data, at the expense of computation power.
-Since LSTM is a more complex structure with an additional gate as well as having the concept of utilizing memory in a cell state, it is no surprise that it has the potential to outperform GRU layers.
-When debating if to use CASTLE or some other model, it would boil down to whether or not increase in certainty is worth the cost in computational power.
+As mentioned earlier, the empirical observations show that LSTM layers are likely to outperform GRU layers when there is enough data at the expense of computation power.
+Since LSTM is a more complex structure with an additional gate and having the concept of utilizing memory in a cell state, it is no surprise that it can outperform GRU layers.
+When debating if to use CASTLE or some other model, it would boil down to whether an increase in certainty is worth the computational power cost.
 
@@
 %Vårt main mål var att förbättra performance med LSTM.
-%Kan inte skriva förrän state-of-the-art på Emory
+%Kan inte skriva förrän state-of the art på Emory
 %Our results are the average of  10 samples since we wanted them to be comparable to COSMICs. 
@@
 
-At the start of this project, the scope was a lot larger than the one explored during this thesis.
-Originally the model was intended to also perform keyword extraction in addition to emotion recognition.
+At the start of this project, the scope was much larger than the one explored during this thesis.
+Originally the model was also intended to perform keyword extraction in addition to emotion recognition.
 This was based on discussions with Telavox, where the idea was a model that could identify the topic of the conversation and the emotions of the parties involved.
-Unfortunately, a corpus fitting for this task does not exist at this time.
+Unfortunately, a corpus fitting for this task did not exist at this time.
 This meant that in order to fulfill the original scope, we would have to both create and annotate such a corpus.
-Such a task could be considered a master's thesis by itself and certainly was not possible to complete in tandem with creating a model, at least not in the given time frame.
-As the project is only intended to run for 20 weeks, we had to make a decision about how to proceed and we decided that a model producing state-of-the-art results but not in customer service would be more rewarding, for both us and Telavox, than a specialized customer service model providing average results.
-
-When deciding the scope, we also underestimated just how time consuming this project was going to be.
+Such a task could be considered a Master's thesis by itself and certainly was not possible to complete in tandem with creating a model, at least not in the given time frame.
+As the project was only intended to run for 20 weeks, we had to decide on how to proceed, and we decided that a model producing state-of-the-art results but not in customer service would be more rewarding for both Telavox and us than a specialized customer service model providing average results.
+
+When deciding the scope, we also underestimated just how time-consuming this project was going to be.
 The initial thought of both emotion recognition and keyword extraction, even with a specialized corpus available from the start, would be a difficult challenge to complete in the time frame.
-As we feel that more time could have been spent on optimization with potential better results for EmoryNLP, adding a wholly new task in the form of keyword extraction would most likely have produced lacking results for both tasks.
-
-As the laptops we were using did not contain a GPU, running our code locally was not possible and we were instead bound by the rules of Google Colab.
+As we feel that more time could have been spent on optimization with potentially better results for EmoryNLP, adding a wholly new task in keyword extraction would most likely have produced lacking results for both tasks.
+
+As the laptops we were using did not contain a GPU, running our code locally was not possible, and the rules of Google Colab instead bound us.
 One of these rules is that no notebook can run for more than 24 hours straight.
-Hyperparameter optimization is a time consuming task that is solved by trial and error, and this restriction meant that we could not run as many trials at a time as we wanted which in turn led to less optimization as there simply was not enough time.
+Hyperparameter optimization is a time-consuming task that is solved by trial and error, and this restriction meant that we could not run as many trials at a time as we wanted, which in turn led to less optimization as there was not enough time.
 
@@
 However, due to the previously discussed hardware limitations, we did not have enough memory to run CASTLE on DailyDialog.
-Despite continuous attempts at optimization, we were unable to complete a finished run with the corpus and instead repeatedly ran out of memory.
+Despite continuous attempts at optimization, we could not complete a finished run with the corpus and instead repeatedly ran out of memory.
 
@@
 Firstly, DailyDialog is close to a magnitude larger than the other corpora used in this experiment.
-Secondly, LSTM is a complex architecture with more gates and calculations which leads to higher demand in hardware.
+Secondly, LSTM is a complex architecture with more gates and calculations, which leads to higher demand in hardware.
 As Google Colab has a limited amount of memory for us to utilize, we did not have to option to run the corpus on the service.
@@
 
-Considering the fact that CASTLE did not produce state-of-the-art results for the EmoryNLP corpus, the most natural continuation for the project is to have CASTLE achieve this goal.
-Our proposal is to simply spend more time with hyperparameter optimization and explore what results that might provide.
-Even though CASTLE managed to achieve state-of-the-results for MELD and IEMOCAP, an extensive hyperparameter optimization could possibly improve the results further.
+\label{futurework}
+
+Because CASTLE did not produce better results for the EmoryNLP corpus, the project's most natural continuation is to have CASTLE achieve this goal.
+We propose to spend more time with hyperparameter optimization and explore what results that might provide.
+CASTLE could probably also improve even further the performance of MELD and IEMOCAP with extensive hyperparameter optimization.
 
@@
 
-Another natural continuation for the project is to expand the scope to what it initially was defined as.
-As stated in \nameref{limitations}, the initial goal of the project was to create a model for analyzing written customer service conversations and using keyword extraction as a complement to said analysis.
+Another natural continuation of the project is to expand the scope to what it initially was defined as.
+As stated in \nameref{limitations}, the project's initial goal was to create a model for analyzing written customer service conversations and using keyword extraction as a complement to said analysis.
 CASTLE could very well be used as a framework for completing this type of task.
@@
 CASTLE is undoubtedly able to analyze written conversations with great success.
-However, the corpora used for training CASTLE are not completely suited for customer service specific dialogues.
+However, the corpora used for training CASTLE are not completely suited for customer service-specific dialogues.
 When searching for appropriate corpora, there was no customer-service themed corpus annotated with emotions available.
-One possibility is to create such a corpus using an already existing unannotated customer service related corpus and labelling it with emotions.
+One possibility is to create such a corpus using an already existing unannotated customer service-related corpus and to label it with emotions.
 A couple of methods has previously been mentioned in section \nameref{corpus_excerpts} for how a corpus can be annotated with emotions.
@@
 %Skapa ett dataset specifikt för customer service
-%Ytterligare ändringar som kan ge state-of-the-art
+%Ytterligare ändringar som kan ge state-of the art
 %Mer utförliga hyperparameter optimization.
 
-\chapter{Conclusions}
-When we started this project, our goal was to find answers to the research questions mentioned in section \ref{research_questions}.
-After completing this project, it is safe to say that creating a model capable of fulfilling our primary aim was possible, as CASTLE is just that.
-CASTLE also proves that an LSTM architecture can in fact outperform a GRU one.
-When comparing CASTLE to the previous state-of-the-art models we show, as presented earlier, that CASTLE surpasses those models.
-Our secondary aim and last research question unfortunately has to remain unanswered due to time limitations.
+\chapter{Conclusion}
+When we started this project, our goal was to find answers to the research questions mentioned in Section \ref{research_questions}:
+
+\begin{itemize}
+    \item \textit{Is it possible to create a model that fulfills our aims?}
+    \item \textit{Can we build an architecture that improves the state of the art?}
+    \item \textit{If yes, how does our model compare to other models in the same field evaluating on the same metrics?}
+    \item \textit{After implementing a fully functioning model, how can we adapt it to predict utterances in real-time and in the context of an interactive service?}
+\end{itemize}
+
+After completing this project, it is safe to say that: 
+\begin{enumerate}
+    \item We could create a model capable of fulfilling our primary aim was possible, as CASTLE is just that.
+    \item CASTLE also proves that, with our datasets, an LSTM architecture outperforms a GRU one. As LSTM is a more complex architecture, one has to consider that CASTLE will require more resources. 
+    \item When comparing CASTLE to the previous state-of-the-art models, we show, as presented earlier, that CASTLE surpasses those models. 
+    \item Our secondary aim and last research question, unfortunately, has to remain unanswered due to time limitations.
+In section \ref{futurework}, we mention how CASTLE can be extended into fulfilling the secondary aim as well.
+\end{enumerate}
+
 All in all, we consider CASTLE a success as it produces state-of-the-art results for two corpora and has the potential of achieving it for the remaining two as well, thus fulfilling our primary aim.
@@
 
+\ifx false
 \begin{appendices}
@@
 \end{appendices}
+\fi
 

### 2021-01-27 msccls/popsci/popsci.tex
report done!

@@
 % Please specify the presentation date
-\presentationsdag{2016-04-28}
+\presentationsdag{2021-02-12}
 
 % use either of these commands to specify the title of your thesis
-% \examensarbete{}
+\examensarbete{Emotion Recognition in Written Conversations Using CASTLE}
 % To create a title in two rows, leave examensarbete blank and fill in examensarbeteTwoRows.
-\examensarbeteTwoRows{Application Specific Instruction-set Processor Using a Parametrizable multi-SIMD}{Synthesizeable Model Supporting Design Space Exploration}
+%\examensarbeteTwoRows{Application Specific Instruction-set Processor Using a Parametrizable multi-SIMD}{Synthesizeable Model Supporting Design Space Exploration}
 \student{Nick Persson, Filip Karabeleski}
@@
 % Your pop-sci title should be different (more catchy) than your thesis title
-\title{Parametrisk processor modell för design utforskning}
+\title{Identifiering av känslor i skriftliga konversationer}
 
@@
 % not more than 4 rows!
-\theabstract{Applikations-specifika processorer är allt mer vanligt för få ut rätt prestanda med så lite resurser som möjligt.
-Detta arbete har en parametrisk modell för att kunna testa hur mycket resurser som behövs för en specifik applikation.}
+\theabstract{Identifiering av känslor i konversationer är ett notoriskt svårt problem för datorer att förstå då konversationer ofta innehåller subtila indikationer som kan vara svåra att urskilja, till och med för en människa.
+Vi har undersökt möjligheten att förbättra datorers kompetens inom detta område.
 
-{\noindent För att öka prestandan i dagens processorer finns det vektorenheter och flera kärnor i processorer.
+}
+\ifx false
+\begin{figure}[!bth] % Use pictures in your pop-vet!
+\includegraphics[width=\linewidth]{msccls/popsci/popsci_image.png} 
+%\caption{En fin bild}
+\end{figure}
+\fi
+
+{\noindent
+
+Identifiering av känslor är en form av kategorisering där utvecklarna sammanställer en lista av känslor att använda som kategorier.
+Därefter går datorn igenom en samling data, eller dataset, och kategoriserar varje mening med den känslan som passar bäst av de känslor i listan.
+
+Detta görs oftast med hjälp av maskininlärning, då det hade varit alldeles för tidskrävande att göra det manuellt.
+Maskininlärning kan låta som magi, men är i själva verket bara en upprepande process där svar på uppgiften matas in till datorn och den lär sig mönster och hur den ska hitta svar på liknande uppgifter.
+Den här processen kallas träning.
+Ett vanligt problem med maskininlärning är att datorn kan få svårigheter att förutse nya sorters uppgifter som skiljer sig från det den tidigare lärt sig av.
+
+%För att öka prestandan i dagens processorer finns det vektorenheter och flera kärnor i processorer.
 Vektorenheten finns till för att kunna utföra en operation på en mängd data samtidigt och flera kärnor gör att man kan utföra fler instruktioner samtidigt.
@@
 
-%För att kunna välja vilken typ av processor som är rätt för den specifika applikationen krävs det ofta att man snabbt kan testa olika prototyper.
-Att implementera dessa till hårdvara kan ofta vara tidskrävande och ifall det visar sig att implementationen inte klarar dem kraven man ställt för prestanda och energieffektivitet, måste man designa för nya parametrar och mer tid har blivit slösat.
-%Om den här processen istället kan göras automatiskt utifrån dessa design-parametrar kan man teoretiskt spara en massa tid.
 Modellen testades med olika multimedia program.
@@
 
-\begin{figure}[!bth] % Use pictures in your pop-vet!
-\includegraphics[width=\columnwidth]{samplePic.pdf} 
-%\caption{En fin bild}
-\end{figure}
 Resultatet visar att det finns en prestanda vinst jämfört med generella processorer men att detta också ökar resurserna som behövs.

### 2021-01-29 msccls/popsci/popsci.tex
to the moon

@@
 
-I mitt examensarbete har jag skrivit en modell som kan användas för att snabbt designa en processor enligt vissa parametrar.
+I vårt examensarbete har vi förbättrat en av de bäst presterande modellerna som finns i dagsläget.
+Vi har ändrat den inre arkitekturen men bibehållit dess logik och beteende.
+Vad detta leder till är en förbättrad modell som vi har döpt till CASTLE.
+Denna "tänker" på samma sätt som tidigare men utför mer sofistikerade beräkningar för att få fram ett mer precist svar.
+CASTLE har testats på några av de mest förekommande dataseten inom området.
+Dataseten är strukturerade som dialoger mellan flertalet personer och kan vara allt ifrån improvisation till direkta utdrag ur den populära 90-tals serien Vänner.
+
+Att testa på samma dataset som de andra gör ger mer tillförlitlighet till resultaten.
+
+%I mitt examensarbete har jag skrivit en modell som kan användas för att snabbt designa en processor enligt vissa parametrar.
 Dessa parametrar väljs utifrån vilket eller vilka program man tänkta köra på den.

### 2021-01-29 msccls/popsci/popsci.tex
still 193.06, but done

@@
 Vi har ändrat den inre arkitekturen men bibehållit dess logik och beteende.
-Vad detta leder till är en förbättrad modell som vi har döpt till CASTLE.
+Detta leder till är en förbättrad modell som vi har döpt till CASTLE.
 Denna "tänker" på samma sätt som tidigare men utför mer sofistikerade beräkningar för att få fram ett mer precist svar.
 CASTLE har testats på några av de mest förekommande dataseten inom området.
-Dataseten är strukturerade som dialoger mellan flertalet personer och kan vara allt ifrån improvisation till direkta utdrag ur den populära 90-tals serien Vänner.
-
-Att testa på samma dataset som de andra gör ger mer tillförlitlighet till resultaten.
+Dataseten är strukturerade som dialoger mellan flertalet personer och innehåller allt ifrån improvisation till direkta utdrag ur den populära 90-tals serien Vänner.
+Att testa på samma dataset som ledande personer inom området använder ger mer tillförlitlighet till våra resultat.
 
@@
 
-Modellen testades med olika multimedia program.
+%Modellen testades med olika multimedia program.
 Den mest beräkningsintensiva och mest upprepande delen av programmen användes.
@@
 
-Resultatet visar att det finns en prestanda vinst jämfört med generella processorer men att detta också ökar resurserna som behövs.
-Detta trots att den generella processorn har nästan dubbelt så hög klockfrekvens än dem applikations-specifika processorerna.
-Resultatet visar också att schemaläggning av instruktionerna i programmen spelar en stor roll för att kunna utnyttja resurserna som finns tillgängliga och därmed öka prestandan.
-Med den schemaläggningen som utnyttjade resurserna bäst var prestandan minst 79\% bättre än den generella processorn.
+Våra resultat visar att CASTLE uppnår rekordhöga värden på hur korrekt modellen förutspår ett yttrandes känsla på majoriteten av de stora dataseten.
+Resultaten är i genomsnitt 1$\%$ bättre än vad som tidigare uppnåtts vilket kan låta insignifikant vid en första anblick.
+Tar man dock i beaktning att andra aktörer inom området använder kluster på hundratals datorer för att räkna ut sina resultat så är det faktum att vi fick bättre resultat med enbart en dator anmärkningsvärt.
+
 }

### 2021-02-23 msccls/MScTemplateCS.tex
literally a new report - tack

@@
 We started from an existing baseline and, using empirical observation; we combined it with a previously untested set of architectures.
-We created a new framework that builds upon the previous state of the art and complements it with more effective recurrent devices.
-As a result, using this framework, CASTLE, we could improve the F1-score of previously published results on two of the three corpora we evaluated.}
+We created a new framework, CASTLE, that builds upon the previous state of the art and complements it with more effective recurrent devices.
+As a result of using this framework, we could improve the F1-score of previously published results on two of the three corpora we evaluated.}
 % produces state-of the art results on several corpora commonly used as benchmarks for these types of tasks.
@@
 
-Concretely, NLP fields of application include machine translation, information extraction, and text categorization.
-For the last field, text categorization consists of automatically assigning a category to a text, for instance, to decide if a message is spam or not.
-
-Since natural languages naturally evolve as time goes on and their adherence to structure diminishes, it is difficult to define explicit rules to follow.
-For this reason, there is a need for machine learning models to aid the computer in identifying these diffuse rules.
+Concretely, NLP fields of application include machine translation \citep{machine_translation}, information extraction \citep{information_extraction}, and text classification \citep{text_classification}.
+For the last field, text classification consists of automatically assigning a class to a text, for instance, to decide if a message is spam or not.
+
+Since natural languages naturally evolve as time goes on and their structure changes \citep{lyons1968introduction}, it is difficult to define explicit rules to follow.
+As a result of this, instead of spending time trying to define these rules, including taking into account all corner cases, machine learning can be used in order to find patterns based on samples of real-world text.
+According to \citet{NLP_challenges} having the machine learning model find patterns in a dataset is more effective than a manual establishment of rules.
+
+% HÄR MÅSTE NÅGOT MER SKRIVAS
 
@@
 This goal can be challenging to achieve, and there is no straightforward way to do so.
-By analyzing prior conversations, companies can gain valuable insights in the most effective way to please the customer.
-However, customer service is a massive industry, and it would take countless hours to analyze the conversations from even a single company manually.
-
-The company sponsoring this project approached us with an idea to try and solve this problem using machine learning.
-A usual technique for these problems is \textit{sentiment analysis}, where utterances are labeled either positive or negative.
-We instead decided to take another route, namely \textit{emotion recognition}.
-Emotion recognition focuses on labeling utterances from a non-binary set of emotions instead of sentiment analysis' binary method.
+By analyzing prior conversations between customers and service-agents, companies can gain valuable insights in the most effective way to please the customer.
+
+The company sponsoring this project approached us with an idea to try and identify emotions in a customer service call using machine learning.
+This is an instance of text classification, where the emotions are the classes.
+This is called \textit{emotion recognition}.
+Emotion recognition focuses on labeling utterances from a non-binary set of emotions instead of a binary representation of for example positive or negative.
 This allows analysis with greater specificity.
 
-To further the academic relevance of this project, we decided that we would attempt to contribute to the field of emotion recognition by achieving state-of-the-art results.
-
-\section{State of the art}
+In order to narrow the project's scope, the conversations analyzed were exclusively in English. %++written conversations
+To further the academic relevance of this project, we decided that we would attempt to contribute to the field of emotion recognition by achieving better than current state-of-the-art results.
+
+\section{Previous work}
 %The progress of emotion recognition has been accomplished by 
@@
 
+At the start of the project, we read some recently published research in the field of deep learning, mainly the article by \citet{edseee.922172720200601} that states that GRU is not as suited for small pieces of text and large corpora as LSTM.
+This is where we got the idea to test whether the architectural change to LSTM could improve the results.
+
 In this report, the term \emph{state of the art} will be recurringly mentioned as it is a fundamental part of our aim and result.
-The term refers to the highest level of achievement accomplished.
-A common goal in machine learning projects is to attempt to raise the bar for the current
-results across various metrics for the task that the project is trying to solve.
+The term refers to the highest level of achievement accomplished at a certain point of time.
+In this project, we will refer to the highest results achieved when we started this project as state of the art.
+A common goal in machine learning projects is to attempt to raise the bar for the current results across various metrics for the task that the project is trying to solve.
 Presenting state-of-the-art results is an excellent approach in proving that an implementation or algorithm is valid and valuable.
@@
 %Description of what it is and why it plays a big part in advancing algorithms and models. Deepminds Alphafold has led to groundbreaking results in a competitive environment.
+
+%Related work, fråga Pierre
+% COSMIC?
+% Bouthillier? 
+% DialogXL?
 
@@
 \label{research_questions}
-This thesis' primary aim is to create an emotion recognition model that can produce state-of-the-art results.
-After this, a secondary aim is to investigate whether it is possible to create a dynamic model to provide interactive support in customer service-related conversations.
+This thesis' primary aim is to create an emotion recognition model that can produce better than state-of-the-art results.
+After this, a secondary aim is to investigate whether it is possible to create a model to assist the service-agent in customer service-related conversations in real time by providing them with the customer's emotions.
 The following questions are used as guidelines on how to realize these aims.
-
+%
 \begin{itemize}
     \item \textit{Is it possible to create a model that fulfills our aims?}
+    %How to predict emotions??
+    %Which XXX architecture 
+    %Can we
 %?? THe first question is can we create such a model??
     \item \textit{Can we build an architecture that improves the state of the art?}
+
+    %What 
+    %What is the performance
 %?? here your statement is too specific as you did not introduce the terms. Use vague words such as What are the different option to design an architecture and train a model?? 
@@
 
+\section{Division of work}
+
+All of the implementation has been done using pair-programming.
+This allowed us both to get a full understanding of the code base.
+In terms of writing, all chapters except \nameref{chap:theory} has had equal contribution on all parts.
+In Theory, Filip wrote sections \ref{sect:ml} to \ref{sect:hyper} while Nick wrote \ref{sect:nlp} and \ref{sect:eval}.
+
+\section{Contributions}
+
 %NLP, ML, Speech-to-text och (STT), API. Modality. RNN, GRU, LSTM, BPTT(?), utterance, speaker, dialogue,
@@
 \chapter{Theory}
+\label{chap:theory}
 
@@
 
-%In order to solve most emotion recognition problems, there are two main concepts that are fundamental to build an accurate model.
-Firstly, the model should be created with deep learning algorithms.
-This enables it to find patterns that it otherwise could not be able to.
-Secondly, to effectively process the utterances it has been tasked to analyze, a word vector is often used to store the most commonly used words for easier analysis.
-??
-I propose this instead??
+%SNACKA OM NLP
 
@@
 
+There are namely two common architectures \citep{staudemeyer2019understanding} that are used in deep learning: Long Short-Term-Memory (LSTM) and Gated Recurrent Unit (GRU).
+There is also the simpler \citep{Sherstinsky_2020} Recurrent Neural Network.
+We will describe these three further and discuss their advantages and drawbacks.
+
 \begin{figure}[h!]
@@
 \section{Machine Learning}
-Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem rather than to program the system explicitly.
+\label{sect:ml}
+Machine learning, according to \citet{franoischollet2017learning}, is a field of study in artificial intelligence that aims to train a system to solve a problem rather than to program the system explicitly.
 By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem.
@@
 
+Supervised learning is used for mostly classification and regression problems \citep{franoischollet2017learning}.
+Regression is the task of finding a function that predicts a continuous value while classification is the task of finding a function that distributes input data into classes \citep{alpaydin2014introduction}.
+In our work, we use supervised learning for classifying emotions in written conversations.
+We design our model to learn how to map utterances to specified emotions.
+
 \subsection{Neural Networks and Deep Learning}
 
+%WHERE/WHEN ML IS NOT ENOUGH
 Neural networks are machine learning systems loosely inspired by biological neural networks.
@@
 
+Deep learning has been shown to be beneficial for solving NLP problems \citep{bhardwaj2018deep} and is predicted to become the norm when solving these types of problems \citep{otter2019survey}.
+Therefore, all the work in this project is done using deep learning.
+Because of the multiple layers in deep learning, the algorithms learn multiple levels of representation \citep{Hinton2007428} opposed to the traditional machine learning NLP systems that heavily depend on manual work \citep{young2018recent}.
+%Deep learning, especially recurrent neural models, is able to capture the sequence information in a much better sense.
+
 Each of the layers in a deep learning network manipulates its input data in some way, and the result is stored in the layer's weights.
@@
 
-One of the simplest neural network classes used in deep learning is \emph{recurrent neural networks} (RNN).
+One of the simpler neural network classes, according to \citet{Sherstinsky_2020}, used in deep learning is the \emph{recurrent neural network} (RNN) \citep{Rumelhart1986LearningRB} and can be observed in Figure \ref{fig:rnn_node}.
 RNN process either sequential data or time-series data \citep{lipton2015critical}.
-Since the inputs are sequential, all the inputs are dependent on each other in one way or another.
-RNNs are suitable for processing sequences of text, audio, and video, where every computation will be performed on each element in the sequence.
+Since the inputs are sequential, all the inputs are dependent on each other.
+RNNs are suitable for processing sequences of text, audio, and video, where every computation will be performed on each element in the sequence \citep{Sherstinsky_2020}.
 
@@
 If the sequence is too long, the network will struggle to try to look back in time longer than a few steps away.
+For example, in emotion recognition, this can be particularly important when trying to interpret emotions if, e.g., tangents from the main of the conversation happen.
 In many of cases, this will result in leaving out important information from the prediction.
 There are instead other alternatives that consider this weakness.
+
+%to understand relationships between distant events described in a book to answer questions about it
 
@@
 \subsection{Long Short-Term Memory -- LSTM}
+\label{LSTM}
 %?? I would have this section before GRU ??
@@
 
+Compared to RNN, shown in Figure \ref{fig:rnn_node}, we can see that LSTM, Figure \ref{fig:lstm_node}, has a higher number gates and computations.
+
 \subsection{Gated Recurrent Unit -- GRU}
+\label{GRU}
 %Byta alla "node" till "cell" för att lättare kunna referera till LSTM-celler i CASTLE
@@
 
-Basic GRU nodes  are primarily constructed of an \textit{update gate} and a \textit{reset gate}.
+Basic GRU nodes  are primarily constructed of an \textit{update gate} and a \textit{reset gate}, as can be seen in Figure \ref{fig:gru_node}.
 The update gate regulates the amount of information passed along to the next state, while the reset gate, much like the forget gate of LSTM, decides the amount of past information that is going to be disregarded henceforth.
@@
 Compared to \textit{LSTM} layers, GRU layers are cheaper to run since they use fewer parameters but have less representational power.
-It has been shown by \citet{Gruber2020AreGC} that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers.
+It has been shown by \citet{Gruber2020AreGC} that GRU layers have had better performance on smaller datasets with less frequent recurrences of words than LSTM layers.
 
@@
 \section{Training Process}
+
 To train a model, there must exist some dataset, either labeled or unlabeled depending on whether supervised or unsupervised learning is the targeted method that feeds the model with input.
@@
 
-The training of a model is done step by step in the form of \textit{epochs}. Epochs represent the number of times 
-the model iterates through a dataset intended to train the model.
+The training of a model is done step by step in the form of \textit{epochs}.
+Epochs represent the number of times the model iterates through a dataset intended to train the model.
 Following each epoch, a \textit{loss function} measures how incorrect the model's previous prediction was and produces a value for the \textit{loss}.
@@
 
+\section{Hyperparameters}
+\label{sect:hyper}
+Many parameters in a machine learning model are determined during the training process.
+There exists, however, parameters that must be fixed before running the algorithm.
+These parameters are called hyperparameters and govern the rules for how the model approaches the learning task \citep{weerts2020importance}.
+As there is too large a number of different hyperparameters to possibly cover them all, a select few has been chosen for us to focus on in this work.
+The ones we chose are \textit{dropout}, \textit{batch size}, and \textit{learning rate} \citep{hyperparameters}.
+
+\begin{enumerate}
+    \item  Batch size governs how many training examples will be utilized in each iteration.
+Batch size is a heavily discussed topic, with scientific conclusions being unanimous regarding a high or low value being preferential.
+\citep{KANDEL2020312}.
+A small batch size can converge faster than a larger one, but the larger batch can potentially find a lower minima that a smaller batch can not reach.
+A small batch size also requires a small learning rate to prevent it from overshooting the minima, but can have a significant regularization effect due to a higher variance.
+    \item  Learning rate govern how aggressively the model is influenced by new information and has a large effect on gradient descent.
+A high learning rate speeds up the learning process at the risk of not converging, while a low value risk taking too long to converge \citep{learning_rate}.
+%Snacka om gradient descent
+     \item Dropout is a value between 0 and 1 that regulates what percentage of learned data the model randomly chooses to forget.
+With a low dropout, or without it altogether, the model becomes heavily reliant on the training data with accurate predictions for it but with worse results for new data.
+This concept of the model becoming too reliant on the training data is called overfitting.
+A too high dropout hampers the learning process and causes the model to guess more than it is predicting \citep{dropout}.
+\end{enumerate}
+
 %bild på training process
@@
 \section{Natural Language Processing -- NLP}
-One of the reasons that make natural language processing a challenge is that natural languages are minted with unstructured types of data such as inflection, dialects, or sarcasm.
+\label{sect:nlp}
+There are several factors that make natural language processing a challenge.
+Natural languages are constantly changing and evolving which makes it especially difficult to define rules that dictate language \citep{neuralnetworkmethods}.
 Natural languages are also oftentimes ambiguous and require a context-based interpretation that computers have a hard time understanding.
@@
 
-Emotion is a key part of human language, and therefore it comes naturally that teaching a computer to understand emotion in human language is a big part of NLP.
-Initially, the focus laid on sentiment analysis, a field of study focused on determining the binary polarization of a piece of text as either positive or negative.
+Emotion is a key part of human language \citep{languageEmotion}, and therefore it comes naturally that teaching a computer to understand emotion in human language is a big part of NLP.
+One way to analyze emotions in text is to use sentiment analysis, which according to \citet{sentimentDefinition} is a field of study focused on determining the binary polarization of a piece of text as either positive or negative.
 %Godtycklig bit om sentiment analysis. 
-Emotion recognition is an extension of sentiment analysis that, instead of just determining text's polarization, attributes emotion to it.
+Instead of just determining text's polarization, it is possible to attribute emotion to it and this is what emotion recognition is used for \citep{introToEmoRec}.
 This process is more challenging than sentiment analysis, as the differences between emotions are subtler than that of just positive and negative.
-Emotion recognition has also been extended to work on different modalities, such as visual and auditory input.
-
-Instead of working directly with raw input, it first needs to be converted into a numerical representation in order for the model to process it \citep{franoischollet2017learning}.
+Emotion recognition has also been extended to work on different modalities, such as visual and auditory input \citep{singh2020emotion}.
+
+Instead of working directly with unprocessed input from the corpora, it first needs to be converted into a numerical representation in order for the model to process it \citep{franoischollet2017learning}.
 In the case of textual input, this is done by transforming it into a \textit{word vector}.
@@
 This means that synonyms and words often used together are embedded closer to each other, while words seldom used together are more distant in the embedding.
-In addition to distance, the direction could also be a useful tool to extract meaning from the embedding.
-%Temp for formatting, don't forget it's here
+In addition to distance, the direction could also be a useful tool to extract meaning from the embedding \citep{whitaker-etal-2019-characterizing}.
 
@@
 
-Typically there are thousands of these types of interpretable direction vectors in a large word embedding.
-
-%\subsection{RoBERTa}
+Typically there are thousands of these types of interpretable direction vectors in a large word embedding \citep{franoischollet2017learning}.
+Since word embeddings are built around the relation of words, these vectors are naturally created by the sample data having a similarity in how people refer to certain things.
+As people refer to queen in relation to king in the same manner as they would actress to actor, this allows such vectors to exist.
+
+\subsection{RoBERTa}
+Despite the numerous advantages of pre-trained word embeddings, they still contain one crucial flaw: each word is represented by one vector only.
+This means that words with multiple meanings depending on context is only embedded in one part of embedding space, which loses information for other use cases.
+In theory this problem is solved by using LSTM or GRU cells as they have memory which can be used to create a different context for the word.
+It has however been shown that this can be insufficient for context-heavy language features \citep{Zhang_deep}.
+A new method for creating context rich information is the BERT architecture provided by Google \citep{BERT}.
+In this work, we are using a modified version of BERT called RoBERTa \citep{RoBERTa}.
+RoBERTa was created by Facebook as they believed that BERT to be significantly undertrained.
+We are using RoBERTa Large, based and structured identically to BERT Large, containing more parameters than the normal version.
+The model we are using is already fine-tuned for emotion label prediction.
 
@@
 \section{Evaluation}
+\label{sect:eval}
 Many different metrics can be used to evaluate a machine learning model, depending on what kind of model and what its purpose is.
@@
     \centering
-    \hbox{\hspace{6em}\includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}}
+    \hspace*{-3cm}\includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}
     \caption{Binary Confusion Matrix}
@@
 For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative.
-
-The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class.
-For instance, when determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{neutral}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.
-
-%kan snacka om Type I och Type II errors om mer text behövs.
+\begin{figure}[h!]
+    \centering
+    \hspace*{-4cm}\includegraphics[width=\textwidth/4*3]{msccls/explanatory_images/Multiclass_conf.png}
+    \caption{Example of a Multi-class Confusion Matrix}
+    \label{fig:multi_confusion}
+\end{figure}
+Figure \ref{fig:multi_confusion} shows an example of how the matrix could look after trying to predict emotions based on 4 classes.
+The rows show the actual class and the column the class the model predicted for that instance.
+The green squares represent the TP for each class: $TP_{Neutral}$, $TP_{Happy}$, $TP_{Angry}$, and $TP_{Sad}$, with the total sum being denoted $TP_{total}$.
+
+Most evaluation in multi-class scenarios is done on a class-based level, meaning that you observe how the model performed for each of the actual classes.
+This allows conclusions to be drawn about the similarity of classes and where the model requires future work.
+For instance, in our multi-class example, we can observe that the model often predicted angry when sad was correct, but only predicted happy in one instance.
+
+Calculating the TN, FP and FN for each class is not as straightforward as the TP.
+Using the neutral class as an example, $TP_{Neutral}$ is easily observed to be 15.
+For the neutral class, FP would be defined as the model predicting neutral and being incorrect.
+Since the first column of our matrix represents the model predicting neutral and the first cell is when the model is correct, the three remaining can be summed together in order to find $FP_{Neutral} = 8+5+5=18$.
+
+Inversely, $FN_{Neutral}$ are the cases where the instance is neutral but the model predicts something else.
+This is represented by the first row in our matrix, again after removing the first cell.
+$FP_{Neutral}$ can then be calculated to $6+4+5=15$.
+
+$TN_{Neutral}$ are the cases when the model does not predict neutral and is correct.
+This is represented by the remaining cells in the matrix, the ones not present in either the first row or column.
+$TN_{Neutral} = 11+4+2+3+9+5+1+8+9 = 52 $.
 
@@
 To counteract this behavior of favoring the minority class, one could instead calculate the weighted F1-score.
-After calculating the F1-score for each class, they are weighted by the number of true instances for each class, making them proportionate before calculating the final score.
-This gives a score more reflecting on how the model performs, but can also favor identifying the majority class heavily, which might not be the model's goal.
+After calculating the F1-score for each class, they are weighted by the number of actual instances for that class, making them proportionate before calculating the final score.
+This gives a score more reflecting on how the model performs, but can also favor identifying the majority class heavily, which typically is not be the model's goal.
+Favoring the majority class or classes is suitable if priority lies on predicting the more common cases.
 
 When calculating the micro F1 score, instead of calculating the F1 scores separately, the F1 score is calculated directly using the total values for TP, FN, and FP in the Precision and Recall formulas places no bias on either class.
-If we have an example dataset with two minority classes, \textit{anger}, and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class.
+If we again look at figure \ref{fig:multi_confusion} and focus on two of the minority classes, \textit{anger}, and \textit{joy}, we can see that for each instance where anger is true, classifying it as joy would be an FN from the perspective of the anger class.
 From the perspective of the joy class, however, it would instead be an FP.
@@
 
-\chapter{Implementation}
-%Summary of what the chapter is about
-This chapter is mainly focused on giving a clear description of what we have done for the project's duration.
-Firstly we will describe the model that we used as a baseline, \textit{COSMIC}.
-We will cover our model \textit{CASTLE}, its implementation, and what changes were made from COSMIC.
-Lastly, we will describe our setup and what corpora we used.
-%??In your text, you present COSMIC first. Change your intro so that it corresponds to your text??.
-
-We based our research strategy on the work by \citet{bouthillier:hal-02447823} that summarizes and reviews the experimental procedures that are most commonly used in the field of machine learning.
-
-\section{Baseline -- COSMIC}
-When trying to improve existing results in many machine learning fields, a baseline is used instead of starting from scratch.
-A baseline is an already functioning model that provides basic results to the problem at hand.
-This provides the opportunity to allocate resources on improving the model instead of building it from the ground up and allows the machine learning community to expand on and draw benefits from each other's work.
-
-This project has been developed using the COSMIC model \citep{ghosal2020cosmic} as a baseline.
-COSMIC is a GRU-based model implemented with PyTorch that has achieved state-of-the-art results across four corpora commonly used in emotion recognition.
-The idea behind the model is that when classifying an utterance, the emotions of the utterances, both preceding and succeeding, are also taken into account.
-This way, the context of the conversation can be used to predict the emotions of individual utterances more accurately.
-
-%COMSICs arkitektur bygger på att de skapar egna GRU-celler och GRU-lager som bygger på PyTorchs respektive. 
-%Filip
-
-Figure \ref{fig:COSMIC_arch} provides a simplified description of how COSMIC operates in a conversation between person A and person B.
-For each time step $t$, there is an utterance said by either A or B with the other person as a listener.
-At the next time-step $t+1$, the roles are alternated.
-For each utterance, multiple GRU-cells exist that predict both the speaker's intent and listener's response at a certain time-step.
-These cells are used to predict the emotion of the current utterance and the prediction of emotions in other utterances related to said utterance.
-
-COSMIC makes use of five different states to determine what emotion to predict. 
-The four following states interact heavily with each other in a codependent manner.
-
-\begin{itemize}
-    \item \textit{Context state} is a state shared between speaker and listener. It manages the current overall emotional state of the conversation. Affects both the Internal and External state.
-    \item \textit{Internal state} tracks for each participant what emotions they are feeling but not necessarily showing. Affects the Intent state and the Context state.
-    \item \textit{External state} portrays, for each participant, what emotions they are expressing to the other participants through both utterances and reactions. Affects the Context state.
-    \item \textit{Intent state} is a state only present for the current speaker.
-As the name suggests, it represents the speaker's objective is for the current utterance.
-For example, if paying the listener a compliment, the intent could be to make the listener happier.
-This does not affect any other of the three states.
-\end{itemize}
-Internal state and External state are especially codependent as they combine into the sum of a participant's emotional state.
-
-\emph{Emotion state}, the final state, considers the current utterance, the summed emotional state through the Internal and External state, and the Intent state to predict the current utterance.
-This process is repeated for every utterance in every conversation in the corpus.
-
-%Softmax, relu mm
-
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[scale=0.3]{msccls/explanatory_images/COSMIC_arch.png}
-    \caption{COSMIC architecture. From \citet{ghosal2020cosmic}.}
-    \label{fig:COSMIC_arch}
-\end{figure}
-
-%?? You could write a bit more on COSMIC ??
-
-\section{CASTLE}
-Building on COSMIC, we designed CASTLE, COSMIC-based AnalySer Taking advantage of LSTM for Emotion recognition, to answer our research questions.
-We have made several improvements on the COSMIC baseline, one of the fundamental being replacing the internal GRU architecture with an LSTM-based architecture, as the name of the model indicates.
-
-At the start of the project, we read some recently published research in the field of deep learning, mainly the article by \citet{edseee.922172720200601} that states that GRU is not as suited for small pieces of text and large corpora as LSTM.
-This is where we got the inspiration to test whether the architectural change to LSTM could improve the results.
-
-COSMIC takes advantage of the \texttt{cudnn} framework in PyTorch to assist in multiple ways, such as the \texttt{backends.cudnn.benchmark}, which attempts to optimize the algorithm during runtime automatically.
-As we aimed to obtain as much control as possible over how the model behaved, we excluded this and other third-party features that restricted our domain of control.
-
-COSMIC is also implemented with a \textit{residual} connection \citep{he2015deep}, which allows the model to redefine the layers as learning residual functions referencing the layer inputs.
-This is supposed to facilitate the training of the model.
-While we originally wanted to implement this feature in CASTLE, there was simply no space to do it within the project's scope.
-
-%Didnt implement residual for CASTLE tidsbrist
-
-%Removed CUDA Backend since we wanted as little noise as possible interfering with our model We didnt want the reason for the results to be out of our reach. Third party.
-
-%COSMIC-based AnalySer Taking advantage of LSTM for Emotion recognition
-
-\section{Experimental Setup}
-
-%We chose COSMIC as our baseline because SOTA, more experienced in PyTorch
-\subsection{Hyperparameters}
-CASTLE's performance depends on three different hyperparameters. The ones we chose are \textit{dropout}, \textit{batch size}, and \textit{learning rate} \citep{hyperparameters}:
-\begin{enumerate}
-    \item Dropout is a value between 0 and 1 that regulates what percentage of learned data the model randomly chooses to forget.
-This is used to combat overfitting and make sure the model is actually predicting instead of just learning the corpus structure.
-    \item  Batch size governs how many training examples will be utilized in each iteration. A larger value increases the accuracy of the gradient in trade off for a higher memory cost.
-    \item  Learning rate defines the speed at which the model updates its parameters.
-A high learning rate speeds up the learning at the risk of not converging, while a low value will more likely converge, sacrificing speed in the process.
-\end{enumerate}
-
-To optimize our hyperparameters, we implemented the \texttt{Tune} framework offered by \texttt{Ray}.
-This framework allowed us to experiment efficiently with many different configurations and provided a simple way to run trials concurrently.
-Most of our optimization was done with a combination of grid search and random search.
-Random search is a function with two parameters, \textit{start} and \textit{end}, and for each trial, it randomly selects a float from in between the two values to use for the given hyperparameter.
-Grid search instead takes in a list of floats as a parameter and runs a trial for each value in the list.
-Multiple grid searches can easily take a large amount of time to run as all possible combinations of values across all grid searches will be tested in separate trials.
-
-Our approach was to begin with random search on parameters to get a sense of what range of values produced optimal results.
-Once we had established the magnitude of which we wanted our parameters to be, we tested out various combinations with grid search, keeping the ones that produced promising results and replacing the others with new trials.
-
-\subsection{Comparison with Current Practices}
+In this work, we are using weighted F1 score as our method of evaluation.
+As weighted F1 score is more suited when predicting the majority class and we expect most conversations to contain the same common emotions, a higher correctness on these emotions are more important than the more rare ones.
+
+\chapter{Approach}
+
+\section{Common Practices}
 When deciding what configuration to use for our experiments, we based our decisions on the empirical studies made by  \citet{bouthillier:hal-02447823}.
@@
 
-\subsection{Equipment}
-As for our hardware setup, we could not run and train the model on our laptops as they lack a GPU.
-To execute our code, we instead utilized Google Colab, a cloud-based service where users can remotely run their code on Google servers built with more powerful hardware and can handle high-demand programs such as machine learning models.
-Using a remote service proved to be perfectly fine for three of the four corpora we aimed to run our model on.
-The fourth one, \textit{DailyDialog}, unfortunately, required more memory than we were allocated and could not complete the run.
-
-%Hur vi "räknade ut" våra resultat, 10 avg på best valid ->( test)
-%Epochs
-%Google Colab
-%Ray Tune för att studera hyperparametrar och automatisera
-%HAL - diskussion
-
-%How many sets of hyperparameters we evaluated? Vi hade i snitt 20:ish per dataset
-%4 dataset for the experiment
-%Reported 10 samples as one averge.
-
-%Stycke där vi jömför med HAL
+%Snacka mer om varje individuell fråga på enkäten
 
@@
 
-%Lack of resources
-
-%Hyperparametrization with Tune
-%Learning rate
-%Dropout
-%Batch_size [RAM limitations]
-%Definitioner finns redan i terminologi
 \section{Corpora}
@@
 Using Joey's first line in the MELD excerpt as an example, \textit{Joy} might be considered an exaggerated description considering the utterance.
+
+\chapter{Implementation}
+%Summary of what the chapter is about
+This chapter is mainly focused on giving a clear description of what we have done for the project's duration.
+Firstly we will describe the model that we used as a baseline, \textit{COSMIC}.
+We will cover our model \textit{CASTLE}, its implementation, and what changes were made from COSMIC.
+Lastly, we will describe our setup and what corpora we used.
+%??In your text, you present COSMIC first. Change your intro so that it corresponds to your text??.
+
+We based our research strategy on the work by \citet{bouthillier:hal-02447823} that summarizes and reviews the experimental procedures that are most commonly used in the field of machine learning.
+
+%?? You could write a bit more on COSMIC ??
+
+\section{Ecosystem}
+As for our hardware setup, we could not run and train the model on our computers as they lack a GPU.
+To execute our code, we instead utilized Google Colab, a cloud-based service where users can remotely run their code on Google servers built with more powerful hardware and can handle high-demand programs such as machine learning models.
+Using a remote service proved to be perfectly fine for three of the four corpora we aimed to run our model on.
+The fourth one, \textit{DailyDialog}, unfortunately, required more memory than we were allocated and could not complete the run.
+
+\section{Baseline -- COSMIC}
+\label{COSMIC_chapter}
+When trying to improve existing results in many machine learning fields, a baseline is used instead of starting from scratch.
+A baseline is an already functioning model that provides basic results to the problem at hand.
+This provides the opportunity to allocate resources on improving the model instead of building it from the ground up and allows the machine learning community to expand on and draw benefits from each other's work.
+
+This project has been developed using the COSMIC model \citep{ghosal2020cosmic} as a baseline.
+COSMIC is a GRU-based model implemented with PyTorch that has achieved state-of-the-art results across four corpora commonly used in emotion recognition.
+The idea behind the model is that when classifying an utterance, the emotions of the utterances, both preceding and succeeding, are also taken into account.
+This way, the context of the conversation can be used to predict the emotions of individual utterances more accurately.
+
+%COMSICs arkitektur bygger på att de skapar egna GRU-celler och GRU-lager som bygger på PyTorchs respektive. 
+%Filip
+
+Figure \ref{fig:COSMIC_arch} provides a simplified description of how COSMIC operates in a conversation between person A and person B.
+For each time step $t$, there is an utterance said by either A or B with the other person as a listener.
+At the next time-step $t+1$, the roles are altered depending on who utters the next utterance.
+For instance, if Person A were to speak multiple utterances in a row, they would remain the speaker until Person B spoke.
+In cases where there are more than two participants in the conversation, the same figure applies but with multiple listeners.
+For each time-step, there exists only one speaker while all other participants are listeners.
+For each utterance, multiple GRU-cells exist that predict both the speaker's intent and listener's response at that certain time-step.
+These cells are used to predict the emotion of the current utterance and the prediction of emotions in other utterances related to said utterance.
+
+These GRU-cells correspond to five different states to determine what emotion to predict.
+These states are shown in the bottom right of Figure \ref{fig:COSMIC_arch} denoted by a color that shows where the state is used in the model.
+Crucial to these states and the contextual emotion recognition of COSMIC is also the Commonsense feature vector created by COMET \citep{Bosselut2019COMETCT}.
+This model is trained on multiple commonsense knowledge graphs in order to perform automatic knowledge base construction.
+The relevant features being used are \textit{Intent of speaker}, \textit{Effect of speaker}, \textit{Reaction of speaker}, \textit{Effect of listeners}, and \textit{Reaction of listeners}.
+Out of the five states mentioned earlier, four of them are heavily interconnected, taking each other as well as parts of the commonsense vector into account when being updated.
+For all states, the previous value of the state is also taken into account.
+\begin{itemize}
+    \item \textit{Context state} is a state shared between speaker and listener.
+It manages the current overall emotional state of the conversation and can be observed as the green cell in between the parties in the figure.
+The state takes input from the speaker at each time-step.
+This input is a vector based on the current utterance as well as the Internal and External state from the previous time-step.
+
+    \item \textit{Internal state} tracks for each participant what emotions they are feeling internally.
+This could remain hidden, depending on whether or not a participant explicitly displays their emotions.
+For the speaker, the internal state is updated using an \textit{attention vector}, formed from the current utterance and previous context state, as well as \textit{effect on speaker} from the commonsense vector.
+For all other participants, \textit{effect on listener} is used instead of \textit{effect on speaker}.
+This state exists as a the blue cell for all participants.
+
+    \item \textit{External state} portrays, for each participant, what emotions they are expressing to the other participants through both utterances and reactions.
+As this state can be observed and understood by other participants, this state can be reacted to, which is taken into account when updating its value.
+For the speaker, the attention vector and \textit{reaction of speaker} are used to update the value.
+For all listeners, \textit{reaction of speaker} is replaced with \textit{reaction of listener}.
+This state exists as the red cell for all participants
+
+    \item \textit{Intent state} is a state only present for the current speaker.
+As the name suggests, it represents the speaker's objective is for the current utterance.
+For example, if paying the listener a compliment, the intent could be to make the listener happier.
+Its value is updated by using the current internal state of the same participant as well as \textit{intent of speaker}.
+This state exists as the orange cell only for the speaker.
+
+\end{itemize}
+Internal state and External state are especially codependent as they combine into the sum of a participant's emotional state and are the only states updated for listeners.
+
+\emph{Emotion-rep}, or \emph{Emotion state}, the final state, considers the current utterance, the summed emotional state through the Internal and External state, and the Intent state to predict the current utterance.
+The state exists as the purplee state only for the speaker.
+This calculation is repeated for every utterance in every conversation in the corpus.
+
+%Softmax, relu mm
+
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[scale=0.3]{msccls/explanatory_images/COSMIC_arch.png}
+    \caption{COSMIC architecture. From \citet{ghosal2020cosmic}.}
+    \label{fig:COSMIC_arch}
+\end{figure}
+
+\section{CASTLE}
+Building on COSMIC, we designed CASTLE, COSMIC-based AnalySer Taking advantage of LSTM for Emotion recognition, to answer our research questions.
+We have made several improvements on the COSMIC baseline, one of the fundamental being replacing the internal GRU architecture with an LSTM-based architecture, as the name of the model indicates.
+This is greater change than it might seem at first glance.
+Both the GRU layer in COSMIC and the LSTM layer in CASTLE contain 768 of their corresponding cells.
+When comparing the LSTM cell to the GRU cell, as can be observed in \ref{LSTM} and \ref{GRU}, we notice that the LSTM cell has two more gates and one more computation than the GRU cell.
+As we have 768 cells, this is clearly a considerable adjustment.
+%An LSTM architecture has, as mentioned in section \ref{GRU}, a higher number of parameters.
+
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[scale=0.3]{msccls/explanatory_images/CASTLE_arch.png}
+    \caption{CASTLE architecture}
+    \label{fig:CASTLE_arch}
+\end{figure}
+
+Figure \ref{fig:CASTLE_arch} demonstrates the timeline of CASTLE, from an utterance to an emotion.
+Initially the data is extracted from the corpora and divided into the train, dev, and test sets.
+RoBERTa is then used together with this data in order to create a word embedding for the model.
+All of this is only pre-processing before any actual training is done.
+CASTLE is also constructed with both \textit{batch normalization} layers \citep{ioffe2015batch} and \textit{layer normalization} layers \citep{Ba2016LayerN}.
+\citet{Ba2016LayerN} further states that the batch normalization layers use the distribution of the summed input to a neuron over a mini-batch of training cases to compute a mean and variance which are then used to normalize the summed input to that neuron on each training case.
+On the other hand, the layer normalization layers are constructed by transposing the batch normalization layer normalization by computing the mean and variance used for normalization from all of the summed inputs to the neurons in a layer on a single training case.
+These two types of layers are used in order to speed up the training process.
+Following the normalization, there is one LSTM layer with 768 LSTM cells as well as two layers with five LSTM-cells each.
+These five cells are corresponding to the five states explained in chapter \ref{COSMIC_chapter}.
+In conjunction with the LSTM layer there occurs a dropout that is set as a hyperparameter.
+The dropout, as mentioned previously, disregards a percentage of the information learnt in order to combat over-reliance on the training data.
+CASTLE takes use of a \textit{linear} layer, also known as a fully connected network.
+A linear layer is a layer that is used to generalize input \citep{TensorflowDeep}, which means that the model does not need any special assumptions about the input, such as whether it consists of text or audio.
+Finally, a \textit{softmax} layer is used and is built upon the softmax function \citep{boltzmann1868studien}.
+The role of the softmax layer is to distribute the output of the network into k number of classes \citep{hess2020softmaxbased}, in the case of CASTLE, the different emotions that are represented in the corpora.
+
+Even though CASTLE and COSMIC have the same amount of hyperparameters, they still differ in how much they affect the model.
+When using the same values for the hyperparameters in CASTLE that produced state-of-the-art results for COSMIC, CASTLE performed worse than post-optimization.
+COSMIC is also implemented with a \textit{residual} connection \citep{he2015deep}, which allows the model to redefine the layers as learning residual functions referencing the layer inputs.
+This is supposed to facilitate the training of the model.
+While we originally wanted to implement this feature in CASTLE, there was simply no space to do it within the project's scope.
+
+%We chose COSMIC as our baseline because SOTA, more experienced in PyTorch
+\section{Hyperparameter optimization}
+\label{sect:hyper_opt}
+
+%The ranges used was X-Y for all corpora
+%Learning rate var: XXX
+%Dropout var: XXX
+%Batch size var: XXX
+
+A large part of our work has been focused on optimizing hyperparameters for optimal results.
+As stated by \citet{KOHAVI1995304}, different hyperparameter configurations tend to work best for different corpora which changes with model.
+This means that not only do we have to optimize in order to find the combination of hyperparameters that produce the best results for CASTLE, this hyperparameter optimization must be done separately for each corpus.
+To optimize our hyperparameters, we implemented the \texttt{Tune} framework offered by \texttt{Ray} \citep{liaw2018tune}.
+This framework allowed us to experiment efficiently with many different configurations and provided a simple way to run trials concurrently.
+Most of our optimization was done with a combination of grid search and random search.
+Random search is a function with two parameters, \textit{start} and \textit{end}, and for each trial, it randomly selects a float from in between the two values to use for the given hyperparameter.
+Grid search instead takes in a list of floats as a parameter and runs a trial for each value in the list.
+Multiple grid searches can easily take a large amount of time to run as all possible combinations of values across all grid searches will be tested in separate trials.
+
+Our approach was to start with random search on our chosen hyperparameters to get a sense of what range of values produce optimal results.
+\citet{bergstra2012random} shows that randomly chosen trials are more efficient than grid search at traversing the search space.
+Once we had established these initial values for what we wanted our hyperparameters to be, we tested out various combinations with grid search, keeping the ones that produced promising results and replacing the others by running new random searches.
+
+%Hur vi "räknade ut" våra resultat, 10 avg på best valid ->( test)
+%Epochs
+%Google Colab
+%Ray Tune för att studera hyperparametrar och automatisera
+%HAL - diskussion
+
+%How many sets of hyperparameters we evaluated? Vi hade i snitt 20:ish per dataset
+%4 dataset for the experiment
+%Reported 10 samples as one averge.
+
+%Stycke där vi jömför med HAL
+
+%Lack of resources
+
+%Hyperparametrization with Tune
+%Learning rate
+%Dropout
+%Batch_size [RAM limitations]
+%Definitioner finns redan i terminologi
 
@@
 EmoryNLP was the only corpus that we managed to run CASTLE on without achieving better results.
-
 There might be several reasons for this, but we believe that one of them is more likely than the others.
+As mentioned in \nameref{sect:hyper_opt}, each corpus needed its own hyperparameter optimization.
 As EmoryNLP was the last corpus that we optimized CASTLE for, there was not enough time to thoroughly tune the hyperparameters.
-The method we used for optimizing hyperparameters was to initially use random search to find relevant sets of parameters, where the leading values were then iterated through with the help of grid search.
-
-Working in this way is somewhat of a scattershot approach, where we depend unpleasantly much on the results of random searches.
+
+%Working in this way is somewhat of a scattershot approach, where we depend unpleasantly much on the results of random searches.
 While the random searches clearly produced competitive results, there is nothing to indicate whether an optimal set of values was overlooked.
@@
 
-The evidence demonstrates that CASTLE exceeded the best F1-scores on the MELD and IEMOCAP corpora while remaining competitive on EmoryNLP and with inconclusive results for DailyDialog.
-This fulfills our thesis' primary aim that we defined in section \ref{research_questions} and makes CASTLE one of the most viable systems to adapt for practical emotion recognition applications.
+The evidence demonstrates that CASTLE exceeded the best F1-scores on the MELD and IEMOCAP corpora while remaining competitive on EmoryNLP and with non-existent results for DailyDialog.
+This fulfills our thesis' primary aim that we defined in section \ref{research_questions} and makes CASTLE one of the most viable systems to adopt for practical emotion recognition applications.
 
@@
 Since LSTM is a more complex structure with an additional gate and having the concept of utilizing memory in a cell state, it is no surprise that it can outperform GRU layers.
-When debating if to use CASTLE or some other model, it would boil down to whether an increase in certainty is worth the computational power cost.
+It is common for results in machine learning to incrementally become better by building on top of each other.
+Similarly our results, albeit just slightly better than the previous state of the art, is the next in a long line of continuous improvements that slowly but surely improve results.
 
@@
 One of these rules is that no notebook can run for more than 24 hours straight.
-Hyperparameter optimization is a time-consuming task that is solved by trial and error, and this restriction meant that we could not run as many trials at a time as we wanted, which in turn led to less optimization as there was not enough time.
+Hyperparameter optimization is a time-consuming task that is solved by trial and error, and this time restriction meant that we could not run as many trials at a time as we wanted, which in turn led to less optimization as we did not have as high an uptime as we potentially could.
 
